---
title: "Second-order Probabilism: Expressive Power and Accuracy"
author: "Rafal Urbaniak and Marcello Di Bello"
date: '`r Sys.Date()`'
format:
  pdf:
    toc: true
    include-in-header: quartoStyle.sty
    mainfont: Times New Roman
    sansfont: Times New Roman
documentClass: scrartcl
classOptions: [dvipsnames, enabledeprecatedfontcommands, headings=big]
font:
  size: 10pt
url:
  color: blue
bibliography: [../../../references/referencesMRbook.bib]
csl: apa-6th-edition.csl
indent: true
numbersections: true
---

  <!-- pdf:
    numberSections: true
    dfPrint: kable 
    keepTex: true
    toc: true
    include-in-header:  quartoStyle.sty
    latex_engine: pdflatex  -->


```{r setup}
#| include: false

knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
library(rethinking)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(tidyr)
library(philentropy)
library(latex2exp)
library(gridExtra)
library(rethinking)
library(bnlearn)
library(gRain)
library(reshape2)
library(truncnorm)
library(ggforce)
library(dplyr)



ps <- seq(0,1, length.out = 1001)
getwd()
source("../../scripts/CptCreate.R")
source("../../scripts/SCfunctions.R")

SCprobsFinal <- readRDS("../../datasets/SCprobsFinal.rds")
attach(SCprobsFinal)      
source("../../scripts/SCplotCPTs.R")
source("../../scripts/SCplotDistros.R")

```




\vspace{2cm}

\noindent \textbf{DISCLAIMER:} \textbf{This is a draft of work in progress, please do not cite or distribute without permission.}

\thispagestyle{empty}

\newpage

\begin{quote} \textbf{Abstract.}  Rational agents are often uncertain about the truth of many propositions. To represent this uncertainty, it is natural to rely on probability theory. Two options are typically on the table, precise and imprecise probabilism, but both fall short in some respect. Precise probabilism is not expressive enough, while imprecise probabilism suffers from belief inertia and the impossibility of proper scoring rules. We put forward a novel version of probabilism, higher-order probabilism, and we show that it outperforms existing alternatives. 

\end{quote}

\todo{check abstract and intro}

# Introduction
\label{sec:introduction}


<!--we form beliefs about a variety of propositions on the basis of the evidence available to us. But believing a proposition is not an all-or-nothing affair; it is a matter of degrees. -->


 As rational agents, we are uncertain about the truth of many propositions since the evidence we possess about them is often fallible. To represent this uncertainty, it is natural to rely on probability theory. Two options are typically on the table: precise and imprecise probabilism. Precise probabilism models an agent's state of uncertainty (or credal state) with a single probability measure: each proposition is assigned one probability value between 0 and 1 (a sharp credence).  The problem is that a single probability measure is not expressive enough to distinguish between intuitively different states of uncertainty rational agents may find themselves in (\S \ref{sec:precise-probabilism}).  To avoid this problem, a *set* of probability measures, rather than a single one, can be used to represent the uncertainty of a rational agent. This approach is known as imprecise probabilism. It outperforms precise probabilism in some respects, but also runs into problems of its own, such as belief inertia and the impossibility of defining proper scoring rules (\S \ref{sec:imprecise-probabilism}). 
 
 To make progress, this paper argues that the uncertainty of a rational agent is to be represented neither by a single probability measure nor a set of  measures. Rather, it is to be represented by a higher-order probability measure, more specifically, a probability distribution over parameter values intepreted as probabilities.  <!---The key insight is that a rational agent's uncertainty (or credal state) is not single-dimensional and thus cannot be mapped onto a one-dimensional scale like the real line. Uncertainty is best modeled by the shape of a probability distribution over multiple probability measures. --->  The theory we propose is not mathematically novel, but addresses many of the problems that plague both precise and imprecise probabilism (\S \ref{sec:higher-order} and \S \ref{sec:proper-scores}). It also fares better than existing versions of probabilism when the probability of multiple propositions, dependent or independent, is to be assessed (\S \ref{sec:higher-order-conjunction} and \ref{sec:higher-order-networks}). 
 
 ]
 
 <!--
Moreover, Bayesian probabilistic programming already provides a fairly reliable implementation framework of this approach.
-->

<!---(1) It is not sufficiently evidence-responsive; (2) it cannot model certain intuitively plausible comparative probability judgments; (3) nor can it model learning when the starting point is complete lack of information; and finally, (4) no inaccuracy measure of an imprecise credal stance exists which satisfies certain plausible formal conditions. 
--->

<!-- Moreover, while it seems to handle some cases of opinion pooling better than PP, it still can't capture the phenomenon of synergy, where slightly disagreeing sources or experts jointly in some sense seem to improve the epistemic situation. -->

<!---# Precise vs. imprecise probabilisms
\label{sec:three-probabilism}
--->

# Precise probabilism
\label{sec:precise-probabilism}

Precise probabilism holds that a rational 
agent's uncertainty about a proposition is to be 
represented as a single, precise probability measure. 
Bayesian updating regulates how the prior probability measure should change in light of new evidence that the agent learns. The updating can be iterated multiple times for multiple pieces of evidence considered successively.
This is an elegant and simple theory with 
many powerful applications. Unfortunately, representing our uncertainty about a proposition in terms of a single, precise probability measure runs into a number of difficulties. 

Precise probabilism fails to 
capture an important dimension of how 
our fallible beliefs reflect the evidence 
we have (or have not) obtained. A couple of stylized 
examples featuring coin tosses should make the point clear. 
Herer is the first:

\begin{quote}
\textbf{No evidence v. fair coin}
You are about to toss a coin, but have no evidence 
about its bias. You are completely ignorant. 
Compare this to the situation in which you know, 
based on overwhelming evidence, that the coin is fair. 
\end{quote}

\noindent On precise probabilism, both scenarios are represented by assigning a probability of .5 to the outcome \emph{heads}. If you are completely 
ignorant, the principle of insufficient evidence suggests that you assign .5 to both outcomes.  Similarly, if you know for sure the coin is fair, assigning .5 seems the best way to quantify the uncertainty about the outcome. The agent's evidence in the two scenarios is quite different, but precise probabilities fail to  capture this difference. 

And now consider a  second scenario:
\begin{quote}
\textbf{Learning from ignorance}
You toss a coin with unknown bias. You toss it 10 times and observe \emph{heads} 5 times. Suppose you toss it further and observe 50 \emph{heads} in 100 tosses. 
\end{quote}

\noindent
Since the coin initially had unknown bias, you should 
presumably assign a probability of .5 to both outcomes if you stick with precise probabilism. 
After the 10 tosses, you again assess the probability to be .5.
You must have learned something, but whatever that is, it is not modeled by precise probabilities. When you toss the coin 100 times and observe 50 heads, you learn something new as well. But your precise probability assessment will again be .5.

These examples suggest that precise probabilism is not appropriately responsive to evidence. Representing an agent's uncertainty by a precise probability measure can fail to track what an agent has learned from new evidence. Precise probabilism assigns the same probability in situations in which one's evidence is quite different:  when no evidence is available about a coin's bias; when there is little evidence that the coin is fair (say, after only 10 tosses); and when there is strong evidence that the coin is fair (say, after 100 tosses). In fact, analogous problems also arise for evidence that the coin is not fair. Suppose the rational agent starts with a weak belief that the coin is  .6 biased towards heads. They can strengthen that belief by tossing the coin repeatedly and observing, say, 60 heads in 100 tosses. But this improvement in their evidence is not mirrored in the .6 probability they are supposed to assign to *heads*.^[Here is another problem for precise probabilism. Imagine a rational agent who does not know the bias of the coin. For precise probabilism, this state of uncertainty should be represented by a .5 probability assignment to the *heads*. Next, the agent learns that the bias towards heads, whatever the bias is, has been slightly increased, say by .001. The addition of this new information is called \emph{sweetening}
in the philosophical literature. This sweetening should now make the agent bet on heads: if the probability of *heads* was initially .5, it must now be slightly above .5. But, intuitively, the new information should leave the agent equally undecided about betting on heads or tails. After sweetening, the agent still does not know much about the actual bias of the coin.]
\todo{add reference about sweetening in footnote}

<!-- 
The general problem is, precise probability captures the value around which your uncertainty should be centered, but fails to capture how centered it should be given the evidence.
-->


 
 

 <!-- ^[Precise probabilism suffers from other difficulties. For example,  it has problems  with formulating a sensible method of probabilistic opinion  aggregation [@Elkin2018resolving,@Stewart2018pooling].  A seemingly intuitive constraint is that if every member agrees that $X$ and $Y$ are probabilistically independent, the aggregated credence should respect this. But this is hard to achieve if we stick to \s{PP} [@Dietrich2016pooling]. For instance, a \emph{prima facie} obvious method of linear pooling does not respect this. Consider probabilistic measures $p$ and $q$ such that $p(X)  = p(Y)  = p(X\vert Y) = 1/3$ and  $q(X)  =  q(Y) = q(X\vert Y) = 2/3$. On both measures, taken separately, $X$ and $Y$ are independent. Now take the average, $r=p/2+q/2$. Then $r(X\cap Y) = 5/18 \neq r(X)r(Y)=1/4$. This inability to capture an important epistemological difference, the impreciser insists, is a serious limitation. Intuitively, a precise stance is not justified by the very scant evidence available.] -->

These problems generalize beyond cases of coin tossing. It is one thing not to know much about whether a proposition is true, for example, whether
an individual is guilty of a crime. It is another thing to have strong evidence that favors a hypothesis and equally strong evidence that favors its negation, for example, strong evidence favoring the guilt hypothesis and equally strong evidence favoring the hypothesis of innocence. Despite this difference, precise probabilism would reccomend that a probability of .5 be assigned to both hypotheses in either case. Here, too, precise probabilities fail to be appropriately responsive to the evidence. 

In addition, evidence can accumulate in a way that does not require changing our initial probability assignments. Suppose that, at first, one's overall evidence favors $A$ over $B$. So the probability assigned to $A$ should be greater than that assigned to $B$. Next, the agent acquires new evidence. The total quantity of evidence has increased, but suppose this larger body of evidence overall still favors $A$ over $B$. So no change in the probabilities seems required. Still, something has changed about the agent's state of uncertainty towards $A$ and $B$: the quantity of evidence on which the agent can make their assessment whether $A$ is more probable than $B$ has become larger. And yet, this change in the quantity of overall evidence is not reflected in the precise probabilities assigned to the propositions $A$ and $B$.^[The distinction here is sometimes formulated in terms of the *balance* of the evidence (that is, whether the evidence available tips in favor a hypothesis or another) as opposed to its *weight* (that is, the overall quantity of evidence regardless of its balance); see @keynes1921treatise and @joyce2005probabilities among others.]



# Imprecise probabilism
\label{sec:imprecise-probabilism}

What if we give up the assumption that probability assignments should be precise? Imprecise probabilism holds that  a rational agent's credal stance towards a hypothesis is to be represented by a set of probability measures, typically called a  representor $\mathbb{P}$, rather than a single measure $\mathsf{P}$. The representor should include all and only those probability measures which are compatible
with the evidence (more on this point later).^[For  the development of imprecise probabilism, see @keynes1921treatise; @Levi1974ideterminate; @Gardenfors1982unreliable; @Kaplan1968decision;  @joyce2005probabilities; @VanFraassen2006vague; @Sturgeon2008grain; @walley1991statistical. @bradley2019imprecise is a good source of further references. Imprecise probabilism shares some similarities with what we might call interval probabilism [@Kyburg1961; @kyburg2001uncertain]. On interval probabilism, precise probabilities are replaced  by intervals of probabilities. On imprecise probabilism, instead, precise probabilities are replaced by sets of probabilities.  This makes imprecise probabilism more general, since the probabilities of a proposition in the representor set do not have to form a closed interval.] It is easy to see that modeling an agent's credal state by sets of probability measures avoids some of the shortcomings of precise probabilism.  For instance, if an agent knows that the coin is fair, their credal state  would  be represented by the singleton set $\{\mathsf{P}\}$, where $\mathsf{P}$ is a probability measure that assigns $.5$ to \emph{heads}. If, on the other hand, the agent knows nothing about the coin's bias, their credal state would be represented by the set of all probabilistic measures, since none of them is excluded by the available evidence. Note that the set of probability measures does not represent admissible options that the agent could legitimately pick from. Rather, the agent's credal state is essentially imprecise and should be represented by means of the entire set of probability measures.


So far so good. But, just as precise probabilism fails to be appropriately 
evidence-responsive in certain scenarios, imprecise probabilism runs in similar difficulties 
in other scenarios. 

\begin{quote}
\textbf{Even v. uneven bias:}
 You have two coins and you know, for sure, that the probability of getting heads is .4, if you toss one coin, and .6, if you toss the other coin. But you do not know which is which. You pick one of the two at random and toss it.  Contrast this with an uneven case. You have four coins and you know that three of them have bias $.4$ and one of them has bias $.6$. You pick a coin at random and plan to toss it. You should be three times more confident that the probability of getting heads is .4. rather than .6.
\end{quote}

\noindent
The first situation can be easily represented by imprecise probabilism. The representor would contain two probability measures, one that assigns .4. and the other that assigns .6 to the hypothesis 'this coin lands heads'.  But imprecise probabilism cannot represent the second situation. Since the probability measures in the set are all compatible with the agent's evidence, no probability measure can be assigned a greater (higher-order) probability than any other.^[Other scenarios can be constructed in which imprecise probabilism fails to capture distinctive intuitions about evidence and uncertainty; see, for example, [@Rinard2013against].  Suppose you know of two urns, \textsf{GREEN} and \textsf{MYSTERY}. You  are  certain \textsf{GREEN} contains only green marbles, but have no information about \textsf{MYSTERY}. A marble will be drawn at random from each. You should be certain that the marble drawn from \textsf{GREEN}  will be green ($G$), and you should be more confident about this than about the proposition that the marble from \textsf{MYSTERY} will be green ($M$). For each $r\in [0,1]$  your representor will contain a $\mathsf{P}$ with $\pr{M}=r$. But then,  it also contains one with $\pr{M}=1$. This means that it is not the case that for any probability measure  $\mathsf{P}$ in the representor set, $\mathsf{P}(G) > \mathsf{P}(M)$, that is, it is not the case that a rational agent should be more confident of $G$ than of $M$. This is counter-intuitive.] 

These examples show that imprecise probabilism is not expressive enough to model the scenario of uneven bias. <!--We believe that the solution is to add a probability distribution over the probability measures (more on this in the next section).--> Defenders of imprecise probabilism might concede this point but prefer their account for reasons of simplicity. They could also point out that imprecise probabilism models scenarios that precise probabilism cannot model, for example, a state of complete lack of evidence. In this respect, imprecise probabilism outperforms precise probabilism in expressive power, and also retains theoretical simplicity. But this is not quite true as imprecise probabilism suffers from several shortcomings that do not affect precise probabilism.  

The first shortcoming has not received extensive discussion in the literature, but it is fundamental. Recall that, for imprecise probabilism, an agent's state of uncertainty is represented by those probability measures that are *compatible* with the agent's evidence. The question is, how should the notion of compatibility be understood here? <!---The idea is that thanks to this feature, imprecise credal stances are evidence-responsive in a way precise probabilistic stances are not.---> Perhaps we can think of compatibility as the fact that the agent's evidence is consistent with the probability measure in question. But mere consistency wouldn't get the agent very far in excluding probability measures, as too many probability measures are consistent with most observations and data. Admittedly, there will be  clear-cut cases: if you see the outcome of a coin toss to be heads, you reject the measure with $\mathsf{P}(H)=0$, and similarly for tails. Another class of cases might arise while randomly drawing objects from a finite set where the objective chances are known. But clear-cut cases aside, what else?  Data will often be consistent with almost any 
probability measure.^[Probability measures can be inconsistent with evidential constraints that agents believe to be true. Mathematically, non-trivial evidential constraints are easy to model [@bradley2012scientific]. They can take the form, for example, of the \emph{evidence of chances} $\{ \mathsf{P}(X) = x\}$ or $\mathsf{P}(X) \in [x,y]$, or \emph{structural constraints} such as  "$X$ and $Y$ are independent" or "$X$ is more likely than $Y$." These constraints are something that an agent can come to accept outright, but only if offered such information by an expert whom the agent completely defers to. <!---Most of the  examples  in the literature start with the assumption that the agent is told by a believable source that the chances are such-and-such, or that certain structural constraint hold.--->  Besides these idealized cases, it is unclear how an agent could come to accept such structural constraints upon observation. There will usually be some degree of uncertainty about the acceptability of these constraints.]


<!---
Bradley suggests that "statistical evidence might inform [evidential] constraints (\dots and that evidence) of causes might inform structural constraints" (125-126). This, however, is not a clear account of how exactly this should proceed. One suggestion might be that once a statistical significance threshold is selected, a given set of observations with a selection of background modeling assumptions yields a credible interval. But this is to admit that to reach such constraints, we already have to start with a second-order approach, and drop information about the densities, focusing only on the intervals obtained with fixed margins of errors. But as we will be insisting, if you have the information about densities to start with, there is no clear advantage to going imprecise instead, and there are multiple problems associated with this move. Moreover, such moves require a choice of an error margin, which is extra-epistemic, and it is not clear what advantage there is to  use extra-epistemic considerations of this sort to drop information contained in densities.^[Relatedly, in forensic evidence evaluation even scholars who disagree about the value of going higher-order agree that interval reporting is problematic, as the choice of a limit or uncertainty level is rather arbitrary  [@Taroni2015Dismissal;@Sjerps2015Uncertainty].] 
--->

A second, related problem for imprecise probabilism is known as belief inertia. Precise probabilism offers an elegant model of learning from evidence: Bayesian updating. Imprecise probabilism,  at least \emph{prima facie}, offers an equally elegant model of learning from evidence, richer and more nuanced. It is a natural extension of the classical Bayesian approach that uses precise probabilities. When faced with new evidence $E$ between time $t_0$ and $t_1$, the representor set should be updated point-wise,  running the  standard Bayesian updating on each probability 
measure in the representor:
\begin{align*} \label{eq:updateRepresentor}
\mathbb{P}_{t_1} = \{\mathsf{P}_{t_1}\vert \exists\, {\mathsf{P}_{t_0} \!\in  \mathbb{P}_{t_0}}\,\, \forall\, {H}\,\, \left[\mathsf{P}_{t_1}(H)=\mathsf{P}_{t_0}(H \vert E)\right] \}.
\end{align*}

\noindent
The hope is that, if we start with a range of probabilities that is not extremely wide, point-wise learning will behave appropriately.
<!---^[The hope is also that \s{IP} offers a feasible aggregation method [@Elkin2018resolving;@Stewart2018pooling]: just put all representors together in one set, and voil\'a! However, this is a very conservative method which quickly leads to extremely few points of agreement, and we are not aware of any successful practical deployment of this method.]
--->
For instance, if we start with a prior probability of \emph{heads} equal to .4 or .6, then those measures should be updated to something closer to $.5$ once we learn that a given coin has already been tossed  ten times with the observed number of heads equal 5 (call this evidence $E$). This would mean that if the initial range of values was $[.4,.6]$ the posterior range of values should be narrower.

<!---But even this seemingly straightforward piece of reasoning is hard to model within the confines of imprecise probabilism. For to calculate $\pr{\s{bias} = k \vert E}$ we need to calculate $\pr{E \vert \s{bias} = k }\pr{\s{bias} = k}$ and divide it by $\pr{E} = \pr{E \vert \s{bias} = k }\pr{\s{bias = k}} + \pr{E \vert  \s{bias} \neq k }\pr{ \s{bias} \neq k}$. The tricky part is obtaining  $\pr{\s{bias} = k}$ or  $\pr{ \s{bias} \neq k}$ in a principled manner without explicitly going second-order, without estimating the parameter value and without using beta distributions. 
--->

Unfortunately, this narrowing of the range of values becomes impossible whenever the starting point is complete lack of knowledge, as imprecise probabilism runs into the problem of belief inertia  [@Levi1980enterprise]. This problem arises in situations in which no amount of evidence could lead the agent to change their belief state, according to a given modeling strategy. Consider a situation in which you start tossing a coin  knowing nothing about its bias. The range of possibilities is $[0,1]$. After a few tosses, if you observed at least one tail and one heads, you can exclude the measures assigning 0 or 1 to \emph{heads}. But what else have you learned?  If you are to update your representor set point-wise, you will end up with the same representor set. For any sequence of outcomes that you can obtain and any probability value in $[0, 1],$ there will exist a probability measure (conditional on the oucomes) that assigns that probability to *heads*. Consequently, the edges of your resulting interval will remain the same. In the end, it is not clear how you are supposed to learn anything if you start from complete ignorance.^[Here is another example of belief inertia by 
@Rinard2013against.  Suppose all the marbles in the urn are green ($H_1$), or exactly one tenth of the marbles are green ($H_2$). Your initial credence about these  hypotheses is completely uncertain, the interval $[0,1]$. Next,  you learn that a marble drawn at random from the urn is green ($E$). After conditioning on this evidence, you end up with the same spread of values for $H_1$ that you had  before learning $E$. This result holds  no matter  how many green marbles are drawn. This is counterintuitive: if you keep drawing green marbles, the hypothesis that all marbles are green should become more probable.]

Some downplay the problem of belief inertia.  After all, if you started with knowing truly nothing, then it is right to conclude that you will never learn anything. Joyce (2010) writes:

> You cannot learn anything in cases of pronounced ignorance simply because a prerequisite for learning is to have prior views about how potential data should alter your beliefs (p. 291) [@Joyce2010defense] 

\noindent
The upshot is that uniform priors should not be used and that imprecise probabilism gives the right results when the priors are non-vacuous. \todo{add citation to Moss} \textbf{Moss CITE} arrives at the same conclusion by drawing the following parallelism. If contigent propositions should not be assigned probabilities of 0 or 1 whenver these extreme values are unrevisable, then by the same token  the uniform interval [0, 1] should not be used  for imprecise probabilities whenever it is impervious to revision.^[Another strategy is to say that, in a state of complete ignorance, a special updating rule should be deployed. @Lee2017impreciseEpistemology suggests the rule of  \emph{credal set replacement} that recommends that upon receiving evidence the agent should  drop measures rendered implausible, and add all non-extreme plausible probability measures. This, however, is tricky. One  needs a separate account of what makes a distribution plausible from a principled account of why one should use a separate special update rule when starting with complete ignorance.] The challenge, however, is to explain in a principled manner which types of priors a rational agent is justified in assigning and why. What is the reason not to assign uniform priors except that they are unrevisable and thus cause belief inertia? While it is true that one cannot learn anything in a state of complete ignorance, the scenarios giving rise to belief inertia are not like that. The agent knows that the coin is two-sided, that the bias of the coin does not change from one toss to the next, etc. As we will soon see, uniform priors are not necessarily unrevisable and can be a starting point for learning. This suggests that the problem lies with imprecise probabilism, not with uniform priors as such.



Finally, even setting aside belief inertia,  imprecise probabilism faces a third major problem that does not arise for precise probabilism. As it turns out, it is impossible to define proper scoring rules for measuring the accuracy of a representor. Workable scoring rules exist for measuring the accuracy of a single, precise probability measure, such as the Brier score. These rules measure the distance between a rational agent's probability measure and the actual value. A  requirement of scoring rules is that they  be \emph{proper}: any rational agent will expect their own probability measure to be more accurate than any other. After all, if an agent thought a different probability measure was more accurate, they should switch to it.  Proper scoring rules are then used to formulate accuracy-based arguments for precise probabilism. These arguments show (roughly) that, if your precise measure follows the axioms of probability theory, no other non-probabilistic measure is going to be more accurate than yours whatever the facts are.  Can the same be done for imprecise probabilism? It cannot. Impossibility theorems demonstrate that no proper scoring rules are available for representor sets [@seidenfeld2012forecasting]. So, as many have noted, the prospects for an accuracy-based argument for imprecise probabilism look dim  [@Mayo-Wilson2016scoring; @CampbellMoore2020accuracy]. Moreover, as shown by  @Schoenfield2017accuracy, if an accuracy measure satisfies certain plausible formal constraints, it will never strictly recommend an imprecise stance, as for any imprecise stance there will be a precise one with at least the same accuracy.


# Higher-order probabilism
\label{sec:higher-order}

Let us take stock. Imprecise probabilism is more expressive than precise probabilism. It can model the difference between a state in which there is no evidence about a proposition (or its negation) and a state in which the evidence for and against a proposition is in equipoise. But imprecise probabilism has its own expressive limitations, for example, it cannot model the case of uneven bias. In addition, imprecise probabilism faces difficulties that do not affect precise probabilism: the notion of compatibility between a probability measure and the evidence is too permissive; belief inertia makes it impossible for a rational agent to learn via Bayesian updating; and no proper scoring rules exist for imprecise probabilism. In this section, we show that higher-order probabilism overcomes the expressive limitations of imprecise probabilism without falling prey to any such difficulties.  

Proponents of imprecise probabilism already hinted at the need of relying on higher order-probabilities. For instance, Bradley   compares the measures in a representor to committee members, each voting on a particular issue, say the true chance or bias of a coin. As they acquire more evidence, the committee members will often converge on a chance hypothesis.

> \dots the committee members are "bunching up". Whatever measure you put over the set of probability functions---whatever "second order probability" you use---the "mass" of this measure gets more and more concentrated around the true chance hypothesis. [@bradley2012scientific, p. 157]


\noindent
But such bunching up cannot be modeled by imprecise probabilism alone: a probability distribution over chance hypotheses is needed.^[In a similar vein, @joyce2005probabilities, in a paper defending imprecise probabilism,  explicates the notion of weight of evidence using a probability distribution over chance hypotheses. 
<!---something that imprecise probabilism was advertised to handle better than precise probabilism: weight of evidence. \todo{briefly add quote from Keynes to clarify weight of evidence} The explication uses a density over chance hypotheses and conceptualizes the  weight of evidence  as an increase of concentration of smaller subsets of chance hypotheses. ---> Oddly, representor sets play no central role in Joyce's account of the weight of evidence.] That one should use  higher-order probabilities  has also been suggested by critics of imprecise probabilism. For example, @Carr2020impreciseEvidence argues that sometimes evidence requires uncertainty about what credences to have. Carr, however, does not articulate this suggestion more fully; does not develop it formally; and does not explain how her approach would fare against the difficulties affecting precise and imprecise probabilism. We now set out to do precisely that. 

<!--Such bunching up cannot be modeled by imprecise probabilism alone. A probability distribution over chance hypotheses is needed. Bradley seems to be aware of that, which would explain the use of scare quotes: when he talks about the option of using second-order probabilities in decision theory, he  insists that  'there is no justification for saying that there is more of your representor here or there.' ~[p.~195]--> 

<!--
\todo{Rafal to correct technical mistakes in this paragraph (e.g. concept of distribution over probability measures seems incorrect for our purposes)}
\todo{Revised this bit, take a look.}
-->

 The central idea of higher-order probabilism is this:  a rational agent's uncertainty is not single-dimensional and thus cannot be mapped onto a one-dimensional scale such as  the real line. Uncertainty is best modeled by the shape of a probability distribution, sometimes even over the parameters which are best construed as probabilities themselves. In some straighforward cases of narrow and symmetric distributions we can get away with using point estimates, but such approximations will fail to be useful in more complex cases. Stated more formally, a rational agent's state of uncertainty (or credal stance) towards a proposition $X$ is not represented by a single probability value $\mathsf{P}(X)$ between 0 and 1, but by a probability density $f(\mathsf{P}(X))$, where the first-order probability of $X$ is the parameter in question, and is treated as a random variable.  \todo{Since f is discretized, can we just say distribution?}Crucially, this representation is general. While the examples used so far may not indicate this, the the first-order probability of $X$ is not restricted to chance hypotheses or the bias of a coin. The probability density $f(\mathsf{P}(X))$ assigns a second-order probability (density) to all possible first-order probabilities $\mathsf{P}(X)$.^[For computational ease, we will use a higher-order density that is discretized and ranges over 1000 first-order probabilities.]

<!---^[Bradley admits this much [@bradley2012scientific, 90], and so does  Konek  [@konek2013foundations, 59]. For instance, Konek disagrees with: (1)  $X$ is more probable than $Y$ just in case $p(X)>p(Y)$, (2)  $D$ positively supports $H$ if $p_D(H)> p(H)$, or (3)  $A$ is preferable to $B$ just in case the expected utility of $A$ w.r.t. $p$ is larger than that of $B$.]
---> 

<!---
The reader might be worried. The examples we discussed so far involve estimation of chances or population frequencies; but how are we to conceptualize higher order probabilities in a more general settings when we think of first-order probabilities as RAs degrees of belief?
--->

How should these second-order probabilities be understood? It is helpful to think of higher-order probabilism as a generalization of imprecise probabilism. Imprecisers already admit that some probability measures are compatible and others incompatible with the agent's evidence at some point. Compatibility is a coarse notion; it is an all-or-nothing affair. <!---For instance, suppose there is no evidence about the bias of a coin. Then, any first-order point probability about *heads* would be compatible with the evidence. If, instead, we know the coin is fair, the evidence clearly selects one preferred value, .5, and all other first-order probabilities would be incompatible with the evidence. But often, evidence is stronger than the former case and weaker than the latter case. --->
But, as seen earlier, evidence can hardly exclude a probability measure in a definitive manner except in clear-cut cases. Just as it is often a matter of degrees whether evidence supports a proposition, the notion of compatibility between evidence and probability measures can itself be a matter of degrees. On this picture, the evidence justifies different values of first-order probability to various degrees. So, second-order probabilities express the extent to which the first-order probabilities are supported by the evidence. 


This higher-order approach at the technical level is by no means novel. Bayesian probabilistic programming languages embrace the well-known idea that parameters can be stacked and depend on each other [@Bingham2021PPwithoutTears]. But, while the technical machinery has been around for a while, it has not been deployed by philosophers to model a rational agent's uncertainty or credal state. Because of its greater expressive power, higher-order probabilism can represent uncertainty in a more fine-grained manner, as illustrated in @fig-evidenceResponse. In particular, the uneven coin scenario in which the two biases of the coin are not equally likely---which imprecise probabilism cannot model---can be easily modeled within high-order probabilism by assigning different probabilities to the two biases.
  

```{r evidenceResponse1,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
p <- seq(from=0 , to=1 , by = 0.001)
FairDensity <- ifelse(p == .5, 1, 0)
FairDF <- data.frame(p,FairDensity)
FairPlot <- ggplot(FairDF, aes(x = p, y = FairDensity))+geom_line()+theme_tufte()+
  xlab("parameter value")+ylim(c(0,1))+ylab("probability")+ggtitle("Fair coin")+
  theme(plot.title.position = "plot")

UniformPlot <- ggplot()+xlim(c(0,1))+stat_function(fun = dunif, args = list(min = 0, max = 1))+
  ylim(0,1)+
  theme_tufte()+
  xlab("parameter value")+
  ylab("probability")+theme(plot.title.position = "plot")+ggtitle("Unknown bias")


p <- seq(from=0 , to=1 , by = 0.001)
TwoDensity <- ifelse(p == .4 | p == .6, .5, 0)
TwoDF <- data.frame(p,TwoDensity)
TwoPlot <- ggplot(TwoDF, aes(x = p, y = TwoDensity))+geom_line()+theme_tufte()+
  xlab("parameter value")+ylim(c(0,1))+ylab("probability")+ggtitle("Two biases (insufficient reason)")+
  theme(plot.title.position = "plot")+scale_x_continuous(breaks = c(.4,.6), labels = c(.4, .6))

p <- seq(from=0 , to=1 , by = 0.001)
TwoUnbalancedDensity <- ifelse(p == .4, .75, ifelse(p ==  .6, .25, 0))
TwoUnbalancedDF <- data.frame(p,TwoUnbalancedDensity)

TwoUnbalancedPlot <- ggplot(TwoUnbalancedDF, aes(x = p, y = TwoUnbalancedDensity))+geom_line()+theme_tufte()+
  xlab("parameter value")+ylim(c(0,1))+ylab("probability")+ggtitle("Two unbalanced biases")+
  theme(plot.title.position = "plot")+scale_x_continuous(breaks = c(.4,.6), labels = c(.4, .6))

```

<!---
```{r}
mean(TwoUnbalancedDensity)
mean(TwoDensity)
mean(TwoDensity)*mean(TwoDensity)
mean(TwoDensity*TwoDensity)
mean(TwoUnbalancedDensity*TwoDensity)
mean(TwoUnbalancedDensity)*mean(TwoDensity)

```
--->

```{r FigevidenceResponse2,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "80%"}
#| label: fig-evidenceResponse
#| fig-cap: "Examples of higher-order distributions for a few  scenarios problematic for both precise and imprecise probabilism."
#| fig-pos: t

grid.arrange(FairPlot,UniformPlot, TwoPlot, TwoUnbalancedPlot)
```

An agent's uncertainty could---perhaps, should---sometimes be represented by a sigle probability value. Higher-order probabilism does not prohibit that. For example, there may well be cases in which an agent's uncertainty is aptly represented by the expectation.^[The 
expectation is usually defined as $\int_{0}^{1} x f(x) \, dx$. In the context of our approach here, $x$ is the first-order probability of a given proposition, and $f$ is the density representing the agent's uncertainty about $x$.]
<!---can be used as  the precise, object-level credence in the proposition itself, where $f$ is the probability density over possible object-level probability values.--->  But this need not always be the case. If the probability distribution is not sufficiently concentrated around a single value, a one-point summary will fail to do justice to the nuances of the agent's credal state.^[This approach lines up with common practice in Bayesian statistics, where the primary role of uncertainty representation is assigned to the whole distribution. Summaries such as the mean, mode standard deviation,  mean absolute deviation, or highest posterior density intervals are only succinct ways for representing uncertainty.] For example, consider again the scenario in which the agent knows that the bias of the coin is  either .4 or .6 but the former is three times more likely. Representing the agent's credal state with the expectation $.75 \times .4 + .25 \times .6 = .45$ would fail to capture the agent's different epistemic attitudes towards the two biases. The agent believes the two biases have different probabilities, and is also certain the bias is *not* .45. 


Besides its greater expressive power in modelling uncertainty, higher-order probabilism does not fall prey to belief inertia. Consider a situation in which you have no idea about the bias of a coin.  You start with a uniform distribution over $[0,1]$ as your prior. Observing  any non-zero number of heads will exclude 0 and observing any non-zero number of  tails will exclude 1 from the basis of the posterior. The posterior distribution will become more centered as the observations come in. This result is a straightforward application of Bayesian updating. Instead of plugging sharp probability values into the formula for Bayes's theorem , the factors to be mutiplied in the theorem will be probability densities (or ratios of densities as needed). @fig-intertia2 illustrates---starting with a uniform prior distribution---how the posterior (beta) distribution changes after successive observations.^[Assuming independence and constant probability for all the observations, learning is modeled the Bayesian way. You start with some prior density $p$ over the parameter values. If you start with complete lack of information, $p$ should be uniform. Then, you observe the data $D$ which is the number of successes $s$ in a certain number of observations $n$. For each particular possible value $\theta$ of the parameter, the probability of $D$ conditional on $\theta$ follows the binomial distribution. The probability of $D$ is obtained by integration. That is:
\begin{align*}
p(\theta \vert D) & = \frac{p(D\vert \theta)p(\theta)}{p(D)}\\
& = \frac{\theta^s (1-\theta)^{(n - s)}p(\theta)}{\int (\theta')^s (1-\theta')^{(n - s)}p(\theta')\,\, d\theta'}.
\end{align*}
] 

```{r Figinertia2, eval = TRUE, echo=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message= FALSE, warning=FALSE}

n <- 1000 #parameters 
s <- 1e5  #sample size

ps <- seq(from=0 , to=1 , length.out=n)

prior <- rep(1/n , n) #uniform prior

likelihood1g <- dbinom( 1 , size=1 , prob=ps)
likelihood2g <- dbinom( 2 , size=2 , prob=ps)
likelihood2g1b <- dbinom( 2 , size=3 , prob=ps)

posterior1g <- likelihood1g * prior
posterior2g <- likelihood2g * prior
posterior2g1b <- likelihood2g1b * prior


posterior1g <- posterior1g / sum(posterior1g)
posterior2g <- posterior2g / sum(posterior2g)
posterior2g1b <- posterior2g1b / sum(posterior2g1b)

upperLimit <- .003

InertiaPriorPlot <- ggplot()+geom_line(aes(x = ps, y = prior))+theme_tufte(base_size = 7)+xlab("parameter")+
  ylab("probability")+theme(plot.title.position = "plot")+ggtitle("Prior")+
  scale_x_continuous()+scale_y_continuous(limits = c(0,upperLimit))


InertiaOneGPlot <- ggplot()+geom_line(aes(x = ps, y = posterior1g))+theme_tufte(base_size = 7)+xlab("parameter")+
  ylab("probability")+theme(plot.title.position = "plot",
                            axis.text.y = element_blank(),
                            axis.title.y = element_blank(),
                          axis.ticks.y =element_blank()
)+ggtitle("Evidence: h")+
  scale_y_continuous(limits = c(0,upperLimit))



InertiaTwoGPlot <- ggplot()+geom_line(aes(x = ps, y = posterior2g))+theme_tufte(base_size = 7)+xlab("p")+
  ylab("probability")+theme(plot.title.position = "plot"#, 
#                        axis.text.y = element_blank(),
#                            axis.title.y = element_blank(),
#                            axis.ticks.y = element_blank()
                  )+ggtitle("Evidence: h, h")+
scale_y_continuous(limits = c(0,upperLimit))

InertiaTwoGoneBluePlot <- ggplot()+geom_line(aes(x = ps, y = posterior2g1b))+theme_tufte(base_size = 7)+xlab("parameter")+
  ylab("probability")+theme(plot.title.position = "plot",     axis.text.y = element_blank(),
  axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
           )+ggtitle("Evidence: h, h, t")+
  scale_y_continuous(limits = c(0,upperLimit))
```



```{r Figinertia3, echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "80%",   message = FALSE, warning = FALSE, results = FALSE}
#| label: fig-intertia2
#| fig-cap: "As observations of heads, heads and tails come in, extreme parameter values drop out of the picture and the posterior is shaped by the evidence."


grid.arrange(InertiaPriorPlot, InertiaOneGPlot,  InertiaTwoGPlot, InertiaTwoGoneBluePlot, ncol = 2, nrow = 2)
```


The impossibility of defining proper scoring rules was another weakness of imprecise probabilism.  This is a significant shortcoming, especially because proper scores do exist for precise probabilism. Fortunately, one can show that there exist proper scoring rules for higher-order probabilism. <!--These rules can then be used to formulate accuracy-based arguments. In addition, recall the point made by  @Schoenfield2017accuracy:  an accuracy measure  will not usually recommend an imprecise stance. This argument 
fails against imprecise probabilism: there are cases in which accuracy considerations recommend an imprecise stance (that is, a multi-modal distribution) over a precise one. --> We defend this claim in the next section. The argument, however, will be more formal and  can be skipped upon first reading.


# Proper scores
\label{sec:proper-scores}

<!--
The overall dialecticts of this section is as follows. (1) we define two imporant divergence measures, (2) we define two inaccuracy measures using those divergencies, (3) we define the notion of expected inaccuracy, emphasizing a particular case, in which one additionally assumes that there are only two possible outcomes, 0 and 1. Then we (4) set up a thought experiment in which intuitively from three candidate subjective distributions only one matches the known true generative process and so should come out, intuitively, as the most accurate. Moreover, if a given accuracy measure is *proper* each of the subjective candidate distributions should expect itself to be the most accurate (in a sense to be specified). (5) We will show that propriety failures arise if we stick to the additional assumption mentioned in (3) and use our example to gesture towards a more general claim that's proven in the appendix, namely that KL divergence without this restrictive assumption is a strictly proper accuracy measure. (6) The example also illustrates that from the two measures discussed, the KL-divergence based one provides more intuitive results. 
-->


A scoring rule or inaccuracy score quantifies the distance (inaccuracy) between a probability distribution and the true state of the world.  A desirable property that any such score should have is strict propriety. In the precise case, let $I(p, w)$ be an inaccuracy score of a probability distribution $p$ relative to the true state $w \in W$. The score $I(p, w)$ is strictly proper if, for any other probability distribution $q$ different from $p$, the following holds: 

$$\sum_{w \in W} p(w) I(p, w) < \sum_{w \in W}  q(w)I(p, w).$$

<!--
A scoring rule or inaccuracy score quantifies the distance (inaccuracy) between a probability distribution and the true state of the world.  A desirable property that any such score should have is strict propriety. In the precise case, let $I(p, \theta_k)$ be an inaccuracy score of a probability distribution $p$ relative to the true chance $\theta_k$ defined over a parameter space $[0,1]$. The score $I(p, \theta_k)$ is strictly proper if, for any other probability distribution $q$ different from $p$, the following holds: 


$$\sum_{i=1}^{n} p_i I(p, Ind^k_i)< sum_{i=1}^{n}  q(w)I(p, Ind^k_i).$$


where  $Ind^k_i$ denotes $Ind^k(\theta_i)$ defined as:

\begin{align*}
Ind^k(\theta_i) & = \begin{cases} 1 & \mbox{if } \theta_i = \theta_k\\
                        0 & \mbox{otherwise}  \end{cases}
\end{align*}
-->


\noindent 
That is, the expected inaccuracy of $p$ from the perspective of $p$ itself should always be smaller than the expected inaccuracy of $p$ from the perspective of another distribution $q$. Other common requirements are: the score $I(p, w)$ should be a function of the probability distribution $p$ and the true state $w$ (extensionality); and the score should be a continuous function around $p$ (continuity). Well-known results demonstrate that the Brier score is an extensional score for precise probabilities that is both proper and continous.^[The Brier score is defined as the squared distance between the true state and the probability forecast, or formally,  $(p(x)-V(x, w))^2$, where $p(x)$ is the probability forecast and $V(x, w)$ determines if a proposition obtains at $w$ ($V(x, w)=1$) or not ($V(x, w)=0$). If, for example, the proposition 'rain' obtains at $w$, the forecast 'rain with .9 probability' would be more accurate at $w$ than the forecast 'rain with .8 probability'. If, on the other hand, the proposition 'not rain' obtained at $w$, the latter forecast would be more accurate.]\todo{Is the formal statement of Brier score in the footnote correct?} \todo{Looks alright, don't remove comment, will pay attention when reading the section at some point.}

<!--
The propriety requirement is that a measure expects itself to be the least accurate---to be able to formalize this requirement, we need the notion of *expected inaccuracy*. To calculate the expected accuracy of a measure, first you need to calculate its inaccuracy w,r,t  every possible hypothesis. Then, you factor in how probable the measure thinks those hypothesis as weights to calculate the expectation. Within our working framework in which we work with a discretized vector $x$ of possible outcomes, outcomes the expected inaccuracy is:
$$\mathbb{E_{discretized}}I(p,q) = \sum  I(p,w) q(w),$$
\noindent where $I(p,w)$ is an inaccuracy score of  distribution $p$ relative to the possible world $w$ (we use $I_{KL}$, but in principle other distance measures can be plugged in here).^[In practice, instead of working with $w$ and saying things like $V(w)=x$, we will be working with scalars and finite vectors, and simply say that the outcome is $x$.]
-->


Can similar results be established for imprecise probabilities? The answer is likely to be negative since several hurdles exist. 
First, a plausible scoring rule for imprecise probabilities cannot be easily found. Suppose an imprecise forecast assigns the [.8, .9] interval to the outcome that it would rain tomorrow, where the true state is 'rain'. Would the [.6, .99] interval be more accurate since its .99 upper bound is closer to the true state? Intuitively, the [.6, .99] interval should not be more accurate, otherwise, the trivial interval [0, 1] would always be more accurate than any other interval. To remedy this problem, a plausible inaccuracy score for imprecise probabilities could be directly proportional to the Brier score computed using the side of the interval closer to the true state, but inversely proportional to the size of the interval.^[If the score increases when the Brier score (thus defined) increases or when the size of the interval increases, this would block the result that the [0, 1] interval is always the most accurate; see @seidenfeld2012forecasting.] 

Still, 
even if a well-behaved score for imprecise probabilities can be found, a second hurdle remains. Recall that the notion of expected inaccuracy is needed to establish strict propriety. Unfortunately, expected inaccuracy cannot be easily defined for imprecise probabilities. Let $I([p_-, p_+], w)$ be an inaccuracy score for the interval $[p_-, p_+]$. What is its expected inaccuracy from the perspective of the interval $[p_-, p_+]$ itself, or from the perspective of another interval $[q_-, q_+]$? There is no standard answer to this question. After all, $I([p_-, p_+], w)$ cannot be multiplied by $[p_-, p_+]$, in the way in which $I(p, w)$ can be multipled by $p(w)$. And if the notion of expected inaccuracy is not defined, the question of whether an imprecise inaccuracy score can be proper is ill-defined.  Perhaps the expected inaccuracy of the interval $[p_-, p_+]$ can be evaluated from the perspective of the precise probabilities $p_-$ or $p_+$ at the edges.^[So the expected inaccuracy of $I_w([p_-, p_+])$ would equal $\sum_{w\in W} p_-(w)I([p_-, p_+], w)$ or $\sum_{w\in W} p_+(w)I([p_-, p_+], w)$.] But now the impreciser would face another problem. If the expected inaccuracy of the interval $[p_-, p_+]$ is evaluated from the perspective of the precise probabilities at the edges, finding a proper inaccuracy score that is also continous turns out to be impossible  [@seidenfeld2012forecasting].  


Despite the difficulties that plague imprecise probabilism in defining scoring rules that are proper, here we put forward an intuitively plausible scoring rule for higher-order probabilities that is both proper and continuous. Building on existing work on this topic (@HersbachDecomp2000, @Pettigrew2012Epistemic-Utili, @GneitRafter2007), the higher-order scoring rule we propose is based on a well-known measure of divergence between probability distributions, the Kullback-Leibler (KL) divergence, which is defined as follows: <!--A key modification to this framework involves abandoning the assumption that only two possible outcomes (truth and false) exist. Once we present the positive proposal, we will illustrate how such an assumption is problematic.--> 
$$ D_{\text{KL}}(q \,||\, p) = \sum_{x} q(x) \log\left(\frac{q(x)}{p(x)}\right) $$

\noindent This is a standard information-theoretic measure of divergence of $p$ from $q$ from the perspective of $q$. For computational ease, we are using a grid approximation instead of continuous distributions, as in practice we are unable to work with infinite precision.\footnote{In the continuous case, we would need to use the so-called differential KL divergence.}   To this end, $x$ will denote the finite vector of discrete outcomes under consideration. <!--- (by default, we will use equally spaced values between 0 and 1 here). ---> 

The goal is to deploy KL divergence as a measure of inaccuracy of $p$ relative to a true state $w\in W$, denoted by $I_{KL}(p, w)$. To this end, let $t_w(x)$ be the omniscient distribution tracking the true state $w$. For any outcome $x$, the distribution $t_w(x)$ will either assign probability 1 (if the outcome obtains in $w$) or 0 (if the outcome does not obtain in $w$). Since $t_w(x)$ will equal one for the true outcome $x$, call it $x_w$, and zero for the others, KL divergence simplifies to:
\todo{!please check that notation is correct throughout! Are the x's a partition?}

$$ I_{KL}(p, w) = D_{\text{KL}}(t_w \,||\, p) = \sum_{x} t_w(x) \log\left(\frac{t_w(x)}{p(x)}\right) = \log\left(\frac{1}{p(x_w)}\right)= -\log p(x_w) $$

\noindent That is, $I_{KL}(p, w)$ is the KL divergence of $p$ from the omniscient probability distribution $t_w$.







<!--- which assigns probability 1 to the outcome $V(w)$ and probability 0 to any other outcome,^[If you prefer to think in continuous terms, you'd need to use Delta-Dirac distributions here.] as seen from the perspective of $w$, the "ground truth" here, conceptually.--->
 
 <!-- Let the outcome under consideration in $w$  be $V(w)$ (in a very simple case you might take there to be two possible oucomes, 0 and 1; the range of $V$ is simply the vector of possible states). So the  KL-inaccuracy of a distribution $p$ with respect to $w$ is defined, as follows:
-->

<!--
 \begin{align*}
I_{KL}(p, w) & = D_{\text{KL}}(f(w)||p)
\end{align*}

\noindent where:
\begin{align*}
f(x) & = \begin{cases} 1 & \text{ if } x = V(w)\\
0 & \text{ o$\,$/w. }
\end{cases}
\end{align*}
-->

The KL inaccuracy score $I_{KL}$ applies to any probability distribution $p$, including higher-order ones. For suppose the possible outcomes are the chance hypotheses $\theta_1, \theta_2, \dots, \theta_n$ about the true bias of a coin. Let $\theta_w$ be the true bias of the coin at $w$, and let $p(\theta_w)$ be the higher-order probability that the distribution $p$ assigns to the bias $\theta_w$. Then, $-\log p(\theta_w)$ is the KL-based inaccuracy score of the higher-order probability $p$ at $w$. If, for example, the true bias of the coin is $.6$ and the higher-order distribution $p$ assigns $.8$ to this bias, the higher-order inaccuracy score of $p$ would be $-\log .8$. Notice that, on this approach, two distributions $p$ and $q$ which assign the same probability to the true chance hypothesis in $w$---$p(\theta_w)=q(\theta_w)$---will have the same inaccuracy score $I_{KL}$ even though they might differ in the probabilities they assign to other chance hypotheses. So the shape of the distribution does not matter for the inaccuracy score; it does matter for expected inaccuracy, as we will soon see.

We will now establish the strict propriety of the scoring rule $I_{KL}$. The first step is to define the score's expected inaccuracy, as follows: 
\todo{summing over w or $x_w$? confused about that? check!}
$$\sum_{w\in W} q(x_w)I_{KL}(p, w) = \sum_{w\in W} q(x_w)(-\log p(x_w)) $$

\noindent
In other words, consider several potential true outcomes $x_w$, each associated with a true state $w$ (say, $n$ true chance hypotheses $\theta_w$); then, compute the inaccuracy scores of $p$ with respect to each of the $x_w$'s (omniscient distributions, chance hypotheses), that is, $-\log p(x_w)$; finally, calculate the expected inaccuracy by summing over the entire distribution $q$. 
Now, to show strict propriety, it is enough to notice two facts. First, the expected inaccurracy of $p$ from the perspective of $p$ itself is the entropy of $p$, 
namely $H(p)$.^[As usual, the expected inaccuracy of $p$ from the perspective of $p$ itself is obtained by replacing $q(x_w)$ with $p(x_w)$ in the formula above: $\sum_{w\in W} p(x_w)I_{KL}(p, w) = -\sum_{w\in W} p(x_w)\log p(x_w) = H(p).$] Second, the expected inaccuracy of $p$ from the perspective of another distribution $q$ is the cross-entropy between $p$ and $q$, namely $H(p, q)$.^[$\sum_{w\in W} q(x_w)I_{KL}(p, w) = H(p, q).$] Since $H(p)$ is always smaller than $H(p, q)$ (see the appendix for details), the inaccuracy score $I_{KL}$ is proper.\todo{Say that CRPS  is used for example by Konek}\footnote{KL divergence is not the only possible score for higher-order probabilities. Another approach relies on the Cramer-Von-Mises measure (CRPS). In the discretized version, it is defined as follows:
\begin{align*}
D_{\text{CM}}(p,q) & = \sum_{x} \vert P(x) - Q(x)\vert^2, 
\end{align*}

\noindent where $P$ and $Q$ are the cumulative distributions corresponding to the probability distribution $p$ and $q$. Looking at cumulative densities ensures that all densities are considered on the same scale. (In the continuous case, this measure is defined as the area under the squared Euclidean distances between the corresponding cumulative density functions. That is, $D_{CM}(p,q)  = \int_{0}^{1} \vert P(x) - Q(x)\vert^2 \, dx$.) If $q$ plays the role of the true distribution, the CRPS measure can be turned into an inaccuracy measure that closely resembles the Brier score. The inaccuracy of $p$ with respect to the true state $w$ is defined, as follows:
\begin{align*}
I_{\text{CRPS}}(p,w) &= \sum \vert \mathsf{P}(x) - \mathbf{ 1 }(x\geq V(w))\vert ^2 
\end{align*}
\noindent where:
\begin{align*}
\mathbf{ 1 }(x \geq V(w)) & = \begin{cases} 1 & \text{ if } x \geq V(w)\\
0 & \text{ o$\,$/w. }
\end{cases}
\end{align*}
\noindent 
Finally, Cramer-Von-Mises can also be used to define the expected inaccuracy, analogously to the case of KL-divergence.
However, for reasons that will soon become clear, we believe that KL divergence is a more natural higher-order inaccuracy measure.} 
\todo{why two acronyms, CM adn CRPS? CM is enough, no?}

<!--
^[Note that there are no readily computable solutions to the integral used in the definition of CRPS, although it can sometimes be evaluated in closed form [ @GneitRafter2007, p. 366].]}
-->

<!---
The crucial characteristic of a scoring rule that we go after is strict-propriety. Our  discussion in this section will be based on an example for the sake of accessibility. We do, however, include the proof of strict propriety of the as an KLD-based inaccuracy measure in the appendix. In a nutshell, the argument demonstrates that for a second-order discretized probability mass $p$ over a parameter space $[0,1]$, with the actual probability denoted as $\theta$,  the Kullback-Leibler divergence of $p$ from the indicator distribution  of $\theta$  (which assigns 1 to $\theta$ and 0 to all other parameter values in the parameter space) is expressed as $\mathcal{I}_{\dkl}^2$.
-->

<!--
^[The argument generalizes 
to parameter spaces that correspond to probabilities of multiple propositions which are Cartesian products of parameter spaces explicitly used in the argument in this section.] This serves as a demonstration of the strict propriety of the inaccuracy measure: each $p$ anticipates itself to be the least inaccurate distribution.\footnote{ The argument has four key moves. Consider a given "point probability", i.e. the unique $x$ such that $V(w)=x$. Call this parameter $\theta$.
-->

<!--   
\begin{enumerate}
\item we show that the inaccuracy of $p$ w.r.t.  $\theta$ can be reduced to $- \log_2 p(\theta)$,
\item  then we show that the expected inaccuracy of $p$ from the perspective of $p$ is the entropy of $p$, $H(p)$,
\item  next we show that the inaccuracy of $q$ from the perspective of $p$ is the cross-entropy $H(p,q)$,
\item we then plug in  the well-established result that cross-entropy is strictly larger than entropy as soon as $p\neq q$.
  \end{enumerate}
}
-->


<!--- While imprecisers face challenges in formulating appropriate scoring rules, the second-order approach, which incorporates KLD and centers its focus on the distribution of all conceivable hypotheses, emerges as a more coherent framework. 
--->

To see that the proposed account of expected inaccuracy works as intended, consider a variation of a scenario by @Schoenfield2017accuracy. A rational agent is invited to engage in a bet by an opponent who has a representative bag of coins coming from a factory where the distribution of bias among the coins produced, the true generative process, is known. It is a mixture of two normal distributions centered at $.3$ and at $.5$, both with standard deviation of $.05$. The opponent randomly selects one of the coins in the bag and flips it. The rational agent who knows this set-up may form a number of higher-order credal states in response to this information. Consider three such credal states, out of many options: first, a faithful bimodal distribution centered at $.3$ and $.5$; second, a unimodal distribution centered at $.4$; third, a wide bimodal distribution centered at $.2$ and $.6$. The three options are depicted in @fig-emc. <!--All of them have expected values at about $.4$. So, if precise probabilities were the only measure of uncertainty, $.4$ would be the most natural value to assign to the probability that coin came up, say, heads. The three distributions, however, differ in how they represent higher-order uncertainty, and it seems that the faithful bimodal distribution gives the best representation, a point to which we will return. -->


```{r calculationsEMC,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}

plotDistroPlain <- function(distro, title, mult = 1.2) {
  plot <-  ggplot()+theme_tufte(base_size = 7)+xlab("parameter values")+
    ylab("probability")+theme(plot.title.position = "plot")+ylim(0,1)+
    ggtitle(title)+
    geom_line(aes(x = ps,y = distro))+
    ylim(c(0,mult * max(distro)))
  return(plot)
}


kld <- function(p,q) kullback_leibler_distance(p,q, testNA = TRUE, unit = "log2",
                                               epsilon = 0.00001)

n <- 1000
ps <- seq(0,1,  length.out =n)
ch1 <- c(rep(0,n-1),1)
indicator1 <- as.numeric(ps >= 1)
ch0 <- c(1, rep(0,n-1))
indicator0 <- as.numeric(ps>=0)
a <- dnorm(ps, .3, .05)
b <- dnorm(ps, .5, .05)
c <- ifelse(ps <= .4, a, b) 

bimodal <- c / sum(c)
bimodalCum <- cumsum(bimodal)
distBi1 <- (bimodalCum - indicator1)^2
distBi0 <- (bimodalCum - indicator0)^2

centered <-   dnorm(ps, .4, .05)
centered <- centered/sum(centered)
centeredCum <- cumsum(centered)
distCe1 <- (centeredCum - indicator1)^2
distCe0 <- (centeredCum - indicator0)^2

aw <- dnorm(ps, .2, .05)
bw <- dnorm(ps, .6, .05)
cw <- ifelse(ps <= .4, aw, bw) 
bimodalWide <- cw / sum(cw)
bimodalWideCum <- cumsum(bimodalWide)
distBiW1 <- (bimodalWideCum - indicator1)^2
distBiW0 <- (bimodalWideCum - indicator0)^2

#now CVM distances from truth and falsehood
#Bi wins in particular distances
dc1 <- sum(distCe1)
db1 <- sum(distBi1)
dbw1 <- sum(distBiW1)
dc0 <- sum(distCe0)
db0 <- sum(distBi0)
dbw0 <- sum(distBiW0)

#now expected values
expBi <- sum(ps * bimodal)
expCe <- sum(ps * centered)
expBiW <- sum(ps * bimodalWide)

expCVMbi <- expBi * db1 + (1-expBi) * db0
expCVMbW <- expBiW * dbw1 + (1-expBiW) * dbw0
expCVMCe <- expCe * dc1 + (1-expCe) * dc0


# now with KLD to omniscient function
kldBi1 <- kld(ch1,bimodal)
kldBi0 <- kld(ch0,bimodal)

kldBiW1 <- kld(ch1,bimodalWide)
kldBiW0 <- kld(ch0,bimodalWide)

kldCe1 <- kld(ch1,centered)
kldCe0 <- kld(ch0,centered)


expKLDbi <- expBi * kldBi1 + (1-expBi) * kldBi0
expKLDbW <- expBiW * kldBiW1 + (1-expBiW) * kldBiW0
expKLDCe <- expCe * kldCe1 + (1-expCe) * kldCe0
```



```{r figEMC,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "85%",   message = FALSE, warning = FALSE, results = FALSE}
#| label: fig-emc
#| fig-cap: 'Three distributions in a vague EMS scenario. The distributions are built from normal  distributions with standard deviation $.05$, the bimodal ones are joint in the middle. All of them have expected values $\approx .4$.'
#| fig-pos: H

grid.arrange(plotDistroPlain(bimodal, "Bimodal, with modes at .3 and .5"),
plotDistroPlain(centered, "Centered around .4"),
plotDistroPlain(bimodalWide, "Wide bimodal, with modes at .2 and .6"), ncol = 1)
```




```{r figinaccuracies2,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
#| label: fig-inaccuracies2
#| fig-cap: 'CRPS and KL divergence based inaccuracies relative to $n$ true chance hypotheses for the three distributions, faithful bimodal, centered unimodal and wide bimodal.'
#| fig-pos: H


point <- function (value) ifelse(abs(ps - value) == min(abs(ps - value)), 1, 0)
indicator <- function(pointf) ifelse( ps >= ps[min(which(pointf != 0))], 1,0  )

cvm <- function(w,p){
  cumw <-  cumsum(w)
  cump <- cumsum(p)
  dist <- (cump - cumw )^2
  return(sum(dist))
}

inaccuracyChancy <- function(distro){
crpss <- numeric(length(ps))
klds <- numeric(length(ps))
for(i in 1:length(ps)){
chance <- point(ps[i])
crpss[i] <- cvm(chance,distro)
klds[i] <- kld(chance,distro)
}
df <- data.frame(ps,distro,crpss,klds)
return(df)
}

InaBi <- inaccuracyChancy(bimodal)
InaCe <- inaccuracyChancy(centered)
InaBiW <- inaccuracyChancy(bimodalWide)



fun_perspective <- function(perspective, expInaccScores){ # vectors!
  return(sum(perspective * expInaccScores))
}

perspectives <- list(
  'centered' = centered,
  'bimodalWide' = bimodalWide,
  'bimodal' = bimodal
)

expInaccScores <- list(
  'InaBiCVM' = InaBi$crpss,
  'InaBiKLD' = InaBi$klds,
  'InaCeCVM' = InaCe$crpss,
  'InaCeKLD' = InaCe$klds,
  'InaBiWCVM' = InaBiW$crpss,
  'InaBiWKLD' = InaBiW$klds
)

combinations <- expand.grid(names(perspectives), names(expInaccScores))

expPersDF <- data.frame(Perspective = character(),
                     ExpInaccScore = character(),
                     Score = numeric(),
                     stringsAsFactors = FALSE)

for (i in 1:nrow(combinations)) {
  perspective_name <- combinations[i, 1]
  expInaccScore_name <- combinations[i, 2]
  perspective <- perspectives[[perspective_name]]
  expInaccScore <- expInaccScores[[expInaccScore_name]]
  score <- fun_perspective(perspective, expInaccScore)
  expPersDF <- rbind(expPersDF, data.frame(Perspective = perspective_name,
                                     ExpInaccScore = expInaccScore_name,
                                     Score = score))
}

expPersDF$Score <- round(expPersDF$Score, 3) 


bcplot <- plotDistroPlain(InaBi$crpss, "Bimodal: CRPS")+ylab("inaccuracy")+
            xlab("true probability")

bkplot <- plotDistroPlain(InaBi$klds, "Bimodal: KLD")+
  ylab("inaccuracy")+xlab("true probability")


ccplot <- plotDistroPlain(InaCe$crpss, "Centered: CRPS")+ylab("inaccuracy")+
  xlab("true probability")

ckplot <- plotDistroPlain(InaCe$klds, "Centered: KLD")+ylab("inaccuracy")+
  xlab("true probability")

bwcplot <- plotDistroPlain(InaBiW$crpss, "Bimodal wide: CRPS")+ylab("inaccuracy")+
  xlab("true probability")

bwkplot <- plotDistroPlain(InaBiW$klds, "Bimodal wide: KLD")+ylab("inaccuracy")+
  xlab("true probability")

library(gridExtra)
grid.arrange(bcplot, bkplot, ccplot, ckplot, bwcplot, bwkplot)
```






\begin{table}[H]
\begin{tabular}{lrrrrrr}
& \multicolumn{3}{c}{CRPS} & \multicolumn{3}{c}{KLD} \\
\toprule
  & bimodal & centered & wide bimodal & bimodal & centered & wide bimodal\\
\midrule
bimodal & 64.670 & 78.145 & 88.380 & 8.577 & 10.655 & 11.336\\
centered & 41.657 & 28.181 & 85.911 & 9.239 & 7.690 & 15.627\\
wide bimodal & 137.699 & 171.719 & 113.989 & 11.541 & 19.231 & 8.689\\
\bottomrule
\end{tabular}
\caption{Expected inaccuracies of the three distributions from their own perspective and that of the other distributions. Each row corresponds to a perspective.}
\label{tbl:expected2}
\end{table}

 
The accuracy scores of these higher-order distributions are in @fig-inaccuracies2.  Each point in the graph reflects the accuracy score calculated relative to a possible omniscient distribution corresponding to the values of $\theta$, the true bias of the coin. \todo{NL: Distribution, not a value representing a coin's bias?}The expected inaccuracies of the three distributions from their own perspective, as well as from the perspective of the other distributions, are in \mbox{Table \ref{tbl:expected2}.} The results are as intended: from their own perspective, the distributions see themselves as the least inaccurate. The strict propriety of the scoring rule is verified.^[One important difference transpires between using CRPS and KLD. Notice how for chance hypotheses between the actual peaks the inaccuracy remains flat. This seems to be an artifice of choosing a squared distance metric. If instead we go with a more principled, information-theory-inspired KL divergence, inaccuracy in fact jumps a bit for values in between the peaks for the bimodal distributions, which seems intuitive and desirable. This seems to be a reason to prefer a KL-based inaccuracy scores.] 


Even though they all recommend themselves, the three distributions are by no means equivalent. The faithful bimodal is the one that best reflects the true generative process, while the others less so. How does the KL-based inaccuracy score we propose to capture the fact that the wide bimodal seems more adequate than the others? This is apparent by looking at the inaccuracy score relative to two chance hypotheses $H_3$, where the true chance is $0.3$, and $H_5$, where the true chance is $0.5$. \todo{Why just those two, and not also .4, another possible true chance hypothesis? The unimodal has a better score of .4. This seems ad hoc. }You can find the inaccuracies for them in Table \ref{tbl:schoen}. To make sure that this favorable outcome isn't due to not using pointed credences, we can redo the calculations using the pointed version. In the pointed version, all the focus is on 0.4, or the weight is evenly divided between 0.3 and 0.5, or between 0.2 and 0.6. As anticipated, when we consider inaccuracy, both of these setups recommend the faithful bimodal distribution (Table \ref{tbl:schoen2}).


```{r tableh3h5,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}
h3 <- rep(0, n)
h3[300] <- 1

h5 <- rep(0, n)
h5[500] <- 1

BKLDH3 <- kld(bimodal, h3)
BKLDH5 <-kld(bimodal, h5)

CKLDH3 <- kld(centered, h3)
CKLDH5 <-kld(centered, h5)

BWKLDH3 <- kld(bimodalWide, h3)
BWKLDH5 <-kld(bimodalWide, h5)

BCVMH3 <-cvm(bimodal, h3)
BCVMH5 <-cvm(bimodal, h5)

CCVMH3 <-cvm(centered, h3)
CCVMH5 <-cvm(centered, h5)

BWCVMH3 <-cvm(bimodalWide, h3)
BWCVMH5 <-cvm(bimodalWide, h5)

```

\begin{table}[H]
\centering
\begin{tabular}{lrrrr}
 & \multicolumn{2}{c}{CRPS} & \multicolumn{2}{c}{KLD} \\
\toprule
&H3 & H5 & H3 & H5\\
\midrule
bimodal &55.475 & 55.378 & 7.935 & 7.935\\
centered &72.281 & 72.090 & 9.836 & 9.825\\
wide bimodal & 86.230 & 86.223 & 10.871 & 10.882\\
\bottomrule
\end{tabular}
\caption{CRPS and KLD inaccuracies of the three distributions with respect to the two hypotheses. On both inaccuracy measures the bimodal distribution dominates the other two.}
\label{tbl:schoen}
\end{table}

```{r tableh3h52,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE, results = FALSE}


point_bimodal <- rep(0, n)
point_bimodal[c(300, 500)] <- 1

pointKLDBH5 <- kld(point_bimodal, h5)   

pointCVMBH3 <- cvm(point_bimodal, h3)

```


\begin{table}[H]
\centering
\begin{tabular}{lrrrr}
 & \multicolumn{2}{c}{CRPS} & \multicolumn{2}{c}{KLD} \\
\toprule
 &H3 & H5 & H3 & H5\\ \midrule
pointed bimodal &49.75 & 49.75 & 1.00 & 1.00\\
pointed centered &100.00 & 100.00 & 16.61 & 16.61\\
pointed wide bimodal & 99.75 & 99.75 & 16.61 & 16.61\\
\bottomrule
\end{tabular}
\caption{CRPS and KLD inaccuracies of the three-pointed distributions with respect to the two hypotheses.}
\label{tbl:schoen2}
\end{table}

 \todo{Not sure this last part is needed.}
Key to the result of strict propriety is summing over all possible chance hypotheses. But the precise probabilist might insist that this is unnecessary. There are ultimately only two possible first-order level outcomes, heads and tails, and thus also only two possible chance hypotheses (or omniscient distributions), one that places all weight on heads and the other that places all weight on tails.  On this view, the accuracy score $I_{KL}$ of a distribution $p$, should only take two possible values, $I(p,\mathsf{heads})$ and $I(p,\mathsf{tails})$, where  '$\mathsf{tails}$' or '$\mathsf{heads}$' is one of the two onmiscient distributions. Thus, expected inaccuracy would be calculated as follows:

$$\mathbb{E_{binary}}(p,q) = I(p,\mathsf{heads}) \mathbb{E}q(\mathsf{heads}) + I(p,\mathsf{tails}) \mathbb{E}q(\mathsf{tails}).$$

\noindent
Instead of calculating inaccuracy relative to all possible chance hypotheses about the coin's bias, expected inaccuracy would result from the sum of the two inaccuracy scores weighted by the probabilities of the two outcomes. To be sure, the three distributions considered so far do not provide the probabilities of the outcome 'heads' or 'tails' directly. They assign probabilities to different values of coin bias, but it is natural enough to take the expected values of these distributions, which equals .4 for all three. Table \ref{tbl:comp1} displays the relevant inaccuracy scores as well as the expected inaccuracy scores. Note that since the probability of heads or tails is the same on all the distributions, those are expected values from the perspective of each of the measures; changing the perspective in this example doesn't change the expected inaccuracy.

<!--^[The expected values of the form $\mathbb{E}_{\mathsf{distribution}}(H) = \sum (x \times \mathsf{distribution}(x))$, where $x$ are the values on the discretized grid  (as in our example these expectations are pretty much the same, we can simply take $\mathsf{distribution}(H)$ to be .4).]
 -->

<!--Now consider what happens if we think of expected inaccuracy of these distributions assuming there are only two possible true outcomes, conceptualized as either heads ($H$) or tails ($T$). Whatever our inaccuracy measure, we will have six inaccuracy scores of the form $I(\mathsf{distribution}, \mathsf{outcome})$, where $\mathsf{outcome}$ is one of two omniscient distributions that give all weight to either heads or tails. 
-->

\noindent  
 
 This approach to expected inaccuracy runs into trouble. As it turns out, the expected KL-inaccuracy score recommends the wide bimodal distribution as the most accurate (or least inaccurate), and the KL divergence from the omniscient measure makes the same recommendation. This is counterintuitive because the faithful bimodal seems the most evidence-responsive. The unimodal distribution, while centering on the expected value, gets the chances wrong, and the wide bimodal has its guesses too close to the truth values and too far from the known chances.  But there is a further problem.  While the wide bimodal distribution expects itself to be the least inaccurate, the other distributions also expect the wide bimodal to be the least inaccurate. In this setting, then,  strict propriety fails.
 since some distributions recommend others as less inaccurate, whatever the true state of the world.

 <!---

```{r tabelPointsCRPSKLD,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%", message = FALSE, warning = FALSE}
#| label: tbl-comp1
#| tbl-cap: 'CPRS and KLD inaccuracies of the three distributions to the TRUE and FALSE omniscient functions, with expected inaccuracies, calculated using the shared point estimates of the probabilities of heads and tails.'

library(knitr)
# library(kable)
library(kableExtra)
table_pointsCRPSKLD <- data.frame(
  distribution = c("bimodal", "centered", "wide bimodal"),
  CRPS1 = c(db1, dc1, dbw1),
  CRPS0 = c(db0, dc0, dbw0),
  KLD1 = c(kldBi1, kldCe1, kldBiW1),
  KLD0 = c(kldBi0, kldCe0, kldBiW0),
  ExpCRPS = c(expCVMbi, expCVMCe, expCVMbW),
  ExpKLD = c(expKLDbi, expKLDCe, expKLDbW)
)

table_pointsCRPSKLD %>%
  kable("latex", booktabs = TRUE)


```

--->

\begin{table}[H]
\centering
\begin{tabular}{lrrrrrr}
\toprule
distribution & CRPS1 & CRPS0 & KLD1 & KLD0 & ExpCRPS & ExpKLD\\
\midrule
bimodal & 534.7305 & 334.9305 & 80.06971 & 33.90347 & 414.8505 & 52.36997\\
centered & 571.2192 & 371.4192 & 110.84220 & 53.13440 & 451.3392 & 76.21752\\
wide bimodal & 485.4052 & 285.6177 & 54.13433 & 19.50965 & 365.5340 & 33.35974\\
\bottomrule
\end{tabular}
\caption{CPRS and KLD inaccuracies of the three distributions to the TRUE and FALSE omniscient functions, with expected inaccuracies, calculated using the shared point estimates of the probabilities of heads and tails.}
\label{tbl:comp1}
\end{table}


 


What are we to make of this result? The same expected value $.4$ is used in the calculations of the expected inaccuracies on the assumption that there are two possible outcomes with respect to which expected inaccuracy is calculated. This approach, however, runs against the spirit of our enterprise. If expected values are often not good representations of a rational agent's uncertainty, it should not be surprising that relying on them fails to deliver plausible expected accuracy scores. By reducing each of the distributions' stances towards heads to a single point value .4, key information is washed away. As emphasized earlier, 
rather than measuring inaccuracy relative to two omniscient distributions that peak at either 0 or 1 and averaging using expected values of the distributions, we should instead utilize a set of $n$ potential true probability hypotheses.  We then compute all the inaccuracies with respect to each of these $n$ values represented by possible omniscient distributions (or true chance hypotheses) and determine the expected inaccuracy scores using the entire distributions rather than relying solely on the expected values of the distributions. As we have seen, this approach delivers the result of strict propriety we were looking for.\todo{take a look at fn}
^[What if we abandon this focus on true distributions being "pointy" and allow them to be any distributions whatsoever? Then we have a bit of work to do to make sure the framework is coherent and the semantics is well-defined (this is possible, as illustrated by the work of CITE DORST). But however we do that, if we then take KL divergence to the true distribution as an inaccuracy measure, since $\log(1)$ is $0$, whatever the weights,
any distribution will expect itself to be the least inaccurate, as expected inaccuracy calculation for a distribution and itself will always involve multiplying by zero.]



# Conjunctions
\label{sec:higher-order-conjunction}

Let us take stock.  In some circumstances, assigning sharp probabilities to events is justified. In others, it is less so, for example, when the bias of a coin is unknown, or when there is evidence that a coin could have a number of biases. In such cases, imprecise probabilities model uncertainty better than precise ones. But imprecise probabilities also fall short in their own way, for example, when the biases of a coin are not equally likely given the evidence available. Higher-order probabilities are better able to model these more complex scenarios. They also avoid many of the problems of imprecise probabilities, such as belief inertia and the difficulty of finding proper scoring rules.

One limitation of the discussion so far, however, is that we only looked at assessing probabilities of individual events, say whether a coin would come up heads or tails. But, of course, rational agents may need to assign probabilities to multiple events, for example, the conjunction of two events. Suppose I am holding two coins, and I have information about their respective biases. What is, then, the probability that they both come up, say, heads? In the precise case, the answer is straightforward: assuming independence, it is enough to multiply the individual probabilities. But what happens in the imprecise case? And how to proceed with higher-order probabilities? Once again, we will see that in assessing probabilities for conjunctions of events higher-order probabilities fare better than precise and imprecise ones.


```{r introStarts,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE}
ps <- seq(0,1,length.out = 1001)

hairMean <-  29/1148
hairA <- 30
hairB <- 1149

dogMean <- 2/78
dogA <- 3
dogB <- 79


lik0 <- hairMean * dogMean
prior <- seq(0,.3, by = 0.001)
priorH0 <- 1-prior
denomin <- lik0 * priorH0 + prior
num <-  lik0 * priorH0
posterior <- 1- num/denomin
threshold <- min(prior[posterior > .99])

pointImpactPlot <- ggplot()+geom_line(aes(x = prior, y = posterior))+xlim(0,.07)+
  theme_tufte(base_size = 10)+labs(title = "Prior vs. posterior, based on point estimates",
                                   subtitle = "Joint evidence: dog & hair")+
  geom_vline(xintercept = threshold, lty = 2, size =.5, alpha = .7)+
  annotate(geom = "label", label = "posterior > .99", x = .067, y =.95, size = 2.5)+
  theme(plot.title.position = "plot")
```



To fix ideas, we will go through a stylized legal example. We selected this example also to illustrate that higher-order probabilities can be useful beyond cases of coin tossing, though anything we say here does apply to coin tossing. Here is some preliminary background. In a murder case, the police recover two items of so-called match evidence: first, hair found at the crime scene matches the defendant's hair; and second, the fur of the defendant's dog matches the fur found in a carpet wrapped around one of the bodies.^[The hair evidence and the dog fur evidence are stylized after two  items of  evidence in the notorious 1981 Wayne Williams case [@deadman1984fiber1; @deadman1984fiber2].]  Clearly, these two matches constitute evidence against the defendant. The most obvious explanation is that the defendant visited the crime scene and contributed both traces. The alternative explanation is that the matches are a coincidence. Maybe another person visited the scene and happened to have the same hair type and a dog with the same fur type. How likely would that be? If the probability of this happening is low, the two matches would be strong incriminating evidence; if it is not low, they would be weak incriminating evidence. Trial experts usually provide coincidental match probabilities (also called random match probabilities). They express the likelihood that, by coincidence, a random person (or a random dog) who is not a contributor would still match. These are the probabilities we are looking for.  

It is customary to rely on database frequencies to assess the coincidental match probabilities, for example, by counting how many matches are found in a sample of the human population or the canine population. Suppose the matching hair type occurs  `r round(hairMean,4)` times in a reference database, and the matching dog fur type occurs  `r round(dogMean,4)` times in a reference database (more on how these numbers are calculated soon). These frequencies give the individual coincidental probabilities. To assess the probability of the two coincidental matches happening jointly, it is enough to multiply the individual probabilities: 
\begin{align*}
\pr{\s{dogMatch}  \vert \neg \s{contributor}} \times
\pr{\s{hairMatch} \vert \neg \s{contributor}}
& =  `r round(hairMean,4)` \times  `r round(dogMean,4)` = {`r round(hairMean * dogMean,6)`}
\end{align*}

\noindent
 Multiplication is allowed on the assumption that the coincidental matches are independent events.^[To put it more carefully, the two matches are independent conditional on the hypothesis that the defendant is not a contributor.] The resulting joint probability is very small. The two matches, combined, are strong evidence against the defendant, or so it would appear.

This is the story told by the precise probabilist.  But this story misses something crucial. As it happens, the coincidental match probability for hair evidence is based on $29$ matches found in a sample database of size $1,148$, while the coincidental match probability for the dog evidence is based on finding $2$ matches in a smaller database of size $78$. The relative frequencies are about $.025$ in both cases, but the two samples differ in size. The smaller the sample, the greater the uncertainty about the probabilities. So, for individual pieces of evidence, simply reporting the exact numbers makes it seem as though the evidential value of the hair and fur matches is the same, but actually, it is not.^[The probabilities in the Wayne Williams case on which our running example is based were $1/100$ for the dog fur, and $29/1148$ for Wayne Williams' hair.  Probabilities have been slightly but not unrealistically modified to be closer to each other in order to make a conceptual point: the same first-order probabilities, even when they sound precise, may come with different degrees of second-order uncertainty.] In the aggregate, multiplying the coincidental match probabilities further washes away this difference. 

A better alternative is easily available: take into account higher-order uncertainty. @fig-densities (upper part) depicts higher-order probability distributions of different coincidental match probabilities given the sample data---the actual number of matches found in the sample databases. <!---By hypothesis, 29 matches were found in the sample database of human hair of size 1,148, and 2 matches were found in the sample database of dog fur of size 78.---> As expected, some coincidental probabilities are more likely than others, and since the sizes of the two databases are different, the distributions have different spreads: the smaller the database the greater the spread, the greater the uncertainty about the coincidental probability. In light of this, @fig-densities (lower part) depicts the probability distribution for the joint coincidental match probabilities associated with both hair and fur evidence. The mathematics here is straightforward: once the higher-order distributions are known, simply multiply them to obtain the higher-order distribution of the joint coincidental match probabilities. 

```{r hair,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
#carpetSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, carpetA, carpetB))
set.seed(231)
hairSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, hairA, hairB))
dogSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, dogA, dogB))

#carpetHPDI <- HPDI(carpetSamples, prob  =.9)
hairHPDI <- HPDI(hairSamples, prob  =.99)
dogHPDI <- HPDI(dogSamples, prob  =.99)
```

```{r charitableImpact,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
lik0l <- .037 * .103
denominL <- lik0l * priorH0 + prior
numL <-  lik0l * priorH0
posteriorL <- 1- numL/denominL
thresholdL <- min(prior[posteriorL > .99])

charitableImpactPlot <- ggplot()+geom_line(aes(x = prior, y = posteriorL))+xlim(0,.32)+
  theme_tufte(base_size = 10)+labs(title = "Prior vs. posterior, charitable reading",
                                   subtitle = "Joint evidence: dog & hair")+
  geom_vline(xintercept = thresholdL, lty = 2, size =.5, alpha = .7)+
  annotate(geom = "label", label = "posterior > .99", x = .305, y =.95, size = 2.5)+ylab(
    "posterior"
  )+
  theme(plot.title.position = "plot")
```




```{r densitiesEvidence,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%"}

jointEvidence <- dogSamples * hairSamples


densities1Plot <- ggplot()+
  geom_line(aes(x = ps, y = dbeta(ps, hairA, hairB)), lty  = 2)+
  geom_line(aes(x = ps, y = dbeta(ps, dogA, dogB)), lty = 3)+xlim(0,.15)+
  xlab("probability")+
  ylab("density")+
  theme_tufte(base_size = 10)+
  #  annotate(geom  = "label", label = "carpet", x =  0.045, y = 140)+
  annotate(geom  = "label", label = "hair", x =  0.035, y = 80)+
  annotate(geom  = "label", label = "dog", x =  0.06, y = 15)+
#  labs(title = "Conditional densities for  individual items of evidence if the source hypothesis is false")+
#  theme(plot.title.position = "plot")
labs(title = "Distributions of individual coincidental probabilities")+
  theme(plot.title.position = "plot")


densities2Plot <- ggplot()+
  xlab("probability")+
  ylab("density")+
  theme_tufte(base_size = 10)+
  geom_density(aes(x= jointEvidence))+
  geom_vline(xintercept = 0.002760, lty = 2, linewidth = .5)+
  geom_vline(xintercept = 0.000023, lty = 2, linewidth  = .5)+
  geom_vline(xintercept = 0.000144, lty = 3, linewidth = .8)+
  geom_vline(xintercept = 0.001742, lty = 3, linewidth  = .8)+
  #labs(title = "Conditional density for joint evidence",
  labs(title = "Distribution for the joint coincidental probability",
       subtitle = "(with .99 and .9 HPDIs)")+
  theme(plot.title.position = "plot") +
  geom_vline(xintercept = hairMean * dogMean)
```


```{r Figdensities,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
#| label: fig-densities
#| fig-cap: "Beta densities for individual items of evidence and the resulting joint density with .99 and .9 highest posterior density intervals, assuming the sample sizes as discussed and independence, with uniform priors."
#| fig-pos: H

grid.arrange(densities1Plot,densities2Plot, ncol = 1 )
```


The precise probabilist might insist that our best assessment of the first-order coincidental match probabilities is still the relative frequency of matches found in the database, whether large or small. All things considered, our best assessment of the match probabilities for both fur and hair evidence should be about $.025$, based on the relative frequencies 2/78 and 29/1,148.   After all, if we were to bet whether a dog or a human picked at random would have the matching fur or hair type, our odds should be $.025$ no matter the size of the database. This argument has some bite for individual events. In fact, the expected values of the coincidental probabilities for hair and match evidence---based on the higher-order distributions in  @fig-densities (upper part)---still end up being about $.025$. If, as the precise probabilist assumes, first-order probabilities are all we should care about, going higher-order would seem a needless complication.

This line of reasoning, however, breaks down when evaluating conjunctions of events. What should our betting odds be for the proposition that a human and a dog, both picked at random, would have the matching fur and hair type in question? For the precise probabilist, the answer is straightforward: on the assumption of independence, multiply the $.025$ individual match probabilities and obtain a joint match probability of `r round(hairMean * dogMean,6)`. The higher-order probabilist will proceed differently. In assessing first-order match probabilities, they will retain information about higher-order uncertainty as much as possible. This can done in two steps: first, aggregate the higher-order distributions for the two match probabilities and obtain a higher-order distribution for the joint match probability (see @fig-densities); next, to obtain our best assessment of the first-order joint match probability, take the expected value of this latter distribution. Interestingly, the higher-order probabilist will assign `r round(mean(jointEvidence),6)` to the joint coincidental match probability, a value greater than what the precise probabilist would assign. 





```{r densitiesEvidenceModified,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%"}

# example 
# illustrates divergence between precise and higher order probabilist 
# on value of first-order joint match probabilities

ps <- seq(0,1, length.out = 1001)
set.seed(231)

# sample frequency, very small sample size

evidence1Mean <-  1/20
evidence1A <- 1
evidence1B <- 20

# sample frequency, big sample size

evidence2Mean <- 1000/20000
evidence2A <- 1000
evidence2B <- 20000

evidence1Samples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, evidence1A, evidence1B))

evidence2Samples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, evidence2A, evidence2B))

# how higher order probabilist assesses joint match probability
# small sample size frequencies for
# four cases: one, three, five, seven items of evidence (same weight each)

jointEvidence1 <- evidence1Samples * evidence1Samples
jointEvidenceThree1 <- evidence1Samples^3
jointEvidenceFive1 <- evidence1Samples^5
jointEvidenceSeven1 <- evidence1Samples^7 

# same as above but larger sample frequencies
jointEvidence2 <- evidence2Samples * evidence2Samples
jointEvidenceThree2 <- evidence2Samples^3
jointEvidenceFive2 <- evidence2Samples^5
jointEvidenceSeven2 <- evidence2Samples^7 

```


So, the higher-order and precise probabilist will disagree about the betting odds for the proposition that a human and a dog, both picked at random, would have the matching fur and hair type in question. The disagreement will become even starker as a larger number of independent items of evidence are evaluated.^[Consider the simple case of independent items of evidence whose individual match probabilities are $.025$. For three, five and seven items of evidence, the joint match probabilities would be: `r round((evidence1Mean)^3,6)`, `r round((evidence1Mean)^5,7)` and `r round((evidence1Mean)^7,10)` (for the precise probabilist); 
`r round(mean(jointEvidenceThree1),6)`, `r round(mean(jointEvidenceFive1),7)` and `r round(mean(jointEvidenceSeven1),10)` (for the higher-order probabilist, based on small databases of size 20); and 
`r round(mean(jointEvidenceThree2),6)`, `r round(mean(jointEvidenceFive2),7)` and `r round(mean(jointEvidenceSeven2),10)` (for the higher-order probabilist, based on larger databases of size 20,000).] Who should be trusted? Since the higher-order probabilist takes into accout more information---that is, the higher-distributions---there is good reason to think that the higher-order probabilist should be trusted more than the precise probabilist.

<!--^[As a further illustration of this point, consider a couple of variations of our running example. First, suppose the match probabilities associated with two matches are both set to $.05$, since they are based on the following relative frequencies: one match occurs in a dog fur database and one match occurs in a human hair database, where both databases are small, say of size 20. By multiplying the individual $1/.05$ likelihood ratios associated with the two matches, their evidential value against the defendant would seem quite strong: $1/.05\times 1/.05=`r 1/(evidence1Mean*evidence1Mean)`$. But the match probabilities are based on frequencies resulting from small databases, so their evidential value should be rather weak. Precise probability here seems to exaggerate the aggregate value of the evidence. Following higher-order probabilism, the  joint likelihood ratio would be `r 1/mean(jointEvidence1)`, a significantly smaller value. On the other hand, if the same $.05$ match probabilities were based on larger databases, the evidential value of the two matches should be correspondingly greater, but precise probabilism would make no difference. If, for example, 1,000 hair and fur matches are found in databases of size 20,000, the higher-order probabilist would assign `r 1/mean(jointEvidence2)` to the joint likelihood ratio, a much greater value than before. This outcome agrees with our intuituions.] 
-->

Consider now how the imprecise probabilism fares in these cases.
It will also run into its own problems when assessing probabilities of multiple events in conjunction. Recall that the probability measures in the representor set are those compatible with the evidence.  The problem is that almost any coincidental match probability will be compatible with any sample data^[Think by analogy to coin tossing: even a coin that has a .99 bias toward tails could come up heads on every toss. This series of outcomes is unlikely but possible. Similarly, a hair type that has a match probability extremely small could still be found several times in a sample population.] So the appropriate interval would be [0,1] for both coincidental match probabilities, and the same for the conjunction. This result would make it impossible to learn anything.

But now suppose we rely on reasonable ranges of coincidental match probabilities, for example, (.015,.037) (.002, .103), for hair and fur match evidence respectively.^[These are 99% credible intervals starting with uniform priors. A 99\% credible interval is the narrowest interval to which the   expert thinks the true parameter belongs with probability .99. For a discussion of what credible intervals are, how they differ from confidence intervals, and why confidence intervals should not be used, see @kruschke2015doing.] 
As expected, the range is wider for dog fur match evidence than hair match evidence: the uncertainty about the dog fur match probability is greater since the sample database was smaller. This is a good feature of the interval approach, unavailable to the precise probabilist. Now, what to do to assess the joint uncertainty? The most natural strategy is to focus on what happens at the edges of the two intervals. Reasoning with representor members at the edges of the intervals will yield the most extreme probability measure the impreciser is committed to, the worst-case and best-case scenarios. Following this strategy yields a new range for the joint match probabilties: (.00003, .003811).\footnote{Redoing the calculations using the upper bounds of the two intervals, $.037$ and $.103$,  yields the following:
\begin{align*}
\mathsf{P}(\s{dogMatch}\wedge \s{hairMatch} \vert \neg \s{contributor})   & =  .037 \times .103 =.003811.
\end{align*}

\noindent
 This number is around `r round(.003811/lik0,2)` times greater than the original, precise estimate. Given this number, the two matches are much weaker evidence for the contributor hypothesis than previously thought.    The calculation for the lower bounds, $.015$ and $.002$, yields the following:
\begin{align*}
\mathsf{P}(\s{dogMatch}\wedge \s{hairMatch} \vert \neg \s{contributor})   & =  .015 \times .002 =.00003
\end{align*}
   
 \noindent  
This number is around `r round(.0003/lik0,2)` times lower than the original estimate.  Given this number, the two matches are much stronger evidence for the contributor hypothesis than previously thought.} This interval does not seem to be particularly informative, but the bigger problem is that relying on ranges for the match probabilities leaves the impression that any value in the interval is just as good as any other. Perhaps we should pick the middle value as representative of the interval. To see why this will not work, consider again @fig-densities (lower part) which depicts the probability distribution for the joint match probability. This distribution is not symmetric. So the most likely value and the bulk of the distribution do not lie in the middle between the edges. <!--Just because the parameter lies in an interval with some posterior probability, it does not mean that the ranges near the edges of the interval are equally likely---the bulk of the density might very well be closer to one of the edges. --> So, only relying on the edges---or taking central values as representative---can lead to overestimating or underestimating the probabilities at play.\footnote{The calculations for the joint interval assume that because the worst- or best-case probability for one event is $x$ and the worst- or best-case probability for another independent event is $y$, the worst- or best-case probability for their conjunction is  $xy$. However, this conclusion does not follow if the margin of error (credible interval) is fixed. Just because the probability of an extreme value $x$ for one variable $X$ is .01, and so it is for the value $y$ of another independent variable $Y$, it does not follow that the probability that those two independent variables take values $x$ and $y$ simultaneously is the same. In general, it is impossible to calculate the credible interval for the joint distribution based solely on the individual credible intervals corresponding to the individual events.}

All in all, precise and imprecise probabilism does not fare well in assessing the probabilities of conjunction of independent events. In the case of individual events, this problem is not readily apparent, but when the probabilities of multiple events are assessed, the divergence between higher-order probabilism and the other versions of probabilism becomes significant. Insisting that all we should care about are first-order probabilities will not work if the values of the first-order probabilities are not assessed in light of all the information available. What precise and imprecise probabilities are ultimately guilty of is neglecting useful information.

<!--
the value of evidence in the aggregate. 
 Instead, the evaluation of multiple items of evidence should take into account higher-order uncertainty (as illustrated in @fig-densities). Whenever probability distributions for the probabilities of interest are available (and they should be available for match evidence and many forms of scientific evidence whose reliability has been studied), those distributions should be reported for assessing the value of the evidence. This approach avoids hiding actual aleatory uncertainties under the carpet. It also allows for a more balanced assessment of the evidence, whereas using point values or intervals may exaggerate or underestimate the value of the evidence.
-->

<!--
A couple of clarifications are in order. First, the problem we are highlighting is not confined to match evidence. Say an eyewitness testifies against the defendant: they saw the defendant near the crime scene at the relevant time. To assess the value of this testimony, one should know something analogous to the match probability: if the defendant was not there, how probable is it that the witness would still say the defendant was there? Or suppose a medical test for a disease turns out positive. Here again, to asses the evidential value of the positive test, one should know how probable it is that the test would still turn out positive even when a patient is actually negative. And so on.  These false positive probabilities are usually derived from sample-based frequencies in surveys or experiments: how often witness misidentify people; how often tests misdiagnose; etc. So, depending on the sample size, the false positive probabilities will have different degrees of uncertainty, and the latter should be taken into account when evaluating eyewitness testimonies, diagnostic test results, and many other forms of evidence. 
-->

One final clarification is in order. While the examples we have used in this section involve match probabilities based on samples of different sizes, the problem we are highlighting is not confined to differences in sample size or match probabilities; it is broader than that. Probabilities can be subject to higher-order uncertainty for other reasons, for example, when they are derived from a probability model that has little support, or when the sample size is unrepresentative. The general point of this section is that,  these higher-order uncertainties may be compounded while assessing the probabilities of events in conjunction. It would be a mistake to ignore them, even if all we cared about were first-order probabilities.  


# Bayesian networks
\label{sec:higher-order-networks}



<!---The reader might be worried: how can we handle the computational complexity that comes with moving to higher-order probabilities? The  answer is, as long as we have decent ways of either basing densities on sensible priors and data, or eliciting densities from experts [@o2006uncertain],  implementation is not computationally unfeasible, as we can approximate densities using sampling. 
--->


We looked at simple cases of conjunctions in which the events in question were probabilistically independent. But realistic scenarios are much more complex.
 Think, for example, of two witnesses testifying in a trial about two propositions, say the defendant's whereabouts and the defendant's motive. These two propositions are likely probabilistically dependent.  In these more complex cases, precise probabilism relies on Bayesian networks. The graphical part of a Bayesian network consists of nodes and arrows. Arrows between nodes visually represent relationships of probabilistic dependence between different hypotheses and items of evidence, each corresponding to a node in the network.  The numerical part of a Bayesian network describes the strengths of these dependencies. On a purely formal level, the numerical part consists of probability tables that are filled in with precise prior probabilities (for nodes without incoming arrows) or conditional probabilities (for nodes with incoming arrows). <!--^[The simple case considered in the previous section would  consist of a network with $k+1$ nodes, with a root node for the hypothesis $H$ and then $k$ arrows going from $H$ node to the evidence nodes $E_1, E_2, \dots, E_k$. The probability tables would be filled in with prior probabilities for $H$ and conditional probabilities  $\pr{E_i \vert H}$ and $\pr{E_i \vert \neg H}$, for any item of evidence $E_i$. Notice that these conditional probabilities are those occuring in the individual likelihood ratios $\frac{\pr{E_1 \vert H}}{\pr{E_1\vert \neg H}}$. There is no need to rely on a Bayesian network in such a simple case because the dependencies between nodes are limited.] --> Equipped with these input probabilities, the network can run the calculations about the output probabilities of interest.^[The calculations can quickly get out of hand, so software exists to perform the calculations automatically.] We might be interested, for example, in the probability of a hypothesis given several items of evidence, while keeping track of dependencies between them. In the standard formulation, Bayesian networks run on precise probabilities but can be extended to handle imprecise and higher-order probabilities. This is the topic of this section. 



<!--
The higher-order framework we are advocating is 
not only applicable to the evaluation of 
individual pieces of evidence. Complex bodies of 
evidence and hypotheses---for example, those often represented by 
Bayesian networks---can also be approached from this perspective. The general strategy is this: (1) capture the uncertainties involving the individual items of evidence in a modular fashion using the standard tools for statistical inference. (2) Elicit other probabilities or densities from experts^[For expert elicitation of densities in a parametric fashion and the discussion of the improvement to which doing so instead of eliciting point values leads, see [@o2006uncertain].], (3) put those together using a structure similar to that of a Bayesian network, except allowing for uncertainties of various levels to be put together --- a usual tool for such a representation is a probabilistic program [@Bingham2021PPwithoutTears], and (4) perform inference evaluating the relevant probabilities or densities of interest. 
-->

<!--
If the reader is more used to thinking in terms of Bayesian networks,  a somewhat restrictive  but fairly straightforward way to conceptualize a large class of such programs is to imagine a probabilistic program as stochastically generating Bayesian networks using our uncertainty about the parameter values, update with the evidence, and propagate  uncertainty to approximate the marginal posterior  for nodes of interest.
-->

As an illustration, let us start with a Bayesian network developed by @Fenton2018Risk. \todo{Why is a node for death is missing in the network? That seems necessary as part of the evidence, no?}The network in Figure \ref{fig-scbnplot} represents the key items of evidence in the infamous British case R. v. Clark (EWCA Crim 54, 2000). Sally Clark, the mother of two sons, witnessed her first son die in 1996 soon after birth. Her second son died in similar circumstances a few years later in 1998.  These two consecutive deaths raised suspicion. One hypothesis about the cause of death is that Sally murdered her children. An alternative explanation is that both children died of Sudden Infant Death Syndrome (SIDS). At trial, however, the pediatrician Roy Meadow testified that the probability that a child from a family like the Clark's would die of SIDS was quite low, 1 in 8,543. Assuming probabilistic independence between the two events, the probability of both children dying of SIDS becomes extremely low. It equals the product of the two probabilities, approximately  1 in 73 million. Based on this low probability and signs of bruising on the bodies, Sally Clark was convicted of murder. The conviction was reversed on appeal thanks to new evidence, specifically, signs of a potentially lethal disease found in one of the bodies.

<!--
\textsf{Amurder} and \textsf{Bmurder} are binary nodes corresponding to whether Sally  Clarks  sons,   call  them A and B, were murdered. These  nodes influence  whether  signs of disease (\textsf{Adisease} and \textsf{Bdisease}) and bruising (\textsf{Abruising} and \textsf{Bbruising}) were present. Also, since A's death preceded in time B's death, whether A was murdered casts some light on the probability that B was also murdered.
-->

```{r scBN,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", warning=FALSE, message = FALSE, dpi = 800}
#create SC DAG
#define the structure of the Sally Clark BN
SallyClarkDAG <- model2network("[Abruising|Acause][Adisease|Acause][Bbruising|Bcause][Bdisease|Bcause][Acause][Bcause|Acause][NoMurdered|Acause:Bcause][Guilty|NoMurdered]")
SCdag <- model2network("[Abruising|Acause][Adisease|Acause][Bbruising|Bcause][Bdisease|Bcause][Acause][Bcause|Acause][NoMurdered|Acause:Bcause][Guilty|NoMurdered]")
#plot 
#graphviz.plot(SallyClarkDAG)


#CPTs as used in Fenton & al.
AcauseProb <-prior.CPT("Acause","SIDS","Murder",0.921659)
AbruisingProb <- single.CPT("Abruising","Acause","Yes","No","SIDS","Murder",0.01,0.05)
AdiseaseProb <- single.CPT("Adisease","Acause","Yes","No","SIDS","Murder",0.05,0.001)
BbruisingProb <- single.CPT("Bbruising","Bcause","Yes","No","SIDS","Murder",0.01,0.05)
BdiseaseProb <- single.CPT("Bdisease","Bcause","Yes","No","SIDS","Murder",0.05,0.001)
BcauseProb <- single.CPT("Bcause","Acause","SIDS","Murder","SIDS","Murder",0.9993604,1-0.9998538)

#E goes first; order: last variable through levels, second last, then first
NoMurderedProb <- array(c(0, 0, 1, 0, 1, 0, 0,1,0,1,0,0), dim = c(3, 2, 2),dimnames = list(NoMurdered = c("both","one","none"),Bcause = c("SIDS","Murder"), Acause = c("SIDS","Murder")))

#this one is definitional
GuiltyProb <-  array(c( 1,0, 1,0, 0,1), dim = c(2,3),dimnames = list(Guilty = c("Yes","No"), NoMurdered = c("both","one","none")))

# Put CPTs together
SallyClarkCPTfenton <- list(Acause=AcauseProb,Adisease = AdiseaseProb,
                      Bcause = BcauseProb,Bdisease=BdiseaseProb,
                      Abruising = AbruisingProb,Bbruising = BbruisingProb,
                      NoMurdered = NoMurderedProb,Guilty=GuiltyProb)

# join with the DAG to get a BN
SCfenton <- custom.fit(SallyClarkDAG,SallyClarkCPTfenton)
```




```{r scBNplot2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning=FALSE, message = FALSE, dpi = 800, fig.height= 5, fig.width= 7}
#| label: fig-scbnplot
#| fig-cap: "Bayesian network for the Sally Clark case, with marginal prior probabilities."
#| fig-pos: H


graphviz.chart(SCfenton, type = "barprob", layout = "dot", #draw.labels = TRUE,
  grid = FALSE, scale = c(0.75, 1.1), col = "black", 
  text.col = "black", bar.col = "black", main = NULL,
  sub = NULL)
```

Much has been written about Sally Clark by philosophers and statisticians. The discussion has often focused on whether Meadow was allowed to assume, as he did, that the two SIDS deaths would be independent events.  The assumption of independence delivered the low probability of 1 in 73 million by multiplying  1 in 8,543 by itself. Another much-discussed point was this: even if it was unlikely that two consecutive SIDS deaths would occur, it does not follow it was likely that Sally murdered her children.^[One could reason that, since 1 in 73 million is a low probability, the alternative explanation, that Sally murdered her children, should be likely. But that a mother would do such a thing is also unlikely, perhaps even less likely than 1 in 73 million.]  A Bayesian network helps to avoid these mistakes. It also helps to view the case holistically. The two consecutive deaths were an important piece of evidence, but other evidence was also important, including signs of bruising and signs of a lethal disease as they were discovered during the appeal process.


Unfortunately, Bayesian networks, in their standard formulation, inherit the shortcomings of precise probabilism. The choice of the input probabilities should be precise, and it is often unclear where the values come from or whether they are justified. Consider, for example, the probability that a death by SIDS would occur. How sure are we that this probability equals, exactly, 1 in 8,543? The figure Meadow used is a sample-based frequency. How big was that sample? How representative? Other input probabilities need to be entered in the network to carry out the calculations, for example, the probability that a mother would kill her son, or the probability that signs of bruising would be found given the hypothesis that Sally was trying to murder her child, and so on. There will be uncertainty about these probabilities, since they are based on sample frequencies or expert elicitation. 

The standard response to these concerns is to invoke *sensitivity analysis*: a range of plausible values is tested. Say we are interested in the output probability that Sally is guilty. The network is updated by the known facts---the items of evidence---following standard Bayesian conditionalization.  The input probabilities in the network are then assigned a range of possible values to see how they impact the output probability of Sally's guilt. Sensitivity analysis is another variant, perhaps more rudimentary, of the interval approach we considered earlier.  \todo{add references to imprecise BNs; see footnote.} In fact, Bayesian networks for reasoning with intervals and imprecise probabilities already exist.^[One can use uniform sampling with Bayesian networks to approximate the impreciser's commitments \textbf{Cite} \url{https://arxiv.org/abs/2302.09656}. Another approach is to rely on probabilistic programs with the restriction that the
 variables corresponding to probabilities are sampled from uniform distributions corresponding to the representor set. A critical survey of approaches along these lines shows that, in complex reasoning situations, ''the imprecision of inferences increases rapidly as new premises are added to an argument''.\textbf{add ref to} \url{https://www.sciencedirect.com/science/article/pii/S0004370296000215}.]
But, as discussed earlier, imprecise probabilism, the interval approach and sensitivity analysis ignore the shape of the underlying distributions. They do not distinguish between probability measures (or point estimates) in terms of their plausibility, even though some will be more plausible than others. Moreover, if the sensitivity analysis is only guided by the values at the edges of the interval, these extremes will often play an undeservedly strong role. 


These concerns can be addressed by recourse to higher-order probabilities.  In a precise Bayesian network, each node is associated with a probability table determined by a finite list of numbers (precise probabilities). In an imprecise Bayesian network, each node is associated with a table determined by an interval of numbers. But suppose that, instead of precise numbers or intervals of numbers, we have distributions over the possible numbers to enter into the probability tables.^[The densities of interests can then be approximated by (1) sampling parameter values from the specified distributions, (2) plugging them into the construction of the BN, and (3) evaluating the probability of interest in that precise BN. The list of the probabilities thus obtained will approximate the density of interest. In what follows we will work with sample sizes of 10k.]   An example of such a higher-order Bayesian network for the Sally Clark case is represented in Figure \ref{fig-scwithhop}. 

The higher-order Bayesian network helps to investigate the impact of different items of evidence on Sally Clark's probability of guilt  (see @fig-scwithhop2). The starting point is the prior distribution for the \s{Guilt} node (first graph). Next, the network is updated with evidence showing signs of bruising on both children (second graph). Next, the assumption that both children lack signs of 
potentially lethal disease is added (third graph). Finally, we consider the state of the evidence at the time of the appellate case: signs of bruising existed on both children, but signs of lethal disease were discovered only in the first child. Interestingly, in the strongest case against Sally Clark (third graph), the median of the posterior distribution is above .95, but the uncertainty around that median is still quite wide.^[The lower limit of the 89\% Highest Posterior Density Intervals (HPDI) is at .83.] This underscores the fact that relying on point estimates can lead to overconfidence. 


```{r scStages,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning=FALSE, message = FALSE, dpi = 800}

SCfJN <- compile(as.grain(SCfenton))

priorFenton <- querygrain(SCfJN, node = "Guilty")[[1]][1]

SCfJNAbruising <- setEvidence(SCfJN, nodes = c("Abruising"), states = c("Yes"))
AbruisingFenton <- querygrain(SCfJNAbruising, node = "Guilty")[[1]][1]

SCfJNAbruisingBbruising <- setEvidence(SCfJN, nodes = c("Abruising","Bbruising"),
                                states = c("Yes","Yes"))
AbruisingBbruisingFenton <- querygrain(SCfJNAbruisingBbruising, node = "Guilty")[[1]][1]

SCfJNAbruisingBbruisingNoDisease <- setEvidence(SCfJN, nodes = c("Abruising","Bbruising","Adisease","Bdisease"),
                                         states = c("Yes","Yes", "No", "No"))
AbruisingBbruisingFentonNoDiseaseFenton <-   querygrain(SCfJNAbruisingBbruisingNoDisease, node = "Guilty")[[1]][1]


SCfJNAbruisingBbruisingDiseaseA <- setEvidence(SCfJN, nodes = c("Abruising","Bbruising","Adisease","Bdisease"), 
                                               states = c("Yes","Yes", "Yes", "No"))
AbruisingBbruisingFentonDiseaseAFenton <- querygrain(SCfJNAbruisingBbruisingDiseaseA, node = "Guilty")[[1]][1]


SCfentonTable <- data.frame(stage = factor(c("prior", "bruising in A", "bruising in both",
                                      "bruising in both, no disease", "bruising in both, disease on A only"),
                                      levels = c("prior", "bruising in A", "bruising in both",
                                                 "bruising in both, no disease", "bruising in both, disease on A only")),
                            probability = c(priorFenton,AbruisingFenton,AbruisingBbruisingFenton,
                                            AbruisingBbruisingFentonNoDiseaseFenton,AbruisingBbruisingFentonDiseaseAFenton))
```


<!---
\begin{figure}[H]
```{r SCfentonTable2,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%", warning=FALSE, message = FALSE, dpi = 800}
ggplot(SCfentonTable) + geom_point(aes(x = stage, y = probability, size = probability * .9))+
  scale_x_discrete(limits=rev, expand = c(0, 2)) +coord_flip()+theme_tufte(base_size = 10)+ scale_size(guide="none")+
  theme(plot.title.position = "plot")+ggtitle("Impact of evidence according to Fenton's BN for the Sally Clark case") +
  geom_text(aes(x = stage, y= probability * 1.05, label= round(probability,2) ,hjust=-.3, vjust=-.3), size  = 3.5)+
  scale_y_continuous(breaks = seq(0,.7, by =.1), limits = c(0,.8))

```

\caption{The prior and posterior probabilities for Fenton's Sally Clark BN.}

\label{fig:SCfentonTable}

\end{figure}
--->

<!--
\begin{figure}[H]
```{r FigSCwithHOPa, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%", warning=FALSE, message = FALSE, dpi = 800}

grid.arrange(AbruisingIfSidsPlot+xlim(0,.4),AbruisingIfMurderPlot+xlim(0,.4))
```
\caption{Example of approximated uncertainties about conditional probabilities in the Sally Clark case.}
\label{fig:SCwithHOPa}
\end{figure}
-->


```{r SCwithHOPa, out.extra='angle=90', echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "125%", out.height= "230%", warning=FALSE, message = FALSE, dpi = 800}
#| label: fig-scwithhop
#| fig-cap: "An illustration of a probabilistic program for the Sally Clark case."
#| fig-pos: h

# out.extra='angle=90'               does not seem to work in quarto
# , 


#SCprobsFinal <- readRDS("../datasets/SCprobsFinal.rds")
#attach(SCprobsFinal)
#source("../scripts/SCfunctions.R")
#source("../scripts/SCplotCPTs.R")
#source("../scripts/SCplotDistros.R")


# AsidsPriorPlotGrob <- ggplotGrob(AsidsPriorPlot+theme_tufte(base_size = 5))
# BcauseSidsIfAsidsPlotGrob <- ggplotGrob(BcauseSidsIfAsidsPlot+theme_tufte(base_size = 5))
# BcauseSidsIfAmurderPlotGrob<- ggplotGrob(BcauseSidsIfAmurderPlot+theme_tufte(base_size = 5)) 

# AbruisingIfSidsPlotGrob <- ggplotGrob(AbruisingIfSidsPlot+theme_tufte(base_size = 5)) 
# AbruisingIfMurderPlotGrob <- ggplotGrob(AbruisingIfMurderPlot+theme_tufte(base_size = 5)) 


# AdiseaseIfSidsPlotGrob <- ggplotGrob(AdiseaseIfSidsPlot+theme_tufte(base_size = 5)) 
# AdiseaseIfMurderPlotGrob <- ggplotGrob(AdiseaseIfMurderPlot+theme_tufte(base_size = 5)) 


# BbruisingIfSidsPlotGrob <- ggplotGrob(BbruisingIfSidsPlot+theme_tufte(base_size = 5)) 
# BbruisingIfMurderPlotGrob <- ggplotGrob(BbruisingIfMurderPlot+theme_tufte(base_size = 5)) 


# BdiseaseIfSidsPlotGrob <- ggplotGrob(BdiseaseIfSidsPlot+theme_tufte(base_size = 5)) 
# BdiseaseIfAmurderPlotGrob <- ggplotGrob(BdiseaseIfAmurderPlot+theme_tufte(base_size = 5)) 


# daigramToRotate <- ggplot(data.frame(a=1)) + xlim(1, 40) + ylim(1, 60)+theme_void()+
#   annotation_custom(AsidsPriorPlotGrob, xmin = 9, xmax = 15, ymin = 49, ymax = 59)+
#   geom_label(aes(label = "Bcause", x = 25, y = 41),
#               size = 3 )+
#   geom_label(aes(label = "Acause", x = 12, y = 48),
#             size = 3 )+
#   geom_curve(aes(x = 13.5, y = 48.2, xend = 25, yend = 42.5), curvature = -.18,size = .3,
#              arrow = arrow(length = unit(.015, "npc")))+
#   annotation_custom(BcauseSidsIfAsidsPlotGrob, xmin = 17, xmax = 23, ymin = 47, ymax = 57)+
#   annotation_custom(BcauseSidsIfAmurderPlotGrob, xmin = 15, xmax = 21, ymin = 37, ymax = 47)+
#   geom_label(aes(label = "Abruising", x = 2, y = 41),
#              size = 3 )+
#   geom_curve(aes(x = 10.5, y = 48.2, xend = 2, yend = 42.5), curvature = .2,size = .3,
#              arrow = arrow(length = unit(.015, "npc")))+
#   annotation_custom(AbruisingIfSidsPlotGrob, xmin = 2, xmax = 8, ymin = 47, ymax = 57)+
#   annotation_custom(AbruisingIfMurderPlotGrob, xmin = 5, xmax = 11, ymin = 37, ymax = 47)+
#   geom_label(aes(label = "Adisease", x = 6, y = 21),
#              size = 3 )+
#   geom_curve(aes(x = 13, y = 46.2, xend = 6, yend = 23), curvature = -.45,size = .3,
#              arrow = arrow(length = unit(.015, "npc")))+
#   annotation_custom(AdiseaseIfSidsPlotGrob, xmin = 4.5, xmax = 10.5, ymin = 27, ymax = 37)+
#   annotation_custom(AdiseaseIfMurderPlotGrob, xmin = 9, xmax = 15, ymin = 17.5, ymax = 27.5)+
#   geom_label(aes(label = "Bbruising", x = 14, y = 12),
#              size = 3 ) +
#   geom_curve(aes(x = 24, y = 39.5, xend = 15.5, yend = 13.5), curvature = -.2,size = .3,
#                                     arrow = arrow(length = unit(.015, "npc")))+
#   annotation_custom(BbruisingIfSidsPlotGrob, xmin = 15.5, xmax = 21.5, ymin = 25, ymax = 35)+
#   annotation_custom(BbruisingIfMurderPlotGrob, xmin = 18.5, xmax = 24.5, ymin = 10, ymax = 20)+
#   geom_label(aes(label = "Bdisease", x = 33, y = 18),
#              size = 3 )  +
#   geom_curve(aes(x = 26, y = 39.5, xend = 33, yend = 19.5), curvature = -.2,size = .3,
#              arrow = arrow(length = unit(.015, "npc")))+
#   annotation_custom(BdiseaseIfSidsPlotGrob, xmin = 31, xmax = 37, ymin = 27, ymax = 37)+
#   annotation_custom(BdiseaseIfAmurderPlotGrob, xmin = 24.5, xmax = 30.5, ymin = 20, ymax = 30)

# print(daigramToRotate, vp=viewport(angle=90), width=7, height=5)


# knitr::include_graphics("imp_philosophical_files/figure-pdf/SCwithHOP-1.pdf") # alternative method


```

\todo{N: I am still searching for a good fix of that plot}



<!--
\begin{figure}[htbp] 
    \centering
    \includegraphics[angle=90, width=\textwidth]{imp_philosophical_files/figure-pdf/SCwithHOP-1.pdf}
    \caption{An illustration of a probabilistic program for the Sally Clark case.}
    \label{fig-scwithhop}
\end{figure} 
-->







<!--
```{r SCwithHOP2, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%", warning=FALSE, message = FALSE, dpi = 800, warning = FALSE, message = FALSE, results= 'hide'}
#| label: fig-scwithhop2
#| fig-cap: 'Impact of incoming evidence in the Sally Clark case.'
#| fig-pos: H

grid.arrange(GuiltPriorPlot, GuiltABbruisingPlot, GuiltABbruisingNoDiseasePlot, GuiltABbruisingDiseaseAPlot, ncol =2)
```
-->

<!---
One question that arises is how this approach relates to the standard method of using likelihood ratios to report the value of the evidence. On this approach, the conditional probabilities that are used in the likelihood ratio calculations are estimated and come in a package with an uncertainty about them. Accordingly, these uncertainties propagate: to estimate the likelihood ratio while keeping track of the uncertainty involved, we can sample probabilities from the selected distributions appropriate for the conditional probabilities needed for the calculations, then divide the corresponding samples, obtaining a sample of likelihood ratios, thus approximating the density capturing the recommended uncertainty about the likelihood ratio. Uncertainty about likelihood ratio is just propagated uncertainty about the involved conditional probabilities. For instance, we can use this tool to gauge our uncertainty about the likelihood ratios corresponding to the signs of bruising in son A and the presence of the symptoms of a potentially lethal disease in son A (@fig-sclrs).
-->



<!--
```{r SClrs, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "90%", warning=FALSE, message = FALSE, dpi = 800, warning = FALSE, message = FALSE, results= 'hide'}
#| label: fig-sclrs
#| fig-cap: 'Likelihood ratios forbruising and signs of disease in child A in the Sally Clark case.'
#| fig-pos: H

AbruisingLR <- AbruisingIfMurder / AbruisingIfSids

AbruisingLRPlot <- plotSample(AbruisingLR,title = "LR: Abruising",  paste("median =", 
                                                            round(median(AbruisingLR),2), ", 89%HPDI = ", 
                                                            round(HPDI(AbruisingLR),2)[1],"-",round(HPDI(AbruisingLR),2)[2],
                                                            sep = "") )+xlim(0,30)

AdiseaseLR <- AdiseaseIfMurder / AdiseaseIfSids

AdiseaseLRPlot <- plotSample(AdiseaseLR,title = "LR: Adisease",  paste("median =", 
                                                                           round(median(AdiseaseLR, na.rm = TRUE),2), ", 89%HPDI = ", 
                                                                           round(HPDI(AdiseaseLR),2)[1],"-",round(HPDI(AdiseaseLR),2)[2],
                                                                           sep = "") )+xlim(0,3)
 grid.arrange(AbruisingLRPlot, AdiseaseLRPlot, ncol =2)
```
-->

<!--
Our approach does involve multiple parameters, uncertainty about them, along with a dependency structure between random variables. So it is only natural to ask whether what we propose 
is not just an old wolf in a new sheep's clothing, as one might think that what looks like a DAG and quacks
like a DAG is always a hierarchical model. In this section we briefly clarify what the answer to this question is.
-->

<!--
First, we need some clarity on what a Bayesian hierarchical model is. In the widest sense of the word,
 these are mathematical descriptions involving multiple parameters such that credible values for some of 
 them meaningfully depend on the values of other parameters, and that dependencies can be re-factored into
  a chain of dependencies.  For instance, think about a joint parameter  space   for two parameters $\theta$
   and $\omega$, where $p(\theta, \omega \vert D) \propto p(D \vert \theta, \omega)p(\theta, \omega)$.
    If, further, some independence-motivated re-factoring of the right-hand side---for instance
     as $p(D\vert \theta)p(\theta \omega)p(\omega)$---is possible, we are dealing with a hierarchical
      model in the wide sense of the word.
-->

<!--
Such models usually come useful when we are dealing with clustered data, such as a cohort study with
repeated measures, or some natural groupings at different levels of analysis. Then, lower-level parameters
are treated as i.i.d. and share the same parameter distribution characterized by some hyper-parameters in
turn characterized by a prior distribution. As a simple example consider a scenario in which we are
dealing with multiple coins created by one mint---each coin has its own bias $\theta_i$, but also
there is some commonality as to what these biases are in this mint, represented by a higher-level 
parameter $\theta$.
Continuing the example,  assume $\theta_i \sim \mathsf{Beta}(a, b)$ and 
$y_{i\vert s} \sim \mathsf{Bern}(\theta_s)$, where the former distribution can be re-parametrized as 
$\mathsf{Beta}(\omega(k-2)+1, (1-\omega)(k-2)+1)$. Let's  keep $k$ fixed,  $\omega$ is our expected value of
the $\theta_i$ parameters, with some dispersion around it determined by $k$. Now, if we also are uncertain
about $\omega$ and express our uncertainty about it in terms a density $p(\omega)$, we got ourselves a
hierarchical model with joint prior distribution over parameters $\prod p(\theta_i \vert \omega) p(\omega)$. 
-->

<!--
As  another example, one can develop a multilevel regression model of the distributions of the random levels 
in various counties, where both the intercept and the slope vary with counties by taking  
$y_i\sim \mathsf{Norm}(\alpha_{\mbox{j[i]}}\, + \beta_{\mbox{j[i]}}  x_i, \sigma^2_y )$, where $j$ is a county index, 
$\alpha_j \sim \mathsf{Norm}(\mu_\alpha,\sigma_\alpha^2 )$,  and
$\beta_j \sim \mathsf{No}rm(\mu_\beta,\sigma_\beta^2 )$. Then, running the regression one
estimates both the county-level coefficients, and the higher-level parameters.
-->

<!--
Our approach is similar to standard hierarchical models in the most general sense: there is a meaningful dependence structure and distributions over parameter values that we are working with.
-->

<!--
However, our  approach is unlike such models in a few respects.
 For one, we are not dealing with clustered data, 
and the random variables are mostly propositions and their truth values. Given a hypothesis $H$ and an item
of evidence $E$ for it, there seems to be no interesting conceptualization on which the underlying data would
be clustered. For example, considering stains at a crime scene as a subgroup of crimes being committed doesn't make logical sense.
Yes, there is dependency between these phenomena, 
but describing it as clustering would be at least misleading.  Second, the dependencies proceed through
 the values of the random variables which are **not** parameters, but rather truth-values, and require 
 also conditional uncertainties regarding the dependencies between these truth-values. 
 -->

<!--
Again, continuing the hypothesis-evidence example, we have $H \sim \mathsf{Bern}(p_h)$, 
$p_h \sim \mathsf{Beta}(a_h, b_h)$, and  $E\sim \mathsf{Bern}(p_e)$. But then we also have the beta distributions
for the probability of the evidence conditional on the actual values of the random variables---the truth-values---thus 
$p_e \vert H = 1 \sim beta(a_{+}, b_{+} )$ and  $p_e \vert H = 0 \sim \mathsf{Beta}(a_{-}, b_{-})$.
But the re-factoring in terms of the actual values of the random variables (which just happen to resemble 
probabilities because they are truth values) makes it quite specific,  at the same time allowing for the 
computational use of a probabilistic program.  Finally, the reasoning we describe is not  a regression the
 way it is normally performed:  the learning task is delegated to the bottom level of whatever happens to the 
 Bayesian networks once updated with evidence. 
We would prefer to reserve the term \emph{hierarchical model} for a class of models dealing with interesting 
cluster structures in the data. 
-->

<!--
A more fitting term, however, for the representation tool we propose should be used here is \emph{probabilistic programs}. We do not claim any originality in devising this tool: it's an 
already existing tool. What we argue for, though, is its ability for being usefully deployed in the
 context of evidence evaluation and integration with other assumptions and hypotheses.  
 -->

# Conclusion



 We have argued that higher-order probabilism outperforms both precise and imprecise probabilism. It is able to model scenarios that the other two cannot model, for example, the case of uneven bias. In addition, higher-order probabilism does not fall prey to difficulties peculiar to imprecise probabilism, such as belief inertia and the lack of proper scoring rules. We have also identified a novel set of problems for precise and imprecise probabilism, mostly stemming from the question of how to evaluate, in the aggregate, the probabilities of multiple propositions. Here again, higher-order probabilism fares better. 

Some might dislike the idea of going higher-order for a number of reasons, for example, unnecessary complexity. This is a line taken by Bradley, who refuses to go higher-order for the following reason:

\begin{quote}
Why are sets of probabilities the right level to stop the regress at? Why not sets of sets? Why not second-order probabilities? Why not single probability functions? This is something of a pragmatic choice. The further we allow this regress to continue, the harder it is to deal with these belief-representing objects. So let's not go further than we need. 131-132\end{quote}


\noindent
We have shown that given the difficulties of precise and imprecise probabilism, we are not going further than we need in using higher-order probabilities. The pragmatic concerns one might have are at best unclear.

<!--  parameter uncertainty, approximations
 and other computational methods are alrady embedded in Bayesian statistical practice and good computational already exist.\footnote{Also, you can insist that instead of going higher order we could just take our sample space to be the cartesian product of the original sample space and parameter space, or use parameters having certain values as potential states of a bayesian network.  If you prefer not to call such approaches first-order, I don't mind, as long as you effectively end up
    assigning probabilities to certain probabilities, the representation means I discussed in this paper
     should be in principle available to you.}
 -->


We should unnderscore that, mathematically, we do not
 propose anything radically new. Concepts from the Bayesian toolkit that can model higher-order uncertainty already exist. Our suggestion is that they have been under-appreciated in formal epistemology and should be more widely used. This is not to say that there is no need for any novel technical work. For example, we still need a proper accuracy argument in defense of higher-order probabilism. Will an agent who relies on higher-order probabilities would accuracy-dominate one who relies on just first-order probabilities? We leave this as an open question. Another concern is the lack of clear semantics for higher-order probabilities. While a more elaborate account is beyond the scope of this paper, the answer 
should gesture at a modification of the framework of 
probabilistic frames [@Dorst2022higher-order;@Dorst2022evidence]. Start with a set of possible worlds $W$.
Suppose you consider a class of probability distributions $D$, a finite list of atomic sentences
$q_1, \dots, q_2$ corresponding to subsets of $W$, and a selection of true probability hypotheses 
$C$ (think of the latter as omniscient distributions, $C\subseteq D$, but in principle this restriction 
can be dropped if need be). Each possible world $w\in W$ and a proposition $p\subseteq W$ come with their 
true probability distribution, $C_{w,p}\in D$ corresponding to the true probability of $p$ in $w$, 
and the distribution that the expert assigns to $p$ in $w$, $P_{w,p}\in D$. Then, various propositions 
involving distributions can be seen as sets of possible worlds, for instance, the proposition that the expert 
assigns $d$ to $p$ is the set of worlds $w$ such that $P_{w,p}=d$.\footnote{There is at least one important 
difference between this approach and that developed by Dorst. His framework is untyped, which allows for 
an enlightening discussion of the principle of reflection and alternatives to it. In this paper, we prefer 
to keep this complexity apart and use an explicitly typed set-up.} 



# Appendix: the strict propriety of $I_{kl}$ {-}

The fact that $I_{KL}$ is strictly proper as applied to second-order probabilities is not very surprising.  However, in the existing literature, the proof is not usually explicitly given,  and some of the pieces are not present in the philosophical literature. So we tried to include the whole chain of thought, warning that some of these results are already known and all we did was making the proofs more presentable, and pointing out new elements in the reasoning.  Let us start with a definition of concavity.

\begin{definition}[concavity]

A function $f$ is convex over an interval $(a,b)$ just in case for all  $x_1, x_2\in (a,b)$ and 
$0 \leq \lambda \leq 1$ we have:
\begin{align*}
f(\lambda x_1 + (1-\lambda)x_2) \leq \lambda f(x_1) + (1-\lambda)f(x_2)
\end{align*}

\noindent A function $f$  is concave just in case:
\begin{align*}
f(\lambda x_1 + (1-\lambda)x_2) \geq \lambda f(x_1) + (1-\lambda)f(x_2)
\end{align*}
\noindent A function $f$  is strictly concave just in case the equality holds only if either $\lambda = 0$ or $\lambda = 1$.
\end{definition}

For us it is important that if a function is twice differentiable on an interval, then it is (strictly) 
concave just in case its second derivative is non-positive (negative). In particular, as $(\log_2(x))'' = -\frac{1}{x^2 ln(2)}$, $\log_2$ is strictly concave over its domain.\footnote{I line with the rest of the paper, we'll work with $\log$ base 2. We could equally well use any other basis.} 




\begin{lemma}[Jensen's inequality]
If $f$ is concave, and $g$ is any function of a random variable, $\mathbb{E}(f(g(x))) \leq f(\mathbb{E}(g(x)))$. If $f$ is 
strictly concave, the equality holds only if $g(x) = \mathbb{E} g(x)$, that is, if $g(x)$ is constant everywhere.
\end{lemma}




\begin{proof}
For the base case consider a two-point mass probability function. Then,
\begin{align*}
p_1f(g(x_1))+ p_2f(g(x_2)) &\leq f(p_1g(x_1) + p_2g(x_2))
\end{align*}
\noindent follows directly from the definition of concativity, if we take $\lambda = p_1$, $(1-\lambda)=p_2$,
 and substitute $g(x_1)$ and $g(x+2)$ for $x_1$ and $x_2$.



Now, suppose that  $ p_1f(g(x_1))+ p_2f(g(x_2)) = f(p_1g(x_1) + p_2g(x_2))$ and that   $f$ is strictly concave.
 That means either $(p_1 = 1\wedge p_2 = 0)$, or $(p_1 = 0 \wedge p_2 =1)$. Then either $x$ always takes
  value $x_1$, in the former case, or always takes value $x_2$, in the latter
   case. $\mathbb{E} g (x) =  p_1 g(x_1) + p_2 g(x_2)$, which equals  $g(x_1)$ in the former case and $g(x_2)$ in the latter.


Now suppose Jensen's inequality and the consequence of strict contativity holds for $k-1$ mass points. 
Write $p_i' = \frac{p_i}{1-p_k}$ for $i = 1, 2, \dots, k-1$. We now reason:
\begin{align*}
\sum_{i=1}^k p_i f(g(x_i)) & =
 p_kf(g(x_k)) + (1-p_k)\sum_{i =1}^{k-1}p_i'f(g(x_i)) &\\
 & \leq p_k f(g(x_k)) + (1-p_k)f\left( \sum_{i = 1}^{k=i}p_i' g(x_i) \right) & \mbox{\footnotesize by 
 the induction hypothesis}\\ &\leq f\left( p_k(g(x_k)) + (1-p_k)\sum_{i = 1}^{k-1} p_i' g(x_i)\right) & 
 \mbox{\footnotesize by the base case} \\
 & = f \left( \sum_{i}^k p_i g(x_i)\right)
 \end{align*}

Notice also that at the induction hypothesis application stage, we know that the equality holds only if 
$p_k =1 \vee p+k = 0$. In the former case $g(x)$ always takes value $x_k = \mathbb{E} g(x)$. In the latter case,
 $p_k$ can be safely ignored and $\sum_{i=1}^{k}p_ig(x_i) = \sum_{i=1}^{k-1}p'g(x_i)$ and by the induction 
 hypothesis we already know that $\mathbb{E} g(x) = g(x)$.


\end{proof}



In particular, the claim holds if we take $g(x)$ to be $\frac{q(x)}{p(x)}$ (were both $p$ and $q$ are 
probability mass functions), and  $f$ to be $\log_2$. Then, given that $A$ is the support set of $p$, we have:
\begin{align*}
\sum_{x\in A}p(x) \log_2 \frac{q(x)}{p(x)} & \leq \log_2 \sum_{x\in A}p(x)\frac{q(x)}{p(x)}
\end{align*}

\noindent Moreover, the equality holds only if $\frac{q(x)}{p(x)}$ is constant, that is, only if $p$ and $q$ 
are the same pmfs. Let's use this in the proof of the following lemma.

\begin{lemma}[Information inequality] For two probability mass functions $p, q$, $\dkl(p,q)\geq 0$ with 
equality iff $p=q$.
\end{lemma}


\begin{proof}
Let $A$ be the support set of $p$, and let $q$ be a probability mass function whose support is $B$.
\begin{align*}
- \dkl(p,q) & = - \sum_{x\in A}p(x) \log_2 \frac{p(x)}{q(x)}& \mbox{\footnotesize (by definition)} \\
&  =  \sum_{x\in A}p(x)  - \left(\log_2 p(x) - \log_2 q(x)\right)& \\
&  =  \sum_{x\in A}p(x)   \left(\log_2 q(x) - \log_2 p(x)\right)& \\
& =  \sum_{x\in A} p(x) \log_2 \frac{q(x)}{p(x)}& \\
& \leq \log_2 \sum_{x\in A} p(x)\frac{q(x)}{p(x)} & \mbox{\footnotesize by Jensen's inequality}\\
& \mbox{(and the equality holds only if $p = q$)}\\
& = \log_2 \sum_{x\in A} q(x)  & \\
& \leq \log_2 \sum_{x\in B} q(x) & \\
& = log (1)  = 0 &\\
\end{align*}
\end{proof}


Observe now that $\dkl$ can be decomposed in terms of cross-entropy and entropy.

\begin{lemma}[decomposition] $\dkl = H(p,q) - H(p)$. \end{lemma}



\begin{proof}
\begin{align*}
\dkl (p, q) & = \sum_{p_{i}} \left( \log_2 p_i - \log_2 q_i \right) \\
& =   - \sum_{p_{i}}\left( \log_2 q_{i} - \log_2 p_{i} \right) \\
& = - \sum_{p_{i}} \log_{2} q_{i} - \sum_{p_{i}} - \log_{2} p_{i}   \\
& -  \underbrace{- \sum_{p_{i}} \log_2 q_{i}}_{H(p,q)}    - \underbrace{- \sum_{p_i}  \log_2 p_{i}}_{H(p)}
\end{align*}
\end{proof}


With information inequality this easily entails  Gibbs' inequality:

\begin{lemma}[Gibbs' inequality] $H(p,q) \geq H(p)$ with identity only if $p = q$.
\end{lemma}



We are done with our theoretical set-up, which is already common knowledge, except presented in an orderly manner in one place. Now we present our argument for the claim that the above entails the propriety of $I_{KL}$. 
First, let's systematize the notation.
Consider a discretization of the parameter space $[0,1]$ into $n$ equally spaced values $\theta_1, \dots, \theta_n$. 
For each $i$ the ``true'' second-order distribution if the true parameter indeed is $\theta_i$---we'll call it 
the indicator of $\theta_i$--- which is defined by
\begin{align*}
Ind^k(\theta_i) & = \begin{cases} 1 & \mbox{if } \theta_i = \theta_k\\
                        0 & \mbox{otherwise}  \end{cases}
\end{align*}
\noindent We will write $Ind^k_i$ instead of $Ind^k(\theta_i)$.


Now consider a probability distribution $p$ over this parameter space, assigning probabilities $p_1, \dots, p_n$
 to $\theta_1, \dots, \theta_n$ respectively. It is to be evaluated in terms of inaccuracy from the perspective 
 of a given 'true' value $\theta_k$. The inaccuracy of $p$ if $\theta_k$ is the 'true' value, is the 
 divergence between $Ind^k$ and $p$. 

\begin{align*}
I_{KL}(p, \theta_k) & = D_\text{KL}(Ind^k||p) \\
& = \sum_{i=1}^n Ind^k_i \left( \log_2 Ind^k_i - \log_2 p_i \right)
\end{align*}
Note now that for $j \neq k$ we have $Ind^k_j = 0$  and so $Ind^k_j \left( \log_2 Ind^k_j - \log_2 p_j \right)=0$. 
Therefore we continue:
\begin{align*}
& = Ind^k_k \left( \log_2 Ind^k_k - \log_2 p_k \right)
\end{align*}
Further, $Ind^k_k= 1$ and therefore $\log_2 Ind^k_k =0$, so we simplify:
\begin{align*}
& =  - \log_2 p_k
\end{align*}

\noindent Now, let's think about expected values. First, what is the inaccuracy of $p$ as expected by $p$, $\mathit{EI}_{\text{DK}}(p,p)$?

\begin{align*}
\mathit{EI}_{\text{DK}}(p,p) & = \sum_{i =1}^n p_i I_{\text{DK}}(p, \theta_i) \\
& = \sum_{i =1}^n  p_i - \log_2 p_i \\
& = - \sum_{i =1}^n  p_i  \log_2 p_i = H(p)
\end{align*}

\noindent Analogously, the inaccuracy of $q$ as expected from the perspective of $p$ is:

\begin{align*}
\mathit{EI}_{\text{DK}}(p, q) & =   \sum_{i =1}^n p_i \left( - \log_2 q_i\right)\\
& = -  \sum_{i =1}^n p_i  \log_2 q_i = H(p,q)
\end{align*}


But that means, by Gibbs' inequality, that $\mathit{EI}_{\text{DK}}(p,q) \geq \mathit{EI}_{\text{DK}}(p,p)$ unless $p=q$, which completes the proof.



#  References {-}

# Evidence aggregation: the simple case - SET ASIDE FOR BOOK

<!--
As we have seen, higher-order probabilism outperforms both precise and imprecise probabilism, at the descriptive as well as the normative level. From a descriptive standpoint, higher-order probabilism can model scenarios that cannot be modeled by the other versions of probabilism. From a normative standpoint, accuracy maximization may sometimes recommend that a rational agent represent their credal state with a distribution over probability values rather than a precise probability measure\todo{add ref to vFraasen in fn; perhaps extend the discussion a bit } (more on this soon).^[Having read van Fraasen's "Laws and Symmetry", you might also worry that going higher order somehow leads to a contradiction; we will address this concern later on.]  \todo{where do we show that accuracy recommends a distribution over probability measures? This bit seems missing.}
-->

 Rational agents are often tasked with aggregating pieces of evidence and assessing their value relative to a hypothesis. In this and the next section, we examine the question of how multiple items of evidence should be evaluated together. This question raises novel difficulties for both precise and imprecise probabilism. We show how higher-order probabilism can handle them. 

For the precise probabilist, a natural measure of the value of the evidence is the likelihood ratio.  This ratio is relative to a pair of competing hypotheses, say $H$ and its negation $\neg H$ (though the two hypotheses need not be one the negation of the other). Relative to these hypotheses, the likelihood ratio of a single piece of evidence $E$ is the probability of $E$ given  $H$ divided by the probability of $E$ given $\neg H$, or in short, $\frac{\pr{E \vert H}}{\pr{E \vert \neg H }}$.
<!--
In qualitative terms, evidence 
$E$ supports $H$ over its negation insofar as the probability of $E$ is 
higher given hypothesis $H$ than given its negation, that is, the likelihood ratio is greater than one. 
-->
Degrees of evidential value (or support, strength) can be expressed as follows:
\begin{quote}
the higher $\frac{\pr{E \vert H}}{\pr{E \vert \neg H }}$ (if greater than one), the more strongly $E$ supports $H$.
\end{quote}

\noindent
The value of the evidence increases whenever $\pr{E \vert H}$ increases or whenever $\pr{E \vert \neg H }$ decreases. The higher $\pr{E \vert H}$, the better the evidence at tracking $H$ (a true positive); the lower $\pr{E \vert \neg H }$, the better the evidence at avoiding $\neg H$ (a true negative).
If the probability of $E$ is the same given hypothesis $H$ as given its negation, that is, the likelihood ratio equals one, the evidence would have no value for $H$.

Likelihood ratios can also be used for assessing the value of multiple pieces of evidence in the aggregate, again relative to a pair of hypotheses of interest. In the simplest case (for more complex cases, see the next section), 
multiple items of evidence all bear on the 
same hypothesis. Then, to obtain their combined evidential value, it is enough to multiply 
their individual likelihood ratios. 
$$
\frac{\pr{E_1 \wedge E_1 \dots E_k \vert H}}{\pr{E_1 \wedge E_2 \dots E_k \vert \neg H}} = \frac{\pr{E_1 \vert H}}{\pr{E_1 \vert \neg H}}\times \frac{\pr{E_2 \vert H}}{\pr{E_2 \vert \neg H}} \times \dots \times \frac{\pr{E_k \vert H}}{\pr{E_k \vert \neg H}}
$$

```{r introStarts-OLD,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE}
ps <- seq(0,1,length.out = 1001)

hairMean <-  29/1148
hairA <- 30
hairB <- 1149

dogMean <- 2/78
dogA <- 3
dogB <- 79


lik0 <- hairMean * dogMean
prior <- seq(0,.3, by = 0.001)
priorH0 <- 1-prior
denomin <- lik0 * priorH0 + prior
num <-  lik0 * priorH0
posterior <- 1- num/denomin
threshold <- min(prior[posterior > .99])

pointImpactPlot <- ggplot()+geom_line(aes(x = prior, y = posterior))+xlim(0,.07)+
  theme_tufte(base_size = 10)+labs(title = "Prior vs. posterior, based on point estimates",
                                   subtitle = "Joint evidence: dog & hair")+
  geom_vline(xintercept = threshold, lty = 2, size =.5, alpha = .7)+
  annotate(geom = "label", label = "posterior > .99", x = .067, y =.95, size = 2.5)+
  theme(plot.title.position = "plot")
```

\noindent
The equality holds provided $E_1, E_2, \dots, E_k$ are probabilistically independent conditional on hypothesis $H$ and its negation. Think, for example, at several diagnostic tests performed by independent laboratories or independent witnesses in a trial testifying about the same issue. 

<!--For the precise probabilist, filling in the precise probabilities in the formula above gives an assessment of the value of aggregate evidence. But, as we will now see, this approach risks over- or under-estimating the value of the evidence. (We will focus on likelihood ratios, but the claim we defend can be generalized to other measures of evidential value and posterior probabilites.^[On different measures of degrees of evidential value (or support, confirmation), see \textbf{CITE FITELSON}. Posterior probabilities are obtained from the likelihood ratios via the prior probabilities. The posterior odds $\frac{\pr{H \vert E}}{\pr{\neg H \vert E }}$ are obtained by multiplying the likelihood ratio $\frac{\pr{E \vert H}}{\pr{E \vert \neg H }}$ by the prior odds $\frac{\pr{H}}{\pr{\neg H}}$. The posterior probability $\pr{H \vert E}$ is obtained from the formula $PO/(1+PO)$, where $PO$ is short for 'posterior odds'. This result only holds if the two hypotheses compared are one the negation of the other.]) To see what the problem is. 
-->

<!--  as follows:
$$
\frac{\pr{H \vert E_1 \wedge E_2\dots E_k}}{\pr{\neg H \vert E_1 \wedge E_1\dots E_k}} = \frac{\pr{E_1 \wedge E_1 \dots E_k \vert H}}{\pr{E_1 \wedge E_2\dots E_k \vert \neg H}} \times \frac{\pr{H}}{\pr{\neg H}} 
$$

We will focus on likelihood ratios for ease of exposition.
--->


To see how likelihood ratios can be deployed, it is worth working through a specific case. In a murder case, the police recover two items of trace evidence, both against the defendant. First, hair found at the crime scene matches the defendant's hair; call this evidence '\textsf{hair}.' Second, the fur of the defendant's dog matches the fur found in a carpet wrapped around one of the bodies; call this evidence '\textsf{fur}.'^[The hair evidence and the dog fur evidence are stylized after two 
 items of  evidence in the notorious 1981 Wayne Williams case [@deadman1984fiber1; @deadman1984fiber2].] 
The two matches favor the hypothesis that the defendant (and the defendant's dog) must be the source of the crime traces; call this hypothesis '$\mathsf{source}$'. If the two matches are independent lines of evidence (conditional on the source hypothesis and its negation), their likelihood ratios can be multiplied:

<!--
[Strictly speaking, it is  possible for $A$ and $B$ to be independent conditional on $H$, but not conditional on $\neg H$. Here, we require both independencies to hold.]
-->
 
$$
\frac{{\pr{\s{fur} \wedge \s{hair}  \vert \s{source}}}}{{\pr{\s{fur} \wedge \s{hair} \vert \neg \s{source}}}} = \frac{{\pr{\s{fur} \vert \s{source}}}}{{\pr{\s{fur}  \vert \neg \s{source}}}} \times
 \frac{{\pr{\s{hair} \vert \s{source}}}}{{\pr{\s{hair} \vert \neg \s{source}}}}
$$



So far so good. But how do we fill in the precise probabilies? The numerators can be equated to one: if the defendant is a contributor, the laboratory will declare a match for sure. This is a simplication, but it will do for our purposes. To fill in the denominators, a trial expert will provide so-called match probabilities. They express the likelihood that, by coincidence, a random person (or a random dog) who is not a contributor would still match.  The match probabilities are approximated by counting how many matches are found in a representative sample of the human population (or the canine population). Suppose the matching hair type occurs  `r round(hairMean,4)` times in a reference database, and the matching dog fur type occurs  `r round(dogMean,4)` times in a reference database (more on how these numbers 
are calculated soon). These frequencies can fill in the match probabilities. 
Putting everything together:
\begin{align*}
\frac{{\pr{\s{dog} \vert \s{source}}}}{{\pr{\s{dog}  \vert \neg \s{source}}}} \times
 \frac{{\pr{\s{hair} \vert \s{source}}}}{{\pr{\s{hair} \vert \neg \s{source}}}}
& =  \frac{1}{`r hairMean`} \times  \frac{1}{`r dogMean`} = {`r 1/(hairMean * dogMean)`}
\end{align*}

\noindent
The resulting ratio is large. The two matches, combined, strongly favor the source hypothesis. 
<!--
```{r impactOfPoint4,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", warning = FALSE, message = FALSE}
#| label: fig-impactOfPoint
#| fig-cap: "Impact of dog fur and human hair evidence on the prior, point estimates."
#| fig-pos: H

pointImpactPlot
```
-->

```{r hair-OLD,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
#carpetSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, carpetA, carpetB))
set.seed(231)
hairSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, hairA, hairB))
dogSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, dogA, dogB))

#carpetHPDI <- HPDI(carpetSamples, prob  =.9)
hairHPDI <- HPDI(hairSamples, prob  =.99)
dogHPDI <- HPDI(dogSamples, prob  =.99)
```

```{r charitableImpact-OLD,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
lik0l <- .037 * .103
denominL <- lik0l * priorH0 + prior
numL <-  lik0l * priorH0
posteriorL <- 1- numL/denominL
thresholdL <- min(prior[posteriorL > .99])

charitableImpactPlot <- ggplot()+geom_line(aes(x = prior, y = posteriorL))+xlim(0,.32)+
  theme_tufte(base_size = 10)+labs(title = "Prior vs. posterior, charitable reading",
                                   subtitle = "Joint evidence: dog & hair")+
  geom_vline(xintercept = thresholdL, lty = 2, size =.5, alpha = .7)+
  annotate(geom = "label", label = "posterior > .99", x = .305, y =.95, size = 2.5)+ylab(
    "posterior"
  )+
  theme(plot.title.position = "plot")
```


This is the story about evidence aggregation told by the precise probabilist.  But this story misses something crucial. As it happens, the match probability for hair evidence is based on 29 matches found in a sample database of size 1148, while the match probability for the dog evidence is based on finding two matches in a smaller database of size 78. The relative frequencies are about $.025$ in both cases, but the two samples differ in size. The smaller the sample, the greater the uncertainty about the match probabilities. So, for individual pieces of evidence, simply reporting the exact numbers makes it seem as though the evidential value of the matches is the same, but actually it is not.^[The match probabilities in the Wayne Williams case  on which our running example is based were 1/100 for the dog fur, and 29/1148 for Wayne Williams' hair. Match probabilities have been slightly but not unrealistically modified to be closer to each other in order to make a conceptual point: the same first-order probabilities, even when they sound precise, may come with different degrees of second-order uncertainty.] In the aggregate, multiplying the individual likelihood ratios further washes away this difference. 

A better alternative is easily available: the evaluation of multiple items of evidence should take into account higher-order uncertainty. @fig-densities (upper part) depicts higher-order probability distributions of different match probabilities given the sample data---the actual number of matches found in the sample databases. <!---By hypothesis, 29 matches were found in the sample database of human hair of size 1,148, and 2 matches were found in the sample database of dog fur of size 78.---> As expected, some random match probabilities are more likely than others, and since the sizes of the two databases are different, the distributions have different spreads: the smaller the database the greater the spread, the greater the uncertainty about the match probability. In light of this, @fig-densities (lower part) depicts the probability distribution for the joint match probability associated with both items of match evidence, hair and fur evidence. The aggregate value of the two pieces of match evidence, then, is given by a distribution over possible likelihood ratios. \todo{add distributions of LRs figure} The shape of this distribution conveys the degree of higher-order uncertainty about the value of the aggregate evidence. \textbf{Marcello (Rafal/Nikodem to add): Can we have a formula of how the two matches are combined in the higher-order approach? In precise probabilism, you multiply the individual LRs, in higher-order probabilism, what do we do formally? Can we also have a distribution of likelihood ratios? What happens if both numerator and denominator in the LR are distributions?} 


```{r LR_dist,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%"}

# initial calculations #################

# individuall likelihood ratio distribution

lh_hair <- rep(1, 1e4) / hairSamples
lh_dog <- rep(1, 1e4) / dogSamples

# joined likelihood ratio distribution
lh_joint <- lh_hair * lh_dog

# taking densities, wrangling, visualizing

dens_hair <- density(lh_hair)

dens_hair_df <- data.frame(
  x = dens_hair$x,
  y = dens_hair$y
  )
  
dens_dog <- density(lh_dog)

dens_dog_df <- data.frame(
  x = dens_dog$x,
  y = dens_dog$y
)

disjoined_dists <- ggplot() +
  geom_density(stat = 'identity')  +
  geom_density(data = dens_hair_df, aes(x= x, y= y, color = "hair"), stat = 'identity') +
  geom_density(data = dens_dog_df, aes(x= x, y= y, color = "dog"), stat = 'identity') +
  labs(title= 'Disjoined LR distribution', y = 'Density', x = 'Likelihood Ratio', color= 'LR distribution:') +
  scale_color_manual(values = c("dog" = "steelblue", "hair" = "purple")) +
  guides(color = guide_legend(override.aes = list(fill = c( "steelblue", "purple"))))+ 
  theme_tufte(base_size = 10) +
  theme(legend.position = c(0.6, 0.5)) 

lh_joint_dens <- density(lh_joint)

lh_joint_dens_df <- data.frame(
  x = lh_joint_dens$x,
  y = lh_joint_dens$y
)

df_hpdi <- lh_joint_dens_df %>% 
  filter(x >= HPDI(lh_joint, 0.9)[1]) %>% 
  filter(x <= HPDI(lh_joint, 0.9)[2])

joined_dists <- ggplot() +
  geom_density(data = lh_joint_dens_df, aes(x= x, y = y), stat = 'identity', linewidth= 1)  +
  geom_density(data = df_hpdi ,aes(x = x, y = y), stat = 'identity', fill= '#cccccc', linewidth= 0)+
  labs(title = 'Joined LR distribution', y = "Density", x = 'Likelihood Ratio')+ theme_tufte(base_size = 10)+
  geom_text(aes(x = 20000, y = 0.00015, label = "Mean: 1607"), color = "steelblue")+ # calculated with mean(lh_joint)
  geom_text(aes(x = 20000, y = 0.0001, label = "HPDI 0.9: 326 | 2976"), color = "steelblue") # HPDI(lh_joint, 0.9)

```



```{r LR_dist_vis,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "90%", warning = FALSE, message = FALSE}
#| label: fig-lrdens
#| fig-cap: "Distributions of dog and hair likelihood ratios and the resulting joint likelihood ratio. Created with the samples from beta distributions. Shaded area on the second one represents HPDI with 0.9 credibility."
#| fig-pos: H

grid.arrange(disjoined_dists,joined_dists, ncol = 2 )
```




```{r densitiesEvidence-OLD,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%"}

jointEvidence <- dogSamples * hairSamples


densities1Plot <- ggplot()+
  geom_line(aes(x = ps, y = dbeta(ps, hairA, hairB)), lty  = 2)+
  geom_line(aes(x = ps, y = dbeta(ps, dogA, dogB)), lty = 3)+xlim(0,.15)+
  xlab("probability")+
  ylab("density")+
  theme_tufte(base_size = 10)+
  #  annotate(geom  = "label", label = "carpet", x =  0.045, y = 140)+
  annotate(geom  = "label", label = "hair", x =  0.035, y = 80)+
  annotate(geom  = "label", label = "dog", x =  0.06, y = 15)+
#  labs(title = "Conditional densities for  individual items of evidence if the source hypothesis is false")+
#  theme(plot.title.position = "plot")
labs(title = "Distributions of individual match probabilities")+
  theme(plot.title.position = "plot")


densities2Plot <- ggplot()+
  xlab("probability")+
  ylab("density")+
  theme_tufte(base_size = 10)+
  geom_density(aes(x= jointEvidence))+
  geom_vline(xintercept = 0.002760, lty = 2, linewidth = .5)+
  geom_vline(xintercept = 0.000023, lty = 2, linewidth  = .5)+
  geom_vline(xintercept = 0.000144, lty = 3, linewidth = .8)+
  geom_vline(xintercept = 0.001742, lty = 3, linewidth  = .8)+
  #labs(title = "Conditional density for joint evidence",
  labs(title = "Distribution for the joint match probability",
       subtitle = "(with .99 and .9 HPDIs)")+
  theme(plot.title.position = "plot") +
  geom_vline(xintercept = hairMean * dogMean)
```


```{r Figdensities-OLD,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
#| label: fig-densities-OLD
#| fig-cap: "Beta densities for individual items of evidence and the resulting joint density with .99 and .9 highest posterior density intervals, assuming the sample sizes as discussed and independence, with uniform priors."
#| fig-pos: H

grid.arrange(densities1Plot,densities2Plot, ncol = 1 )
```


The precise probabilist might insist that the value of the evidence---one item or multiple items of evidence---is naturally captured by a precise likelihood ratio. In our running example, this ratio equals one over the first-order match probabilty, and our best assessment of this first-order probability is still the relative frequency of matches found in the database, whether large or small. If this is right, our best assessment of the match probabilities for both fur and hair evidence should be about $.025$, based on the relative frequencies 2/78 and 29/1148.   If we were to bet whether a dog or a human picked at random would have the matching fur or hair type, our odds should be $.025$ no matter the size of the database. This argument has some bite when evaluating single items of evidence. In fact, the expected values of the match probabilities for hair and match evidence---based on the higher-order distributions in  @fig-densities (upper part)---still end up being about $.025$. If, as the precise probabilist assumes, first-order probabilities is all we should care about, going higher-order would seem a needless complication.

This line of reasoning, however, breaks down when evaluating two or more items of evidence. What should our betting odds be for the proposition that a human and a dog, both picked at random, would have the matching fur and hair type in question? For the precise probabilist, the answer is straightforward: on the assumption of independence, it is enough to multiply the $.025$ individual match probabilities and obtain a joint match probability of `r hairMean * dogMean`. The higher-order probabilist will proceed differently. In assessing first-order match probabilities, they will retain information about higher-order uncertainty as much as possible. This can done in two steps: first, aggregate the higher-order distributions for the two match probabilities and obtain a higher-order probability distribution for the joint match probability (see @fig-densities); next, to obtain our best assessment of the first-order joint match probability, take the expected value of this latter  distribution. The higher-order probabilist will assign `r mean(jointEvidence)` to the joint match probability, a value greater than what the precise probabilist would assign. 



```{r densitiesEvidenceModified-OLD,echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%"}

# example with slightly different numbers 
# illustrate divergence between precise and higher order probabilist 
# on value of first-order joint match probabilities

# sample frequency, very small sample size
ps <- seq(0,1, length.out = 1001)
set.seed(231)

evidence1Mean <-  1/20
evidence1A <- 1
evidence1B <- 20

evidence2Mean <- 1000/20000
evidence2A <- 1000
evidence2B <- 20000

evidence1Samples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, evidence1A, evidence1B))

evidence2Samples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, evidence2A, evidence2B))


jointEvidence1 <- evidence1Samples * evidence1Samples
jointEvidenceThree1 <- evidence1Samples^3
jointEvidenceFive1 <- evidence1Samples^5
jointEvidenceSeven1 <- evidence1Samples^7 

jointEvidence2 <- evidence2Samples * evidence2Samples
jointEvidenceThree2 <- evidence2Samples^3
jointEvidenceFive2 <- evidence2Samples^5
jointEvidenceSeven2 <- evidence2Samples^7 

```


So, the higher-order and precise probabilist will disagree  about the betting odds for the proposition that a human and a dog, both picked at random, would have the matching fur and hair type. The disagreement will become even starker as a larger number of independent items of evidence are evaluated.^[Consider the simple case of independent items of evidence whose individual match probabilities are $.025$. For three, five and seven items of evidence, the joint match probabilities would be: `r (evidence1Mean)^3`,`r (evidence1Mean)^5` and `r (evidence1Mean)^7` (for the precise probabilist) and `r mean(jointEvidenceThree1)`, `r mean(jointEvidenceFive1)` and `r mean(jointEvidenceSeven1)` (for the higher-order probabilist, based on small databases of size 20).] Who should be trusted? Since the higher-order probabilist takes into accout more information---that is, the higher-distributions---there is good reason to think that the higher-order probabilist should be trusted more than the precise probabilist.^[As a further illustration of this point, consider a couple of variations of our running example. First, suppose the match probabilities associated with two matches are both set to $.05$, since they are based on the following relative frequencies: one match occurs in a dog fur database and one match occurs in a human hair database, where both databases are small, say of size 20. By multiplying the individual $1/.05$ likelihood ratios associated with the two matches, their evidential value against the defendant would seem quite strong: $1/.05\times 1/.05=`r 1/(evidence1Mean*evidence1Mean)`$. But the match probabilities are based on frequencies resulting from small databases, so their evidential value should be rather weak. Precise probability here seems to exaggerate the aggregate value of the evidence. Following higher-order probabilism, the  joint likelihood ratio would be `r 1/mean(jointEvidence1)`, a significantly smaller value. On the other hand, if the same $.05$ match probabilities were based on larger databases, the evidential value of the two matches should be correspondingly greater, but precise probabilism would make no difference. If, for example, 1,000 hair and fur matches are found in databases of size 20,000, the higher-order probabilist would assign `r 1/mean(jointEvidence2)` to the joint likelihood ratio, a much greater value than before. This outcome agrees with our intuituions.] 

Imprecise probabilism will also run into its own problems when assessing the value of aggregate evidence. Recall that the probability measures in the representor set are those compatible with the evidence.  The problem is that almost any random match probability will be compatible with any sample data---with any number of matches found in a reference database. This point should be familiar from the earlier discussion. Think by analogy to coin tossing: even a coin that has a .99 bias toward tails could come up heads on every toss. This series of outcomes is unlikely, but possible. Similarly, a hair type that has a match probability extremely small could still be found several times in a sample population. So, it is not clear how to proceed if one takes seriously the binary notion of compatibility. Imprecise probabilism is too permissive because almost any match probability will be compatible with the data. 

Another option for the imprecise probabilist is to rely on reasonable ranges of match probabilities. Suppose these ranges are (.015,.037) (.002, .103), for hair and fur evidence respectively in our original case.^[These are 99% credible intervals starting with uniform priors. A 99\% credible interval is the narrowest interval to which the   expert thinks the true parameter belongs with probability .99. For a discussion of what credible intervals are, how they differ from confidence intervals, and why confidence intervals should not be used, see @kruschke2015doing.] 
<!-- Now, the representor covers
  the convex hull of the probability measures that result in probabilities that are the edges of the intervals. For this reason, to investigate what span of probabilities the imprecise probabilities will end with,
  -->
As expected, the range is wider for dog fur match evidence than hair match evidence: the uncertainty about the dog fur match probability is greater since the sample database was smaller. This is a desirable feature of the interval approach.  Now, to assess the joint uncertainty, it is enough to focus on what happens at the edges of the two intervals. Reasoning with representor members at the edges of the intervals will yield the most extreme probability measure the impreciser is committed to, the worst-case and best-case scenarios. We end up with a new range for the joint match probabilties, (.00003, .003811).\footnote{Redoing the calculations using the upper bounds of the two intervals, $.037$ and $.103$,  yields the following:
\begin{align*}
\mathsf{P}(\s{dog}\wedge \s{hair} \vert \neg \s{source})   & =  .037 \times .103 =.003811.
\end{align*}

\noindent
 This number is around `r round(.003811/lik0,2)` times greater than the original estimate. Given this number, the two matchs are much weaker evidence for the source hypothesis than previously thought.    The calculation for the lower bounds, $.015$ and $.002$, yields the following:
\begin{align*}
\mathsf{P}(\s{dog}\wedge \s{hair} \vert \neg \s{source})   & =  .015 \times .002 =.00003
\end{align*}
   
 \noindent  
This number is around `r round(.0003/lik0,2)` times lower than the original estimate.  Given this number, the two matchs are much stronger evidence for the source hypothesis than previously thought.}
The corresponding likelihood ratios could be as high as  `r round(1/0.00003,4)` or as low as `r round(1/0.003811,4)`. The two matches could be much stronger or much weaker evidence than previously thought.

 <!--Now the prior probability of the source hypothesis needs to be higher than `r thresholdL` for the posterior probability to be above .99 (@fig-impactofcharitable). -->

<!--
```{r FigcharitableImpact75,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", warning = FALSE, message = FALSE}
#| label: fig-impactofcharitable
#| fig-cap: "Impact of dog fur and human hair evidence on the prior, charitable reading."
#| fig-pos: H

charitableImpactPlot
```
-->


<!--
In general, it is impossible to calculate the credible interval for the joint distribution
 based solely on the individual credible intervals corresponding to the individual events.
   We need additional information: the distributions that were used to calculate the intervals 
   for the probabilities of the individual events. In our example, if you additionally knew, 
   for instance, that the expert used  beta distributions (as, arguably, they should in this context),
    you could in principle calculate the  99\% credible interval for the joint distribution.
     It usually will not be the same as whatever the results of multiplying the individual
      interval edges, and it is unlikely that a human fact-finder would be able to correctly 
      run such calculations in their head even if they knew the functional form of the distributions 
      used.^[Also, in principle, in more complex contexts, we need  further information about how the 
      items of evidence are related if we cannot take them to be independent.] 
      So providing the fact-finder with individual intervals, even if further information 
      about the distributions is provided, might easily mislead.^[Investigation of the 
       extent to which the individual interval presentation is misleading  would be an interesting psychological study.]
-->

<!--
As it turns out, given the reported sample sizes, the 99\% credible interval
 for the probability $\mathsf{P}(\s{dog}\wedge \s{hair} \vert \neg \s{source})$ is $(0.000023,  0.002760)$.
The upper bound of this interval would then require the prior probability of the source hypothesis to be above .215
 for the posterior to be above .99. On this interpretation, the two items of match evidence 
 are still not quite as strong as you initially thought, but stronger than what your second calculation indicated. 
-->



Using plausible ranges for the match probabilities leaves the impression that any value in the interval is just as good as any other. Perhaps we should pick the middle value as representative of the interval. However, relying on the entire interval or the middle value will misrepresent the evidence.  To see why, consider again @fig-densities (lower part) which depicts the probability distribution for the joint match probability. Interestingly, this distribution is not symmetric. So the most likely value (and the bulk of the distribution, really) does not lie in the middle between the edges. <!--Just because the parameter lies in an interval with some posterior probability, it does not mean that the ranges near the edges of the interval are equally likely---the bulk of the density might very well be closer to one of the edges. --> Therefore, only relying on the edges---or taking central values as representative of the interval---can lead to overestimating or underestimating the probabilities at play.\footnote{The calculations for the joint interval assume that because the worst- or best-case probability for one event is $x$ and the worst- or best-case probability for another independent event is $y$, the worst- or best-case probability for their conjunction is  $xy$. However, this conclusion does not follow if the margin of error (credible interval) is fixed. Just because the probability of an extreme value $x$ for one variable $X$ is .01, and so it is for the value $y$ of another independent variable $Y$, it does not follow that the probability that those two independent variables take values $x$ and $y$ simultaneously is the same. In general, it is impossible to calculate the credible interval for the joint distribution based solely on the individual credible intervals corresponding to the individual events.}
 

```{r, lratiosCalc, echo= FALSE, eval=TRUE}

# imprecise intervals: hair - (.015,.037), dog - (.002, .103)
hair_low <- .015
hair_high <- .037

dog_low <- .002
dog_high  <- .103

hair_LR_width <- 1/hair_low - 1/hair_high
dog_LR_width <- 1/dog_low - 1/dog_high

# aggregating two items of evidence

aggregated_low <- 1/ (hair_low * dog_low)
aggregated_high <- 1/ (hair_high * dog_high)

# posteriors ???



```


Another problem in taking intervals as representative of the value of the evidence is that they will tend to widen as more items of evidence are evaluated. The size of the likelihood ratio interval was initially `r round(hair_LR_width, 2)` (hair evidence)  and `r round(dog_LR_width, 2)` (fur evidence). After aggregating the two items of evidence, the likelihood ratio interval widened to `r aggregated_low - aggregated_high`. The size of the match probability interval was initially `r hair_low - hair_high` (hair evidence)  and `r dog_low - dog_high` (fur evidence). 
After aggregating the two items of evidence, the match probability interval narrowed to `r hair_low * dog_low - hair_high * dog_high`. Posterior interval (starting with 1:1 prior odds) was initially `r (1/.015)/(1+1/.015) - (1/.037)/(1+1/.037)` (hair evidence)  and `r (1/.002)/(1+1/.002) - (1/.103)/(1+1/.103)` (fur evidence). After aggregating the two items of evidence, the posterior interval narrowed to `r (1/.00003)/(1+1/.00003) - (1/.003811)/(1+1/.003811)`. 

\todo{LR intervals widen but match and posterior probability intervals do not? How does that work? How can we claim that uncertainty increases?}

 All in all, precise and imprecise probabilism do not fare well in modelling the value of evidence in the aggregate. 
 Instead, the evaluation of multiple items of evidence should take into account higher-order uncertainty (as illustrated in @fig-densities). Whenever probability distributions for the probabilities of interest are available (and they should be available for match evidence and many forms of scientific evidence whose reliability has been studied), those distributions should be reported for assessing the value of the evidence. This approach avoids hiding actual aleatory uncertainties under the carpet. It also allows for a more balanced assessment of the evidence, whereas using point values or intervals may exaggerate or underestimate the value of the evidence.

A couple of clarifications are in order. First, the problem we are highlighting is not confined to match evidence. Say an eyewitness testifies against the defendant: they saw the defendant near the crime scene at the relevant time. To assess the value of this testimony, one should know something analogous to the match probability: if the defendant was not there, how probable is it that the witness would still say the defendant was there? Or suppose a medical test for a disease turns out positive. Here again, to asses the evidential value of the positive test, one should know how probable it is that the test would still turn out positive even when a patient is actually negative. And so on.  These false positive probabilities are usually derived from sample-based frequencies in surveys or experiments: how often witness misidentify people; how often tests misdiagnose; etc. So, depending on the sample size, the false positive probabilities will have different degrees of uncertainty, and the latter should be taken into account when evaluating eyewitness testimonies, diagnostic test results, and many other forms of evidence. At the same time---and this is the second clarification---this discussion is not meant to suggest that the problem we are highlighting is confined to differences in sample size; it is broader than that. Probabilities can be subject to uncertainty for other reasons, for example, when they are derived from a probability model for which there is little support, or when the sample size is large but unrepresentative. So, in short, the problem of higher-order uncertainty is widespread and goes beyond match evidence and questions of sample size.



<!--
```{r densLines, echo=FALSE,eval=TRUE,fig.align = "center",cache=FALSE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
prior <- seq(0,1, by = 0.001)
priorH0 <- 1-prior

#-----

jointPosterior <- list()
minima <- numeric(1e4)


# for each likelihood from sample,
# calculate the posterior based on what the prior is
# and find the minimum above threshold
for (s in 1:1e4){
  lik <- jointEvidence[s]
  denomin <- lik * priorH0 + prior
  num <-  lik * priorH0
  posterior <- 1- num/denomin
  jointPosterior[[s]] <- posterior
  minima[s] <- min(prior[posterior > .99])
  }



#minimaPlot <- ggplot()+geom_density(aes(x = minima))+
#  theme_tufte(base_size = 10)+ggtitle("Minimal priors sufficient for posterior >.99")#+xlab("minimal prior")+
#  theme(plot.title.position = "plot")

#minimaGrob <- ggplotGrob(minimaPlot)



#jointPosteriorDF <-   as.data.frame(do.call(cbind, jointPosterior))



#jointPosteriorSubsample <- jointPosteriorDF[,1:300]

#jointPosteriorDF$ps <- ps
#jointPosteriorSubsample$ps <- ps

#jointPosteriorLong <- melt(jointPosteriorSubsample,
#                           id.vars = c("ps"))
#colnames(jointPosteriorLong) <- c("prior", "sample","probability")

#alpha = .25
#size = .2

#densitiesLinesPlot <- ggplot(jointPosteriorLong)+
#  geom_line(aes(x = prior, y = probability,group = sample),
#            alpha = alpha, linewidth = size)+
#  theme_tufte()+xlim(0,.2) +
#  annotation_custom(minimaGrob, xmin = .085, xmax = .185,
#                  ymin = 0.01, ymax = 0.8) +
#  ggtitle("Posterior vs prior (300 sampled lines)") +
#  theme(plot.title.position = "plot")
#densitiesLinesPlot





minimaPlot <- ggplot()+geom_density(aes(x = minima))+
  theme_tufte(base_size = 10)+ggtitle("Minimal priors sufficient for posterior >.99")+xlab("minimal prior")+
  theme(plot.title.position = "plot", plot.title = element_text(size = 10))

jointPosteriorDF <-   as.data.frame(do.call(cbind, jointPosterior))

jointPosteriorSubsample <- jointPosteriorDF[,1:300]

jointPosteriorDF$ps <- ps
jointPosteriorSubsample$ps <- ps

jointPosteriorLong <- melt(jointPosteriorSubsample,
                           id.vars = c("ps"))
colnames(jointPosteriorLong) <- c("prior", "sample","probability")

alpha = .25
size = .2

densitiesLinesPlot <- ggplot(jointPosteriorLong)+
  geom_line(aes(x = prior, y = probability,group = sample),
            alpha = alpha, linewidth = size)+
  theme_tufte()+xlim(0,.2) + theme_tufte(base_size = 10)+
  ggtitle("Posterior vs prior (300 sampled lines)") +
  theme(plot.title.position = "plot", plot.title = element_text(size = 10))

```
-->


<!--
```{r Figlines,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
#| label: fig-lines
#| fig-cap: "300 lines illustrating the uncertainty about the dependence of the posterior on the prior given aleatory uncertainty about the evidence, with the distribution of the minimal priors required for the posterior to be above .99."
#| fig-pos: H

grid.arrange(densitiesLinesPlot, minimaPlot, ncol= 2)
```
-->







```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
priorA <- 1
priorB <- 1
set.seed(215)
trueH <- runif(1,0,1)

sampleSize <- sample(10:20,size = 1)
testSize <- sampleSize

set.seed(319)
successes <- rbinom(1, sampleSize, trueH)
pointEstimate <- successes/sampleSize

pointPredictions <-  rbinom( 1e4 , size = testSize , prob = pointEstimate )

ps <- seq(0,1,length.out = 1001)

testPredictions <- rbinom( 1e4 , size = testSize , prob = trueH )

posterior <- function(x) dbeta(x, priorA + successes,
                               priorB + sampleSize - successes)

posteriorSample <- sample( ps , size=1e4 ,
                           replace=TRUE , posterior(ps)/sum(posterior(ps)))

posteriorPredictions <- rbinom( 1e4 , size=testSize , prob=posteriorSample )


testProbs <- table(factor(testPredictions, levels = seq(0,sampleSize)))/1e4
pointProbs <- table(factor(pointPredictions, levels = seq(0,sampleSize)))/1e4
posteriorProbs <- table(factor(posteriorPredictions, levels = seq(0,sampleSize)))/1e4

testPlot <- ggplot()+geom_bar(aes(x= testPredictions, y = ..prop..))+
  ggtitle(paste("Predictions based on the true parameter = ", round(trueH,2), sep = ""))+xlim(0,testSize)+ylab("probability")+xlab("successes")+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")


pointPlot <- ggplot()+geom_bar(aes(x= pointPredictions,y = ..prop..))+
  labs(title = paste("Predictions based on the point estimate = ", round(pointEstimate,2)), subtitle = paste(successes, " successes in ", sampleSize, " observations", sep = ""))+xlim(0,testSize)+ylab("probability")+xlab("successes")+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")


samplesPlot <- ggplot()+geom_density(aes(x= posteriorSample))+
  ggtitle(paste("Posterior sample from beta(", 1+successes, ",", 1+testSize - successes,
                ")", sep = ""))+xlab(
                  "parameter value"
                )+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")

posteriorPlot <- ggplot()+geom_bar(aes(x= posteriorPredictions,y = ..prop..))+
  ggtitle("Predictions based on the posterior sample")+xlim(0, testSize)+ylab("probability")+xlab("successes")+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")


```



```{r kldsCalculations,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
klds <- numeric(1000)
for(i in 1:1000){
trueH <- runif(1,0,1)
sampleSize <- sample(10:25,size = 1)
testSize <- sampleSize
successes <- rbinom(1, sampleSize, trueH)
pointEstimate <- successes/sampleSize
pointPredictions <-  rbinom( 1e4 , size = testSize , prob = pointEstimate )
ps <- seq(0,1,length.out = 1001)
testPredictions <- rbinom( 1e4 , size = testSize , prob = trueH )
posteriorSample <- sample( ps , size=1e4 ,
                           replace=TRUE , posterior(ps)/sum(posterior(ps)))
posteriorPredictions <- rbinom( 1e4 , size=testSize , prob=posteriorSample )
testProbs <- table(factor(testPredictions, levels = seq(0,sampleSize)))/1e4
pointProbs <- table(factor(pointPredictions, levels = seq(0,sampleSize)))/1e4
posteriorProbs <- table(factor(posteriorPredictions, levels = seq(0,sampleSize)))/1e4
klds[i] <- kld(testProbs,pointProbs) - kld(testProbs,posteriorProbs)
}

```




