---
title: "Title"
author: "Marcello Di Bello and Rafal Urbaniak"
date: '`r Sys.Date()`'
output:
  bookdown::pdf_book: 
    keep_tex: yes
  bookdown::gitbook:
    lib_dir: "book_assets"
documentclass: book
classoptions: dvipsnames, usenames
site: bookdown::bookdown_site
bibliography: [bibliographyRafal.bib]
biblio-style: 
link-citations: yes
csl: apa-5th-edition.csl
---

<!-- github-repo: rstudio/bookdown-demo -->


```{r setup, include=FALSE, echo=FALSE}
library(bookdown)
library(tidyverse)
library(knitr)

options(kableExtra.latex.load_packages = FALSE)
library(kableExtra)
```



Here we can scribble the outline draft!  


\begin{enumerate}

\item Chapter blah.

\end{enumerate}


# Preface {-}


testing again



\begin{align} 
  f\left(k\right) = \binom{n}{k} p^k\left(1-p\right)^{n-k}
  (\#eq:binom)
\end{align} 


This is a citation [@diamond90] which uses keys from the bib file listed in the preamble.

Equation \@ref(eq:binom)^[This is a footnote containing a double citation [@diamond90; @dahlmanNakedStat2020].]

Note that chapter files are found and compiled automatically, but the file names have to contain chapter numbers first. For instance, we used `01-intro.Rmd`, placed in the same folder. Observe how we included r code inline.



```{r fig-margin, fig.margin=TRUE}
plot(cars)
```

