---
title: "Second-order Probability, Accuracy and Weight of Evidence"
author: "Rafal Urbaniak and Marcello Di Bello"
date: "November 24, 2022"
output:
  pdf_document:
    number_sections: true
    df_print: kable 
    keep_tex: true
    toc: yes
    includes:
      in_header:
        - Rafal_latex7.sty
classoption: dvipsnames,enabledeprecatedfontcommands
fontsize: 10pt
documentclass: scrartcl
urlcolor: blue
bibliography: [../../references/referencesMRbook.bib]
csl: [../../references/apa-6th-edition.csl]
indent: true
---





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dagitty)
library(rethinking)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(tidyr)
library(philentropy)
library(latex2exp)
library(gridExtra)
library(rethinking)
library(bnlearn)
library(gRain)
library(reshape2)
library(truncnorm)
library(ggforce)

ps <- seq(0,1, length.out = 1001)
getwd()
source("../scripts/CptCreate.R")
source("../scripts/SCfunctions.R")

SCprobsFinal <- readRDS("../datasets/SCprobsFinal.rds")
attach(SCprobsFinal)      
source("../scripts/SCplotCPTs.R")
source("../scripts/SCplotDistros.R")
```



\vspace{2cm}

\noindent \textbf{DISCLAIMER:} \textbf{This is a draft of work in progress, please do not cite or distribute without permission.}

\thispagestyle{empty}

\newpage

\begin{quote} \textbf{Abstract.}  \todo{need to write one when done}

\end{quote}

\inbook{this is what a comment about what will go into book looks like; will be globally supressed when generating latex for the journal paper, don't worry about deleting them.}

# Introduction

<!---
\textbf{M's comment: Nice exmaple and discussion. The only thing I would change (besides being clearer on certain stuff, see comments below) is make clear that we are after three strategies: precise probabilism (use precise probabilities for rando match probabilities), interval approach (senstivity analysys), and then higher-order approach. I think that sequence of moves should emerge from the discussion as it foreshadoes what is to come.)}
--->


```{r introStarts,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE}
ps <- seq(0,1,length.out = 1001)

hairMean <-  29/1148
hairA <- 30
hairB <- 1149

dogMean <- 2/78
dogA <- 3
dogB <- 79

lik0 <- hairMean * dogMean
prior <- seq(0,.3, by = 0.001)
priorH0 <- 1-prior
denomin <- lik0 * priorH0 + prior
num <-  lik0 * priorH0
posterior <- 1- num/denomin
threshold <- min(prior[posterior > .99])

pointImpactPlot <- ggplot()+geom_line(aes(x = prior, y = posterior))+xlim(0,.07)+
  theme_tufte(base_size = 10)+labs(title = "Prior vs. posterior, based on point estimates",
                                   subtitle = "Joint evidence: dog & hair")+
  geom_vline(xintercept = threshold, lty = 2, size =.5, alpha = .7)+
  annotate(geom = "label", label = "posterior > .99", x = .067, y =.95, size = 2.5)+
  theme(plot.title.position = "plot")
```


A defendant in a criminal case may face multiple 
items of incriminating evidence whose strength 
can at least  sometimes be assessed using probabilities. 
For example, consider a murder case 
in which the police recover trace 
evidence that matches the defendant. Hair found at the crime 
scene matches the defendant's hair (call this evidence \textsf{hair}).
In addition, the defendant owns a dog whose 
fur matches the dog fur found in a carpet wrapped around 
one of the bodies (call this evidence \textsf{dog}).^[The hair evidence and the dog fur evidence are stylized after two items of  evidence in the notorious 1981 Wayne Williams case [@deadman1984fiber1; @deadman1984fiber2].] 
The two matches suggest that the defendant (and the defendant's dog) 
must be the source of the crime traces (call this hypothesis $\mathsf{source}$). But how strong is this evidence, really?  What are the fact-finders to make of it? 
<!---Consider two different  items of match evidence:^[These are stylized after two items of  evidence in the notorious Wayne Williams case. Probabilities have been slightly but not unrealistically shifted to be closer to each other to make a conceptual point. The original probabilities were  1/100 for the dog fur, and 29/1148 for Wayne Williams' hair.]  The suspects dog's fur matches the dog fur found in a carpet wrapped around one of the bodies (\textsf{dog}). A hair found on one of the victims matches that of the suspect (\textsf{hair}). What are the fact-finders to make of this evidence? To start with, some probabilistic evaluation thereof should be useful. 
--->

The standard story among legal probabilists 
goes something like this. To evaluate the strength of the two items of match evidence, we must find the value of the likelihood ratio:
\[\frac{\pr{\s{dog}\wedge \s{hair} \vert \s{source}}}{\pr{\s{dog}\wedge \s{hair} \vert \neg \s{source}}}\]
For simplicity, the numerator can be equated to one. 
To fill in the denominator, an expert provides the relevant random match probabilities. Suppose the expert testifies that the probability of a random person's hair matching the reference sample is about `r round(hairMean,4)`, and the probability of a random dog's hair matching the reference sample happens to be about the same, `r round(dogMean,4)`.^[Probabilities have been slightly but not unrealistically modified to be closer to each other in order 
to make a conceptual point. The original probabilities were  1/100 for the dog fur, and 29/1148 for Wayne Williams' hair. We modified the actual reported probabilities slightly to emphasize the point that we will elaborate further on: the same first-order probabilities, even when they sound precise, may come with different degrees of  second-order uncertainty.] 
<!---result from various items of evidence connected to various levels of second-order uncertainty.--->
<!---You assume that the probabilities of matches if the suspect (respectively, the suspect's dog) is the source is one, and that --->
Presumably, the two matches are independent lines of evidence. In other words, their random match probabilities <!---of a match---> must be independent of each other conditional on <!---either truth value of---> the source hypothesis. <!---, namely that the defendant (or the defendant's dog) is the source of the hair (or dog fur) found at the scene --->  <!----and $\neg \mathsf{source}$).---> Then, to evaluate the overall impact of the evidence on the source hypothesis, you calculate: 
\begin{align*}
\pr{\s{dog}\wedge \s{hair} \vert \neg \s{source}} & = \pr{\s{dog} \vert \neg \s{source}} \times \pr{\s{hair} \vert \neg \s{source}} \\
& =  `r hairMean` \times  `r dogMean` = `r hairMean * dogMean`
\end{align*}
This is a very low number. Two such random matches would be quite a coincidence. Following our advice from Chapter 5, the expert  facilitates your understanding  of how this low number should be interpreted. They show you how
the  items of match evidence change the probability of the source hypothesis given a range of possible priors (Figure \ref{fig:impactOfPoint}).  The posterior of .99 is reached as soon as the prior is higher than  `r threshold`.^[These calculations assume that the probability of a match if the suspect and the suspect's dog are the sources is one.]
While perhaps not sufficient for outright belief in the source hypothesis, the evidence seems extremely strong: a minor additional piece of evidence could make the case against the defendant overwhelming. 


\begin{figure}[H]
```{r impactOfPoint4,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", warning = FALSE, message = FALSE}
pointImpactPlot
```
\caption{Impact of dog fur and human hair evidence on the prior, point estimates.}
\label{fig:impactOfPoint}
\end{figure}

Unfortunately, this  analysis 
leaves out something crucial. You reflect on what you have been told and ask the expert: how can you know the random match probabilities with such precision? Shouldn't we also be mindful of the uncertainty that may affect these numbers? The expert agrees, and tells you that in fact the random match probability for the hair evidence  is based on 29 matches found in a database of size 1148, while the random match probability for the dog evidence is based on finding two matches in a reference database of size 78. 


```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
#carpetSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, carpetA, carpetB))
set.seed(231)
hairSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, hairA, hairB))
dogSamples <- sample(ps, 1e4, replace = TRUE, prob = dbeta(ps, dogA, dogB))

#carpetHPDI <- HPDI(carpetSamples, prob  =.9)
hairHPDI <- HPDI(hairSamples, prob  =.99)
dogHPDI <- HPDI(dogSamples, prob  =.99)
```

```{r charitableImpact,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
lik0l <- .037 * .103
denominL <- lik0l * priorH0 + prior
numL <-  lik0l * priorH0
posteriorL <- 1- numL/denominL
thresholdL <- min(prior[posteriorL > .99])

charitableImpactPlot <- ggplot()+geom_line(aes(x = prior, y = posteriorL))+xlim(0,.32)+
  theme_tufte(base_size = 10)+labs(title = "Prior vs. posterior, charitable reading",
                                   subtitle = "Joint evidence: dog & hair")+
  geom_vline(xintercept = thresholdL, lty = 2, size =.5, alpha = .7)+
  annotate(geom = "label", label = "posterior > .99", x = .305, y =.95, size = 2.5)+ylab(
    "posterior"
  )+
  theme(plot.title.position = "plot")
```


The expert's answer makes apparent that the precise random match probabilities do not tell the whole story. Perhaps, the information about sample sizes is good enough  and now you know how to use the evidence properly.^[This is what, effectively, CITE TARONI seem to suggest when they insist the fact-finders should be simply given point estimates and information about the study set-up, such as sample size. As will transpire, we disagree.] But if you are like most human beings, you can't. What to do, then?  
\todo{added this bit to draw attention to this aspect of the Taroni debate, to come back to this}

You ask the expert for guidance:  what are reasonable ranges of the random match probabilities? What are the worst-case and best-case scenarios? The expert responds with 99% credible intervals---specifically, starting with uniform priors, the ranges of the random match probabilities are (.015,.037) for hair evidence and (.002, .103) for fur evidence.^[Roughly, the 99\% credible interval is the narrowest interval to which the expert thinks the true parameter belongs with probability .99. For a discussion of what credible intervals are, how they differ from confidence intervals, and why confidence intervals should not be used, see Chapter 3.] With this information, you redo your calculations using the upper bounds of the two intervals: $.037$ and $.103$. The rationale for choosing the upper bounds is that these numbers result in random match probabilities that are most favorable to the defendant. Your new calculation yields the following:
\begin{align*}
\mathsf{P}(\s{dog}\wedge \s{hair} \vert \neg \s{source})   & =  .037 \times .103 =.003811.
\end{align*}
 This number is around `r round(.003811/lik0,2)` times greater than the original estimate. Now the prior probability of the source hypothesis needs to be higher than `r thresholdL` for the posterior probability to be above .99 (Figure \ref{fig:impactOfCharitable}). So you are no longer convinced that the two items of match evidence are strongly incriminating.

\begin{figure}[H]
```{r fig:charitableImpact7,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", warning = FALSE, message = FALSE}
charitableImpactPlot
```
\caption{Impact of dog fur and human hair evidence on the prior, charitable reading.}
\label{fig:impactOfCharitable}
\end{figure}


This result is puzzling.  Are the two items of match evidence strongly incriminating evidence (as you initially thought) or somewhat weaker (as the new calculation suggests)? For one thing,  using precise random match probabilities might be too unfavorable toward the defendant. On the other hand, your new assessment of the evidence based on the upper bounds might be too *favorable* toward them. Is there a middle way that avoids overestimating and underestimating the strength of the evidence?



To see what this middle path looks like, 
we should reconsider the calculations you just did. 
You made an important blunder: you assumed that because the worst-case probability for one event is $x$ and the worst-case probability for another independent event is $y$, the worst-case probability  for their conjunction is  $xy$. But this conclusion does not follow if the margin of error (credible interval) is fixed. The intuitive reason is  simple: just because the probability of an extreme (or larger absolute) value $x$ for one variable $X$ is .01, and so it is for the value  $y$ of another independent variable $Y$, it does not follow that the probability that those two independent variables take values $x$ and $y$ simultaneously is the same. This probability is actually much smaller.  The interval presentation instead of doing us good led us into error. 


In general, it is impossible to calculate the credible interval for the joint distribution based solely on the individual credible intervals corresponding to the individual events.  We need additional information: the distributions that were used to calculate the intervals for the probabilities of the individual events. In our example, if you additionally knew, for instance, that the expert used  beta distributions (as, arguably, they should in this context), you could in principle calculate the  99\% credible interval for the joint distribution. It usually will not be the same as whatever the results of multiplication of individual interval edges, and it is unlikely that a human fact-finder would be able to correctly run such calculations in their head even if they knew the functional form of the distributions used. ^[Also, in principle, in more complex contexts, we need  further information about how the items of evidence are related if we cannot take them to be independent.] So providing the fact-finder with individual intervals, even if further information about the distributions is provided, might easily mislead.^[Investigation of the  extent to which the individual interval presentation is misleading  would be an interesting psychological study.]
\todo{Can you google to see if there is any such study?}


As it turns out, given the reported sample sizes, the 99\% credible interval for the probability $\mathsf{P}(\s{dog}\wedge \s{hair} \vert \neg \s{source})$ is $(0.000023,  0.002760)$.
<!-- ^[The 99\% credible interval (or a 99\% margin of error) is not the 99\% confidence interval known from classical statistics. There are various reasons not to use these, already discussed in Chapter 3. Another sense, in which we mean it here, is the range to which the true value belongs with posterior probability of 99\% given the evidence.  Normally we mean highest posterior density intervals, that is the narrowest intervals with this property.] -->
\todo{the fn was repetitive, compare to fn 5}

The upper bound of this interval would then require the prior probability of the source hypothesis to be above .215 for the posterior to be above .99. On this interpretation, the two items of match evidence are still not quite as strong as you initially thought, but stronger than what your second calculation indicated. 




Still, the interval approach---even the corrected version just outlined---suffers from a more general problem. 
Working with intervals might be useful if the underlying distributions are fairly symmetrical. But in our case, they might not be. For instance, Figure \ref{fig:densities} depicts beta densities for dog fur and human hair, together with sampling-approximated density for the joint evidence.  The distribution for the joint evidence is not symmetric. <!---, and so switching the margin of error moves the right edge of the interval much faster towards lower values.---> If you were only informed about the edges of the interval, you would be oblivious to the fact that the most likely value (and the bulk of the distribution, really) does not simply lie in the middle between the edges. Just because the parameter lies in an interval with some posterior probability, it does not mean that the ranges near the edges of the interval are equally likely---the bulk of the density might very well be closer to one of the edges. Therefore, only relying on the edges  can lead one to either overestimate or underestimate the probabilities at play. 
This also means that---following our advice on how to illustrate the impact of evidence on prior probabilities---a better representation of the dependence of the posterior on the prior should comprise multiple possible sampled lines whose density mirrors the density around the probability of the evidence (Figure \ref{fig:lines}).


```{r densitiesEvidence,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}

jointEvidence <- dogSamples * hairSamples


densities1Plot <- ggplot()+  
  geom_line(aes(x = ps, y = dbeta(ps, hairA, hairB)), lty  = 2)+
  geom_line(aes(x = ps, y = dbeta(ps, dogA, dogB)), lty = 3)+xlim(0,.15)+
  xlab("probability")+
  ylab("density")+
  theme_tufte(base_size = 10)+
  #  annotate(geom  = "label", label = "carpet", x =  0.045, y = 140)+
  annotate(geom  = "label", label = "hair", x =  0.035, y = 80)+
  annotate(geom  = "label", label = "dog", x =  0.06, y = 15)+
  labs(title = "Conditional densities for  individual items of evidence if the source hypothesis is false")+
  theme(plot.title.position = "plot")

densities2Plot <- ggplot()+  
  xlab("probability")+
  ylab("density")+
  theme_tufte(base_size = 10)+
  geom_density(aes(x= jointEvidence))+
  geom_vline(xintercept = 0.002760, lty = 2, linewidth = .5)+
  geom_vline(xintercept = 0.000023, lty = 2, linewidth  = .5)+
  geom_vline(xintercept = 0.000144, lty = 3, linewidth = .8)+
  geom_vline(xintercept = 0.001742, lty = 3, linewidth  = .8)+
  labs(title = "Conditional density for joint evidence", 
       subtitle = "(with .99 and .9 HPDIs)")+
  theme(plot.title.position = "plot")
```



\begin{figure}[H]
```{r fig:densities,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning = FALSE, message = FALSE}
grid.arrange(densities1Plot,densities2Plot, ncol = 1 )
```
\caption{Beta densities for individual items of evidence and the resulting joint density with .99 and .9 highest posterior density intervals, assuming the sample sizes as discussed and independence, with uniform priors.}
\label{fig:densities}
\end{figure}




```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", warning = FALSE, message = FALSE}
prior <- seq(0,1, by = 0.001)
priorH0 <- 1-prior

#-----

jointPosterior <- list()
minima <- numeric(1e4)


#for each likelihood from sample, 
#calculate the posterior based on what the prior is
#and find the minimum above threshold
for (s in 1:1e4){
  lik <- jointEvidence[s] 
  denomin <- lik * priorH0 + prior
  num <-  lik * priorH0
  posterior <- 1- num/denomin
  jointPosterior[[s]] <- posterior
  minima[s] <- min(prior[posterior > .99])
  }



minimaPlot <- ggplot()+geom_density(aes(x = minima))+
  theme_tufte(base_size = 10)+ggtitle("Minimal priors sufficient for posterior >.99")+xlab("minimal prior")+
  theme(plot.title.position = "plot")

minimaGrob <- ggplotGrob(minimaPlot)



jointPosteriorDF <-   as.data.frame(do.call(cbind, jointPosterior))



jointPosteriorSubsample <- jointPosteriorDF[,1:300]

jointPosteriorDF$ps <- ps
jointPosteriorSubsample$ps <- ps

jointPosteriorLong <- melt(jointPosteriorSubsample,
                           id.vars = c("ps"))
colnames(jointPosteriorLong) <- c("prior", "sample","probability")

alpha = .25
size = .2

densitiesLinesPlot <- ggplot(jointPosteriorLong)+
  geom_line(aes(x = prior, y = probability,group = sample),
            alpha = alpha, linewidth = size)+
  theme_tufte()+xlim(0,.2) + 
  annotation_custom(minimaGrob, xmin = .085, xmax = .185, 
                  ymin = 0.01, ymax = 0.8) +
  ggtitle("Posterior vs prior (300 sampled lines)") + 
  theme(plot.title.position = "plot")

```




\begin{figure}[H]
```{r fig:lines5,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "60%", warning = FALSE, message = FALSE}
densitiesLinesPlot
```

\caption{300 lines illustrating the uncertainty about the dependence of the posterior on the prior given aleatory uncertainty about the evidence, with the distribution of the minimal priors required for the posterior to be above .99.}

\label{fig:lines}

\end{figure}


This, then, is the main claim of this chapter: whenever density estimates for the probabilities of interest are available (and they should be available for match evidence and many other items of scientific evidence if the reliability of a given type of evidence has been properly studied), those densities should be reported for assessing the strength of the evidence. This approach avoids hiding actual aleatory uncertainties under the carpet. It also 
allows for a balanced assessment of the evidence, whereas using point estimates  or intervals may exaggerate or underestimate the value of the evidence. 


In what follows, we  expand on this idea in different directions. Section \ref{sec:three-probabilism} engages with the philosophical debate about precise and imprecise probabilism. We argue that both options are problematic and should be superseded by a higher-order approach to probability whenever possible. Section \ref{sec:objections}
revisits a recent discussion in the forensic science literature. A prominent view has it  that trial experts, even when they use densities, should present only first-order probabilities. We disagree and show that reasons of accuracy maximization sometimes recommend relying on higher-order probabilities. Section \ref{sec:legal-applications} turns to some legal applications of higher-order probabilism. We focus on two topics: first, the role of higher-order probabilities and false positive rates in the evaluation of DNA evidence; second, how complex bodies of evidence can be represented by what we call higher-order Bayesian networks. 

Before we dive in, one more remark: ost of the time, mathematically, we do not propose anything radically new---we just put together some of the items from the standard Bayesian toolkit. The novelty is rather in our arguing that that these tools are under-appreciated in the legal scholarship and should be properly used to incorporate second-order uncertainties in evidence evaluation and incorporation. Perhaps a minor exception is our explication of the notion of weight, but even here many related notions are available in information theory, and the novelty here is not technical, but rather in the argument that they also are under-appreciated in legal scholarship.
\todo{added this par to preemt Kadane's style pickiness}









# Three probabilisms
\label{sec:three-probabilism}

 In introduction we sketched three probabilistic approaches that one might take for assessing the value of the evidence presented at trial. The first approach uses precise probabilities; the second uses intervals; the third uses distributions over probabilities. By relying on an example featuring two items of match evidence, we suggested that the third approach is preferable. This section buttresses this claim by providing principled, philosophical reasons in favor of the third approach.  

The three approaches we considered correspond (roughly) to three ways in which probabilities can be deployed to model a rational agent's fallible and evidence-based beliefs about the world.  The first approach, known in the philosophical literature as precise probabilism, posits that an agent's credal state is modeled by a single, precise probability measure. The second approach, known as imprecise probabilism, replaces precise probabilities by sets of probability measures. The third approach, what we call higher-order probabilism, relies on distributions over parameter values.  There are good reasons to abandon precise probabilism and endorse higher-order probabilism. Imprecise probabilism is a step in the right direction, but also suffers from too many difficulties of its own.


## Precise probabilism

Precise probabilism (\textsf{PP}) holds that a rational 
agent's  uncertainty about a hypothesis is to be 
represented as a single, precise probability measure. 
This is an elegant and simple theory. But representing our uncertainty about 
a proposition in terms of a single, precise probability runs into a number of difficulties. Precise probabilism fails to 
capture an important dimension of how 
our fallible beliefs reflect the evidence 
we have (or have not) obtained. A couple of stylized 
examples should make the point clear. (For the sake of simplicity, 
we will use examples featuring coins, but biases of coins can be thought of as  random match probabilities in the forensic context.)

<!---but the points are general and hold for all sampling based frequency estimation methods, including those of random match probability for various pieces of forensic evidence):--->

\begin{quote}
\textbf{No evidence v. fair coin}
You are about to toss a coin, but have no evidence 
whatsoever about its bias. You are completely ignorant. 
Compare this to the situation in which you know, 
based on overwhelming evidence, that the coin is fair. 
\end{quote}

\noindent On precise probabilism, both scenarios are represented by assigning a probability of .5 to the outcome \emph{heads}. If you are completely 
ignorant, the principle of insufficient evidence suggests that you assign .5 to both outcomes.  Similarly, if you know for sure the coin is fair, assigning .5 seems the best way to quantify the uncertainty about the outcome. The agent's evidence in the two scenario is quite different, but precise probabilities cannot capture this difference. 

\begin{quote}
\textbf{Learning from ignorance}
You toss a coin with unknown bias. You toss it 10 times and observe \emph{heads} 5 times. Suppose you toss it further and observe 50 \emph{heads} in 100 tosses. 
\end{quote}

\noindent
Since the coin initially had unknown bias, you should 
presumably assign a probability of .5 to both outcomes. 
After the 10 tosses, you end up again with an estimate of .5.
You must have learned something, but whatever that is, it is not modeled by precise probabilities. When you toss the coin 100 times and observe 50 heads, you learn something. But your precise probability assessment will again be .5.

<!---Clearly,  precise probabilism has difficulties modeling such  situations. ---> These  examples suggest that precise probabilism is not appropriately responsive to evidence. It ends up assigning the same probability in  situations in which one's evidence is quite different:  when no evidence is available about the coin's bias; when there is little evidence that the coin is fair (say, after only 10 tosses); and  when there is strong evidence that the coin is fair (say, after 100 tosses). The general problem is, precise probability captures the value around which your uncertainty should be centered, but fails to capture how centered it should be given the evidence.^[Precise probabilism suffers from other difficulties. For example,  it has problems  with formulating a sensible method of probabilistic opinion  aggregation [@Elkin2018resolving,@Stewart2018pooling].  A seemingly intuitive constraint is that if every member agrees that $X$ and $Y$ are probabilistically independent, the aggregated credence should respect this. But this is hard to achieve if we stick to \s{PP} [@Dietrich2016pooling]. For instance, a \emph{prima facie} obvious method of linear pooling does not respect this. Consider probabilistic measures $p$ and $q$ such that $p(X)  = p(Y)  = p(X\vert Y) = 1/3$ and  $q(X)  =  q(Y) = q(X\vert Y) = 2/3$. On both measures, taken separately, $X$ and $Y$ are independent. Now take the average, $r=p/2+q/2$. Then $r(X\cap Y) = 5/18 \neq r(X)r(Y)=1/4$. ]


## Imprecise probabilism

What if we give up the assumption that probability assignments should be precise? Imprecise probabilism (\textsf{IP}) holds that  an agent's credal stance towards a hypothesis is to be represented by means of a *set of probability measures*, typically called a  representor $\mathbb{P}$, rather than a single measure $\mathsf{P}$. The representor should include all and only those probability measures which are compatible
<!-- (in a sense to be specified)  -->
with the evidence.  For instance, if an agent knows that the coin is fair, their credal state  would  be represented by the singleton set $\{\mathsf{P}\}$, where $\mathsf{P}$ is a probability measure which assigns $.5$ to \emph{heads}. If, on the other hand, the agent knows nothing about the coin's bias, their credal state would be represented by the set of all probabilistic measures, since none of them is excluded by the available evidence. Note that the set of probability measures does not represent admissible options that the agent could legitimately pick from. Rather, the agent's credal state is essentially imprecise and should be represented by means of the entire set of probability measures.^[For  the development of imprecise probabilism, see @keynes1921treatise; @Levi1974ideterminate; @Gardenfors1982unreliable; @Kaplan1968decision;  @joyce2005probabilities; @VanFraassen2006vague; @Sturgeon2008grain; @walley1991statistical. @bradley2019imprecise is a good source of further references. Imprecise probabilism shares some similarities with what we might call **interval probabilism** [@Kyburg1961; @kyburg2001uncertain]. On interval probabilism, precise probabilities are replaced  by intervals of probabilities. On imprecise probabilism, instead, precise probabilities are replaced by sets of probabilities.  This makes imprecise probabilism more general, since the probabilities of a proposition in the representor set do not have to form a closed interval. <!---Moreover, learning on Kyburg's approach is somewhat  idiosyncratic and is strongly connected to reference classes and selection and reshaping rules for intervals. See [PEDDEN] for a introduction.---> As we have already noted, intervals do not contain probabilistic information sufficient to guide reasoning with multiple items of evidence. So we focus on \s{IP}, which is the more promising approach.] 


Imprecise probabilism, at least \emph{prima facie}, offers a straightforward picture of learning from evidence, that is a natural extension of the classical Bayesian approach. When faced with new evidence $E$ between time $t_0$ and $t_1$, the representor set  should be updated point-wise,  running the  standard Bayesian updating on each probability measure in the representor:
\begin{align*} \label{eq:updateRepresentor}
\mathbb{P}_{t_1} = \{\mathsf{P}_{t_1}\vert \exists\, {\mathsf{P}_{t_0} \!\in  \mathbb{P}_{t_0}}\,\, \forall\, {H}\,\, \left[\mathsf{P}_{t_1}(H)=\mathsf{P}_{t_0}(H \vert E)\right] \}.
\end{align*}

\noindent
The hope is that, if we start with a range of probabilities that is not extremely wide, point-wise learning will behave appropriately.
<!---^[The hope is also that \s{IP} offers a feasible aggregation method [@Elkin2018resolving;@Stewart2018pooling]: just put all representors together in one set, and voil\'a! However, this is a very conservative method which quickly leads to extremely few points of agreement, and we are not aware of any successful practical deployment of this method.]
--->For instance, if we start with a prior probability of \emph{heads} equal to .4 or .6, then those measure should be updated to something closer to $.5$ once we learn that a given coin has already been tossed  ten times with the observed number of heads equal 5 (call this evidence $E$). This would mean that if the initial range of values was $[.4,.6]$ the posterior range of values should be more narrow. But even this seemingly straightforward piece of reasoning is hard to model without using densities. For to calculate $\pr{\s{heads}\vert E}$ we need to calculate $\pr{E \vert \s{heads}}\pr{\s{heads}}$ and divide it by $\pr{E} = \pr{E \vert \s{heads}}\pr{\s{heads}} + \pr{E} = \pr{E \vert \neg \s{heads}}\pr{\neg \s{heads}}$. The tricky part is obtaining the conditional probabilities $\pr{E \vert  \s{heads}}$ and $\pr{E \vert \neg \s{heads}}$ in a principled manner without explicitly going second-order, estimating the parameter value and using beta distributions. 


The situation is even more difficult if we start with complete lack of knowledge, as imprecise probabilism runs into the problem of **belief inertia**  [@Levi1980enterprise]. Say  you start tossing a coin  knowing nothing about its bias. The range of possibilities is $[0,1]$. After a few tosses, if you observed at least one tail and one heads, you can exclude the measures assigning 0 or 1 to \emph{heads}. But what else have you learned? <!---Let's agree that each particular measure from your initial representor set is updated to a measure closer to .5, but also now each value in your original interval  can be obtained by updating   some *other* measure in your original representor on the evidence, and the picture does not change  no matter how many observations you have made.  For instance, some measure that initially assigned .4 to \s{heads} might now assign \s{.45} to \s{heads}, but now a measure that assigned .37 to heads has been updated to one that assigns $.4$ to \s{heads}.---> If you are to update your representor set point-wise, you will end up with the same representor set.  Consequently, the edges of your resulting interval will remain the same. In the end, it is not clear how you are supposed to learn anything if you start from complete ignorance.^[Here's another example from @Rinard2013against. Either all the marbles in the urn are green ($H_1$), or exactly one tenth of the marbles are green ($H_2$). Your initial credence $[0,1]$ in each. Then you learn that a marble drawn at random from the urn is green ($E$). After conditionalizing each function in your representor on this evidence, you end up with the  the same spread of values for $H_1$ that you had  before learning $E$, and  no matter  how many marbles are sampled from the urn and found to be green.]

<!---Some replies in defense of imprecise probabilism are available.--->
Some downplay the problem of belief inertia. They insist that vacuous priors should not be used and that imprecise probabilism gives the right results when the priors are non-vacuous. After all, if you started with knowing truly nothing, then perhaps it is right to conclude that you will never learn anything. Another strategy is to say that, in a state of complete ignorance, a special updating rule should be deployed.^[@Lee2017impreciseEpistemology suggests the rule of  \emph{credal set replacement} that recommends that upon receiving evidence the agent should  drop measures rendered implausible, and add all non-extreme plausible probability measures. This, however, is tricky. One  needs a separate account of what makes a distribution plausible or not, as well as a principled account of why one should use a separate special update rule when starting with complete ignorance.] But no matter what we think about belief inertia, other problems plague imprecise probabilism.  Two  more problems are particularly pressing. 


<!---Elkin admits that he has no solution to this: "But how do we determine what the set of plausible probability measures is relative to $E$? There is no precise rule that I am aware of for determining such set at this moment, but I might say that the set can sometimes be determined fairly easily" [p. 83] He goes on to a trivial example of learning that the coin is fair and dropping extreme probabilities. This is far from a general account.--->


<!-- \textbf{Marcello's comment: I did not add the bit about how we learn about constraints in imprecise probabilism. Might need to go here as a further difficulty about imprecise probabilism. I could not understand the argment. To be discussed.} -->

One problem is that imprecise probabilism fails to capture intuitions 
we have about evidence and uncertainty in a number of scenarios. Consider this example:

\begin{quote}
\textbf{Even v. uneven bias:}
 You have two coins and you know, for sure, that the probability of getting heads is .4, if you toss one coin, and .6, if you toss the other coin. But you do not know which is which. You pick one of the two at random and toss it.  Contrast this with an uneven case. You have four coins and you know that three of them have bias $.4$ and one of them has bias $.6$. You pick a coin at random and plan to toss it. You should be three times more confident that the probability of getting heads is .4. rather than .6.
\end{quote}

\noindent
The first situation can be easily represented by imprecise probabilism. The representor would contain two probability measures, one that assigns .4. and the other that assigns .6 to the hypothesis 'this coin lands heads'.  But imprecise probabilism cannot represent the second situation, at least not without moving to higher-order probabilities or assigning probabilities to chance hypotheses, in which case it is no longer clear whether the object-level imprecision performs any valuable task.^[Other scenarios can be constructed in which imprecise probabilism fails to capture distinctive intuitions about evidence and uncertainty; see, for example, [@Rinard2013against].  Suppose you know of two urns, \textsf{GREEN} and \textsf{MYSTERY}. You  are  certain \textsf{GREEN} contains only green marbles, but have no information about \textsf{MYSTERY}. A marble will be drawn at random from each. You should be certain that the marble drawn from \textsf{GREEN}  will be green ($G$), and you should be more confident about this than about the proposition that the marble from \textsf{MYSTERY} will be green ($M$). In line with how lack of information is to be represented on \textsf{IP}, for each $r\in [0,1]$  your representor contains a $\mathsf{P}$ with $\pr{M}=r$. But then,  it also contains one with $\pr{M}=1$. This means that it is not the case that for any probability measure  $\mathsf{P}$ in your representor, $\mathsf{P}(G) > \mathsf{P}(M)$, that is, it is not the case that RA is more confident of $G$ than of $M$. This is highly counter-intuitive.] 

Second, besides descriptive inadequacy, an even deeper, foundational problem exists for imprecise probabilism. This problem arises when we attempt to measure the accuracy of a representor set of probability measures. Workable *scoring rules* exists for measuring the accuracy of a single, precise credence function, such as the Brier score. These rules measure the distance between one's credence function (or probability measure) and the actual value. A  requirement of scoring rules is that they  be \emph{proper}: any agent will score their own credence function to be more accurate than every other credence function. After all, if an agent thought a different credence was more accurate, they should switch to it.  Proper scoring rules are then used to formulate accuracy-based arguments for precise probabilism. These arguments show (roughly) that, if your precise credence follows the axioms of probability theory, no other credence is going to be more accurate than yours whatever the facts are.  Can the same be done for imprecise probabilism? It seems not. Impossibility theorems demonstrate that no proper scoring rules are available for representor sets. So, as many have noted, the prospects for an accuracy-based argument for imprecise probabilism look dim  [@seidenfeld2012forecasting; @Mayo-Wilson2016scoring; @Schoenfield2017accuracy; @CampbellMoore2020accuracy]. Moreover, as shown by  @Schoenfield2017accuracy, if an accuracy measure satisfies certain plausible formal constraints, it will never strictly recommend an imprecise stance, as for any imprecise stance there will be a precise one with at least the same accuracy. 


<!---
The third problem with imprecise probabilism has 
to do with spelling how when a probability measures is incompatible with the evidence. For recall that the probability measures allowed in a representor set are only those compatible with the agent's evidence. The idea is that thanks to this feature, imprecise credal stances are evidence-responsive in a way precise probabilistic stances are not. But how, exactly, does the evidence exclude probability measures?  This is not a  mathematical question: mathematically [@bradley2012scientific], evidential constraints are easy to model. They can take the form, for example, of the \emph{evidence of chances} $\{ \mathsf{P}(X) = x\}$ or $\mathsf{P}(X) \in [x,y]$, or \emph{structural constraints} such as  "$X$ and $Y$ are independent" or "$X$ is more likely than $Y$." While it is clear that these constraints are something that an agent can come to accept if offered such information by an expert to which the agent completely defers, it  is not trivial to explain how  non-testimonial evidence can result in such constraints.  Most of the  examples  in the literature start with the assumption that the agent is told by a believable source that the chances are such-and-such, or that the experimental set-up is such that the agent  knows that such and such structural constraint is satisfied. But, besides ideal circumstances, it is unclear how an agent could come to accept such structural constraints upon observation.^[Even if you were lucky enough to run into an expert that you completely trust that provides you with a constraint like this, how exactly did the expert come to learn the constraint? The chain of testimonial evidence has to end somewhere. Admittedly, there are straightforward degenerate cases: if you see the outcome of a coin toss to be heads, you reject the measure with $\mathsf{P}(H)=0$, and similarly for tails. Another class of cases might arise if you are randomly drawing objects from a finite set where the real frequencies are already known, because this finite set has been inspected. But such extreme cases aside, what else?  Mere consistency constraint wouldn't get the agent very far in the game of excluding probability measures, as way too many probability measures are strictly speaking still consistent with the observations for evidence to result in epistemic progress.] 
--->

<!---
Bradley suggests that "statistical evidence might inform [evidential] constraints [\dots and that evidence] of causes might inform structural constraints" [125-126]. This, however, is far cry from a clear account of how exactly this should proceed. Now, one suggestion might be that once a statistical significance threshold is selected, a given set of observations with a selection of background modeling assumptions yields a credible interval. But this is to admit that to reach such constraints we already have to start with a second-order approach, and drop information about the densities, focusing only on the intervals obtained with fixed margins of errors. But as we already illustrated, if you have the information about densities to start with, there is no clear advantage to going imprecise instead, and there are multiple problems associated with this move. Moreover, such moves require a choice of an error margin, which is extra-epistemic,^[In forensic evidence evaluation even scholars who disagree about the value of going higher-order agree that interval reporting is problematic, as the choice of a limit or uncertaintly level is rather arbitrary [@Taroni2015Dismissal;Sjerps2015Uncertainty].] and it is not clear what advantage there is to dropping information contained in second-order probabilities based on extra-epistemic considerations of this sort. 
--->


## Higher-order probabilism

There is, however, a view in the neighborhood that fares better: a second-order perspective. In fact, some of the comments by the proponents of imprecise probabilism tend to go in this direction. For instance,  Seamus Bradley   compares the measures in a representor to committee members, each voting on a particular issue, say the true bias of a coin. As they acquire more evidence, the committee members will often converge on a specific chance hypothesis. He writes [@bradley2012scientific, p. 157]:
\begin{quote}
\dots the committee members are ''bunching up''. Whatever measure you put over the set of probability functions---whatever ''second order probability'' you use---the ''mass'' of this measure gets more and more concentrated around the true chance hypothesis'.
\end{quote}

\noindent
Note, however, that such bunching up cannot be modeled by imprecise probabilism. <!---^[Bradley seems to be aware of that, which would explain the use of scare quotes: when he talks about the option of using second-order probabilities in decision theory, he  insists that  'there is no justification for saying that there is more of your representor here or there.' ~[p.~195]]--->   @joyce2005probabilities, in a paper defending imprecise probabilism, in fact uses a density over chance hypotheses to account for the notion of evidential weight.
<!-- (on this, more later).  TODO -->
<!---and conceptualizes the  weight of evidence  as an increase of concentration of smaller subsets of chance hypotheses, without any reference to representors in his explication of the notion of weight (we will get back to his explication when we discuss weight of evidence). ---> 
The idea that one should use  higher-order probabilities  has also been suggested by critics of imprecise probabilism. For example, @Carr2020impreciseEvidence argues that sometimes   evidence requires uncertainty about what credences to have.  <!---On Carr's approach, one should use vague credences, assigning various weights to probabilities---agent's credence in propositions about either what credences the evidence supports, or about objective chances. ---> Carr, however, does not articulate this suggestion more fully, does not develop it formally, and does not explain how her approach would fare against the difficulties affecting precise ad imprecise probabilism.


The key idea of the higher-order approach we propose is that uncertainty is not a single-dimensional thing to be mapped on a single one-dimensional scale such as a real line. It is the whole shape of the whole distribution over parameter values that should be  taken under consideration.^[Bradley admits this much [@bradley2012scientific, 90], and so does  Konek  [@konek2013foundations, 59]. For instance, Konek disagrees with: (1)  $X$ is more probable than $Y$ just in case $p(X)>p(Y)$, (2)  $D$ positively supports $H$ if $p_D(H)> p(H)$, or (3)  $A$ is preferable to $B$ just in case the expected utility of $A$ w.r.t. $p$ is larger than that of $B$.] From this perspective, when an agent is asked about their credal stance towards $X$, they can refuse to summarize it in terms of a point value $\mathsf{P}(X)$. They can instead express their credal stance in terms of a probability (density) distribution $f_x$ treating  $\mathsf{P}(X)$ as a random variable.  To be sure, an agent's credal state toward $X$ could sometimes be usefully represented by the expectation  
\[\int_{0}^{1} x f(x) \, dx\] 
as the precise, object-level credence in $X$, where $f$ is the probability density over possible object-level probability values. But this need not always be the case. If the probability density $f$ is not sufficiently concentrated around a single value, a one-point summary might fail to do justice to the nuances of the agent's credal state.^[This approach lines up with common practice in Bayesian statistics, where the primary role of uncertainty representation is assigned to the whole distribution. Summaries such as the mean, mode standard deviation,  mean absolute deviation, or highest posterior density intervals are only succinct ways for representing the uncertainty of a given scenario. Whether the expectation should be used in betting behavior is a separate problem. Here we focus on epistemic issues.]   For example, consider again the scenario in which the agent knows that the bias of the coin is  either .4 or .6 but the former is three times more likely. Representing the agent's credal state with the expectation $\mathsf{P}(X) = .75 \times .4 + .25 \times .6 = .45$ would be inadequate as it would fail to capture the agent's belief that the two biases are uneven. 

The higher-order approach can easily model all the challenging scenarios we discussed so far in the manner illustrated in Figure \ref{fig:evidenceResponse}. In particular, the scenario in which the two biases of the coin are not equally likely---which imprecise probabilism cannot model---can be easily modeled within high-order probabilism by assigning different probabilities to the two biases.
  

```{r evidenceResponse1,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "100%",   message = FALSE, warning = FALSE, results = FALSE}
p <- seq(from=0 , to=1 , by = 0.001)
FairDensity <- ifelse(p == .5, 1, 0)
FairDF <- data.frame(p,FairDensity)
FairPlot <- ggplot(FairDF, aes(x = p, y = FairDensity))+geom_line()+theme_tufte()+
  xlab("parameter value")+ylim(c(0,1))+ylab("probability")+ggtitle("Fair coin")+
  theme(plot.title.position = "plot")

UniformPlot <- ggplot()+xlim(c(0,1))+stat_function(fun = dunif, args = list(min = 0, max = 1))+
  ylim(0,1)+
  theme_tufte()+
  xlab("parameter value")+
  ylab("probability density")+theme(plot.title.position = "plot")+ggtitle("Unknown bias")


p <- seq(from=0 , to=1 , by = 0.001)
TwoDensity <- ifelse(p == .4 | p == .6, .5, 0)
TwoDF <- data.frame(p,TwoDensity)
TwoPlot <- ggplot(TwoDF, aes(x = p, y = TwoDensity))+geom_line()+theme_tufte()+
  xlab("parameter value")+ylim(c(0,1))+ylab("probability")+ggtitle("Two biases (insufficient reason)")+
  theme(plot.title.position = "plot")+scale_x_continuous(breaks = c(.4,.6), labels = c(.4, .6))




p <- seq(from=0 , to=1 , by = 0.001)
TwoUnbalancedDensity <- ifelse(p == .4, .75, ifelse(p ==  .6, .25, 0))
TwoUnbalancedDF <- data.frame(p,TwoUnbalancedDensity)

TwoUnbalancedPlot <- ggplot(TwoUnbalancedDF, aes(x = p, y = TwoUnbalancedDensity))+geom_line()+theme_tufte()+
  xlab("parameter value")+ylim(c(0,1))+ylab("probability")+ggtitle("Two unbalanced biases")+
  theme(plot.title.position = "plot")+scale_x_continuous(breaks = c(.4,.6), labels = c(.4, .6))

```


\begin{figure}[t]
```{r fig:evidenceResponse2,echo=FALSE,eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "80%"}
grid.arrange(FairPlot,UniformPlot, TwoPlot, TwoUnbalancedPlot)
```
\caption{Examples of higher-order distributions for scenarios brought up in the literature.}
\label{fig:evidenceResponse}
\end{figure}


Besides its flexibility in modelling uncertainty, higher-order probabilism does not fall prey to belief inertia. Consider a  situation in which you have no idea about the bias of a coin. So you start with a uniform density over $[0,1]$ as your prior. By using binomial probabilities as likelihoods, observing  any non-zero number of heads will exclude 0 and observing any non-zero number of  tails will exclude 1 from the basis of the posterior. The posterior distribution will become more centered around the parameter estimate as the observations come in. <!---Let's see an example with a grid approximation ($n=1k$) and coin tossing (grid approximation allows us also to talk about probabilities rather than densities).  Our prior is uniform, and then, in subsequent steps, we observe heads,  another heads, and then tails. This is what  happens with the posterior as we go (Figure \ref{fig:intertia2}).
--->
Figure \ref{fig:intertia2} shows---starting with a uniform prior distribution--- how the posterior distribution changes after successive observations of heads,  heads again, and then tails.^[More generally, learning about frequencies, assuming independence and constant probability for all the observations, is modeled the Bayes way. You start with some prior density $p$ over the parameter values. If you start with complete lack of information, $p$ should be uniform. Then, you observe the data $D$ which is the number of successes $s$ in a certain number of observations $n$. For each particular possible value $\theta$ of the parameter, the probability of $D$ conditional on $\theta$ follows the binomial distribution. The probability of $D$ is obtained by integration. That is:
\begin{align*}
p(\theta \vert D) & = \frac{p(D\vert \theta)p(\theta)}{p(D)}\\
& = \frac{\theta^s (1-\theta)^{(n - s)}p(\theta)}{\int (\theta')^s (1-\theta')^{(n - s)}p(\theta')\,\, d\theta'}.
\end{align*}
]





```{r fig:inertia2, eval = TRUE, echo=FALSE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%", message= FALSE, warning=FALSE}

n <- 1000 #parameters 
s <- 1e5  #sample size

ps <- seq(from=0 , to=1 , length.out=n)

prior <- rep(1/n , n) #uniform prior

likelihood1g <- dbinom( 1 , size=1 , prob=ps)
likelihood2g <- dbinom( 2 , size=2 , prob=ps)
likelihood2g1b <- dbinom( 2 , size=3 , prob=ps)

posterior1g <- likelihood1g * prior
posterior2g <- likelihood2g * prior
posterior2g1b <- likelihood2g1b * prior


posterior1g <- posterior1g / sum(posterior1g)
posterior2g <- posterior2g / sum(posterior2g)
posterior2g1b <- posterior2g1b / sum(posterior2g1b)

upperLimit <- .003

InertiaPriorPlot <- ggplot()+geom_line(aes(x = ps, y = prior))+theme_tufte(base_size = 7)+xlab("parameter")+
  ylab("probability")+theme(plot.title.position = "plot")+ggtitle("Prior")+
  scale_x_continuous()+scale_y_continuous(limits = c(0,upperLimit))


InertiaOneGPlot <- ggplot()+geom_line(aes(x = ps, y = posterior1g))+theme_tufte(base_size = 7)+xlab("parameter")+
  ylab("probability")+theme(plot.title.position = "plot",
                            axis.text.y = element_blank(),
                            axis.title.y = element_blank(),
                          axis.ticks.y =element_blank()
)+ggtitle("Evidence: h")+
  scale_y_continuous(limits = c(0,upperLimit))



InertiaTwoGPlot <- ggplot()+geom_line(aes(x = ps, y = posterior2g))+theme_tufte(base_size = 7)+xlab("p")+
  ylab("probability")+theme(plot.title.position = "plot"#, 
#                        axis.text.y = element_blank(),
#                            axis.title.y = element_blank(),
#                            axis.ticks.y = element_blank()
                  )+ggtitle("Evidence: h, h")+
scale_y_continuous(limits = c(0,upperLimit))

InertiaTwoGoneBluePlot <- ggplot()+geom_line(aes(x = ps, y = posterior2g1b))+theme_tufte(base_size = 7)+xlab("parameter")+
  ylab("probability")+theme(plot.title.position = "plot",     axis.text.y = element_blank(),
  axis.title.y = element_blank(),
      axis.ticks.y = element_blank()
           )+ggtitle("Evidence: h, h, t")+
  scale_y_continuous(limits = c(0,upperLimit))
```



\begin{figure}[t]
```{r fig:inertia3, echo=FALSE, eval=TRUE,fig.align = "center", cache=TRUE, fig.show = "hold", out.width = "80%",   message = FALSE, warning = FALSE, results = FALSE}
grid.arrange(InertiaPriorPlot, InertiaOneGPlot,  InertiaTwoGPlot, InertiaTwoGoneBluePlot, ncol = 2, nrow = 2)
```
\caption{As observations of heads, heads and tails come in, extreme parameter values drop out of the picture and the posterior is shaped by the evidence.}
\label{fig:intertia2}
\end{figure}


A further advantage of high-order probabilism over imprecise probabilism is that the prospects for accuracy-based arguments are not foreclosed. This is a significant shortcoming of imprecise probabilism, especially because such arguments exist for precise probabilism. One can show that there exist proper scoring rules for higher-order probabilism. These rules can then be used to formulate accuracy-based arguments. Another interesting feature of the framework is that the point made by Schoenfield against imprecise probabilism does not apply: there are cases in which accuracy considerations recommend an imprecise stance (that is, a multi-modal distribution) over a precise one (Urbaniak, 2022 manuscript).

All in all, higher-order probabilism outperforms both precise and imprecise probabilism, at the descriptive as well as the normative level. From a descriptive standpoint, higher-order probabilism can easily model a variety of scenarios that cannot be adequately modeled by the other versions of probabilism. From a normative standpoint, accuracy maximization may sometimes recommend that a rational agent represent their credal state with a distribution over probability values rather than a precise probability measure (more on this in the next section). 





# Objections 
\label{sec:objections}


This section addresses a number of conceptual difficulties that may arise in using higher-order probabilities, with focus on  those brought up by prominent legal evidence scholars. In discussing these conceptual issues, we will formulate an accuracy-based argument that higher-order probabilities are preferable to precise probabilities.


## The Taroni-Sjerps debate

Our treatment will be centered around a discussion initiated by 
@Taroni2015Dismissal, who argue extensively that trial experts should avoid report higher-order densities, and should only report point estimates. Their point of departure is a reflection on match evidence. 


Say an expert reports at trial that the sample from the crime scene matches the defendant. The significance of this match should be evaluated in light of the population frequency $\theta$ of the matching profile.  This frequency, however, cannot be known 
for sure and must instead be estimated. 


The expert will estimate the true parameter $\theta$ by means of a probability distribution $p(\theta)$ over its possible values. For example, if the observations are realizations of independent and identically distributed Bernoulli trails given $\theta$, the expert's uncertainty about $\theta$ can be modeled as $\s{beta}(\alpha + s + 1 ,\beta + n - s)$, where $s$ is the number of observed successes, $n$ the number of observations in the database (1 is added to the first shape parameter to include the match with the suspect), and $\alpha$ and $\beta$ reflect the expert's priors.

<!---However, they insist, in front of the fact-finders, the expert is supposed to report their epistemic uncertainty, which is to be understood according to the subjective interpretation.  On this interpretation, probabilities express an agent's epistemic attitude towards a proposition, their uncertainty about its truth value. Such probabilities, @Taroni2015Dismissal insist, unlike frequencies, are not states of nature, but states of mind associated with individuals. For this reason, they claim, it makes no sense to talk about second-order uncertainty about subjective probabilities, as there is no underlying state of the nature to estimate. Further, they insist this also applies to likelihood ratios:
\begin{quote}
\dots there is no meaningful state of nature equivalent for the likelihood ratio in its entirety, as it is given by a ratio of two conditional probabilities. [p. 12]
\end{quote}
\noindent In line with this perspectives, in the elicitation of probabilities @Taroni2015Dismissal recommend investigating an agent's betting preferences, and that a proper elicitation of this form will lead to a single number.^["Clearly,
one can adjust the measure of belief of success in the reference gamble in such
a way that one will be indifferent with respect to the truth of the event about
which one needs to give ones probability. This understanding is fundamental, as
it implies that probability is given by a single number. It may be hard to define,
but that does not mean that probability does not exist in an individuals mind. One cannot logically have two different numbers because they
would reflect different measures of belief." [p. 7]]
\noindent Moreover, while they claim  that it is fine to use distributions to talk about chance, they deny this possibility for  personal uncertainty, under the threat of infinite regress: 
\begin{quote}
One can, in fact, have probabilities for events, or probabilities for propositions, but not probabilities of probabilities, otherwise one would have an infinite regression. [p. 8]
\end{quote}
\noindent Accordingly,  @Taroni2015Dismissal insist that given a frequentist estimation of the probability of the evidence given the defense hypothesis, \dots
\begin{quote}
\dots personal beliefs [\dots] can be computed as:
\begin{align*}\pr{E} & = \int_{\theta} \pr{E\vert \theta} \pr{\theta}\,\, d\theta \\
& =  \int_\theta  \theta \pr{\theta}\,\, d\theta
\end{align*}
\end{quote}
\noindent In particular, for the DNA match example, they advice that the personal belief  is the expected value of the $\s{beta}$ distribution, which reduces to $\nicefrac{\alpha + s + 1}{\alpha + \beta +n + 1}$. They claim that it satisfactorily expresses the posterior uncertainty about $\theta$, and that it is  solely this probability that should be used in the denominator in the calculation and reporting of the likelihood ratio. 
--->

Nothing so far should be controversial. However, the question arises of how the expert should report their own uncertainty about $\theta$, especially in the light of the usual practice of reporting likelihood ratios.

To fix the notation, let the prosecution hypothesis $H_p$ be that the suspect is the source of the trace, and the defense hypothesis $H_d$ that another person, unrelated to the suspect, is the source. For simplicity, assume that if $H_p$ holds, the laboratory will surely report a match $M$, so that $\pr{M\vert H_p}=1$. 
The likelihood ratio, then,  reduces to $\nicefrac{1}{\pr{M \vert H_d}}$---but given that $\theta$ was estimated using density over its possible values, it is not obvious how a single value $\pr{M \vert H_d}$ is to be obtained and whether its use in the reporting does not hide the uncertainty involved in the estimation of $\theta$ under the carpet. 


@Taroni2015Dismissal claim  that the point estimate for the match evidence given the defense hypothesis should be calculated as follows:
\begin{align*}\pr{M \vert H_d} & = \int_{\theta} \pr{M\vert \theta} \pr{\theta}\,\, d\theta \\
& =  \int_\theta  \theta \pr{\theta}\,\, d\theta
\end{align*}
In case of a DNA match, they recommend that the expert report the expected value of the $\s{beta}$ distribution, which reduces to $\nicefrac{\alpha + s + 1}{\alpha + \beta +n + 1}$. They claim that this number satisfactorily expresses the posterior uncertainty about $\theta$. For them, it is this probability alone that should be used in the denominator in the calculation and reporting of the likelihood ratio. 


@Sjerps2015Uncertainty disagree. In reporting a single value, the expert would refrain from providing the fact-finders with relevant information  that can make a difference in the proper evaluation of the evidence.  There is a difference between (a) an expert who is certain $\theta$ is $.1$; (b) an expert whose best estimate of $\theta$ is $.1$ based on thousands of observations; and (c) an expert whose best estimate of $\theta$ is again $.1$ but based on only ten observations. 

These three scenarios mirror scenarios we discussed earlier: (a) the bias of a coin is known for sure; (b) the bias is estimated on the basis of a large number of tosses; and (c) the bias is estimated using a small set of observations. As our critique of precise probabilism makes clear, a simple point estimate (or precise probability) would fail to capture the differences among the three scenarios. This concern might be slightly mitigated by the fact that @Taroni2015Dismissal admits that the expert, besides providing a point estimate, should also informally explain how the estimate was arrived at. They grant that this additional information can be helpful so long as the recipients are instructed on "the nature of probability, the importance of an understanding of it and its proper use in dealing with uncertainty" [p. 16]. But why stop at an informal presentation? It is unclear why the fact-finders should be deprived of quantifiable information about the aleatory uncertainty of the parameter of interest and only be given an informal description of what the expert did, along with some remarks about the nature of probability. It is wildly optimistic to assume that an informal description of how the point estimate has been arrived at is enough to secure a proper assessment of the evidence. We hope to have convinced the reader already in the introduction that informal treatment and bare intuitions are not good enough  even when it  comes to the evaluation of the impact of a rather simple combination of  two items of evidence if all the fact-finder has to go by is point estimates and and informal description of how the estimates have been obtained.






Somewhat surprisingly, most of the concerns raised by  @Taroni2015Dismissal are  philosophical. They argue that if  probabilities express an agent's epistemic attitude towards a proposition probabilities are not states of nature, but states of mind associated with individuals. They think this claim  has two consequences. First, it makes no sense sense to talk about second-order uncertainty about subjective probabilities, as there is no "underlying state of the nature" to estimate. Second, if these subjective probabilities can be elicited by examining an agent's betting preferences, a proper elicitation will lead to a single number.^[They write: "Clearly, one can adjust the measure of belief of success in the reference gamble in such a way that one will be indifferent with respect to the truth of the event about which one needs to give ones probability. This understanding is fundamental, as it implies that probability is given by a single number. It may be hard to define, but that does not mean that probability does not exist in an individuals mind. One cannot logically have two different numbers because they would reflect different measures of belief." [@Taroni2015Dismissal, p. 7] ] 


In response to the philosophical argument, @Dahlman2022Information have also emphasized that the distinction is not so clear-cut. They argue that, if a probability assessment is a subjective attitude that is elicited via a betting preference, a probability assessment is itself a state of nature, "the formation of a betting preference by a certain person at a certain time" [p. 15]. While we will have something to say about the philosophical dimension of this debate, let us first develop  a less philosophically involved  argument for the position taken by @Sjerps2015Uncertainty.


## An ccuracy-based argument

 With this argument, we hope to break the stalemate in the debate by proving an argument to which both parties should be receptive. 
It is an accuracy-based argument in favor of using higher-order probabilities---roughly, it says, if you discard relevant information that you already have contained in the densities resulting from the estimation and rely on point estimates only, your predictions about the world will be less accurate in a very precise and quantifiable sense.

```{r,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
priorA <- 1
priorB <- 1
set.seed(215)
trueH <- runif(1,0,1) 

sampleSize <- sample(10:20,size = 1)
testSize <- sampleSize

set.seed(319)
successes <- rbinom(1, sampleSize, trueH)
pointEstimate <- successes/sampleSize

pointPredictions <-  rbinom( 1e4 , size = testSize , prob = pointEstimate )

ps <- seq(0,1,length.out = 1001)

testPredictions <- rbinom( 1e4 , size = testSize , prob = trueH )

posterior <- function(x) dbeta(x, priorA + successes,
                               priorB + sampleSize - successes)

posteriorSample <- sample( ps , size=1e4 ,
                           replace=TRUE , posterior(ps)/sum(posterior(ps)))

posteriorPredictions <- rbinom( 1e4 , size=testSize , prob=posteriorSample )


testProbs <- table(factor(testPredictions, levels = seq(0,sampleSize)))/1e4
pointProbs <- table(factor(pointPredictions, levels = seq(0,sampleSize)))/1e4
posteriorProbs <- table(factor(posteriorPredictions, levels = seq(0,sampleSize)))/1e4 

testPlot <- ggplot()+geom_bar(aes(x= testPredictions, y = ..prop..))+
  ggtitle(paste("Predictions based on the true parameter = ", round(trueH,2), sep = ""))+xlim(0,testSize)+ylab("probability")+xlab("successes")+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")


pointPlot <- ggplot()+geom_bar(aes(x= pointPredictions,y = ..prop..))+
  labs(title = paste("Predictions based on the point estimate = ", round(pointEstimate,2)), subtitle = paste(successes, " successes in ", sampleSize, " observations", sep = ""))+xlim(0,testSize)+ylab("probability")+xlab("successes")+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")


samplesPlot <- ggplot()+geom_density(aes(x= posteriorSample))+
  ggtitle(paste("Posterior sample from beta(", 1+successes, ",", 1+testSize - successes,
                ")", sep = ""))+xlab(
                  "parameter value"
                )+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")

posteriorPlot <- ggplot()+geom_bar(aes(x= posteriorPredictions,y = ..prop..))+
  ggtitle("Predictions based on the posterior sample")+xlim(0, testSize)+ylab("probability")+xlab("successes")+theme_tufte(base_size = 8)+theme(plot.title.position = "plot")


```

First, let us go over a particular example. Suppose we randomly draw a true population frequency from the uniform distribution. In our particular case, we obtained `r round(trueH,3)`. Then, we randomly draw a sample size as a natural number between 10 and 20. In our particular case, it is `r sampleSize`. Next, we simulate an experiment in which we draw that number of observations from the true distribution. We observe `r successes` successes and use this number to calculate the point estimate of the parameter, which is `r round(pointEstimate,3)`. 

What is the probability mass function (PMF) for all 
possible outcomes of an observation of the same size? 
Two PMF are initially relevant: first, the true probability mass based on the true parameter; second, the probability mass function based on the point estimate which is binomial around the point estimate. This latter PMF, however, does not take into account the uncertainty about the point estimate. To take this uncertainty seriously, continuing our example, we take a sample distribution of size 16 of possible parameter values from the posterior  $\s{beta}(1+\s{successes}, 1+\s{sample size} - \s{successes})$ distribution (we assume uniform prior for the sake of an example). Then, we use this sample of parameter values to simulate observations, one simulation for each parameter value in the sample. This simulation yields the so-called *posterior predictive distribution* (or posterior predictive PMF), which instead of a point estimate, propagates the uncertainty about the parameter value into the predictions about the outcomes of possible observations. Finally, we take simulated frequencies as our estimates of probabilities. This distribution is more honest about uncertainty and wider than the one obtained using the point estimate. The three PMFs are displayed in Figure \ref{fig:posteriorPrediction}.



\begin{figure}[H]
```{r fig:posteriorPrediction2, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "80%", warning=FALSE, message = FALSE, dpi = 800, warning = FALSE, message = FALSE, output = FALSE}
grid.arrange(testPlot,pointPlot,samplesPlot,posteriorPlot, ncol = 1)
```


\caption{Real probability mass, probability mass calculated using a point estimate, sampling distribution from the posterior, and the posterior predictive distribution based on this sampling distribution.}
\label{fig:posteriorPrediction}
\end{figure}

The PMF based on a point estimate is further off from the real PMF than the posterior predictive distribution. For instance, if we ask about the probability of the outcome being at least 9 successes, the true answer is `r mean(testPredictions >= 9)`, the point estimate PMF tells us it is `r mean(pointPredictions >= 9 )`, while the posterior predictive distribution gives a somewhat better guess at `r mean(posteriorPredictions >= 9 )`. A similar thing happens when we ask about the probability of the outcome being at most 9 successes. The true answer is `r mean(testPredictions <= 9)`, the point-estimate-based answer is `r mean(pointPredictions <= 9 )`, while the posterior predictive distribution yields  `r mean(posteriorPredictions <= 9 )`. More generally, we can use an information-theoretic measure, Kullback-Leibler divergence,  to quantify how far the point-estimate PMF and the posterior predictive PMF are from the true PMF.^[A bit of explanation of this divergence measure. Suppose we are dealing with a variable $X$ with $n$ distinct possible discrete states $x_1, \dots, x_n$ and consider two probability mass functions $p$ and $q$ which express uncertainty about the true value of $X$ so that, say, on $p$, $\pr(X=x_i)=p_i$. First, the uncertainty of a given distribution $p$, its **entropy**, is given by the sum of the logarithms of surprise $\nicefrac{1}{p_i}$ for all the possible values,  $H(p) = \sum x_i \log \frac{1}{p_i} = - \sum p_i \log p_i$.  Next, suppose events arise according to $p$, but we predict them
using $q$. The **cross-entropy**  is then  
$ \mathsf{H}(p, q) & = \sum p_i \log_(q_i)$. This value is going to be higher than the entropy of $p$ if $q$ is different from it. Think of it as the uncertainty involved in using $q$ to predict events that arise according to $p$. 

Third, **Kullback-Leibler**  divergence is the additional entropy introduced by
using $q$ instead of $p$ itself, that is, the difference between cross-entropy and entropy:
\begin{align*}
\mathsf{DKL}(p, q) & = H(p, q) - H(p)\\
&= - \sum p_i \log q_i   - \left(   - \sum p_i \log p_i \right) \\
& = - \sum p_i\left( \log q_i - \log p_i\right)\\
& =  \sum p_i\left( \log p_i - \log q_i\right)\\
& = \sum p_i \log \left( \frac{p_i}{q_i}\right)
\end{align*} \noindent  As it turns out, KL divergence is also the expected
difference in log probabilities. In particular, if
\(p=q\) we get
\(DKL(p,p) = \sum p_i (\log_2 p_i - \log_2 p_i) = 0\),
which works out as it intuitively should be.]


 In our particular case, the former distance is `r kld(testProbs,pointProbs)` and the latter is `r kld(testProbs,posteriorProbs)`. <!---That is, the posterior predictive distribution is information-theoretically closer to the true distribution.--->

```{r kldsCalculations,echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "100%"}
klds <- numeric(1000)
for(i in 1:1000){
trueH <- runif(1,0,1) 
sampleSize <- sample(10:25,size = 1)
testSize <- sampleSize
successes <- rbinom(1, sampleSize, trueH)
pointEstimate <- successes/sampleSize
pointPredictions <-  rbinom( 1e4 , size = testSize , prob = pointEstimate )
ps <- seq(0,1,length.out = 1001)
testPredictions <- rbinom( 1e4 , size = testSize , prob = trueH )
posteriorSample <- sample( ps , size=1e4 ,
                           replace=TRUE , posterior(ps)/sum(posterior(ps)))
posteriorPredictions <- rbinom( 1e4 , size=testSize , prob=posteriorSample )
testProbs <- table(factor(testPredictions, levels = seq(0,sampleSize)))/1e4
pointProbs <- table(factor(pointPredictions, levels = seq(0,sampleSize)))/1e4
posteriorProbs <- table(factor(posteriorPredictions, levels = seq(0,sampleSize)))/1e4 
klds[i] <- kld(testProbs,pointProbs) - kld(testProbs,posteriorProbs)
}

```


This result can be generalized.  We repeat the simulation 1000 times, each time with a new true parameter, a new sample size, and a new sample. Every time the three PMFs are constructed using the methods we described. <!---, and measure their KLD distances---> Figure \ref{fig:kldsPlots} displays the empirical distribution of the results of such a simulation, where the differences are measured by the Kullback-Leibler divergence. A positive difference indicates that the distribution based on the point-estimate was further from the true PMF than the posterior predictive distribution. Notably, the mean difference is `r round(mean(klds),3)`, the median difference is `r round(median(klds),3)`, and the distribution is asymmetrical, as there are multiple cases of huge differences favoring posterior predictive distributions over point-based predictions.  All in all, accuracy-wise, point-estimate-based PMFs are systematically worse than the posterior predictive distribution. 




\begin{figure}[H]
```{r fig:kldsPlots, echo=FALSE,eval=TRUE,fig.align = "center",cache=TRUE, fig.show = "hold", out.width = "70%", warning=FALSE, message = FALSE, dpi = 800, warning = FALSE, message = FALSE, output = FALSE}
ggplot()+geom_density(aes(x = klds))+labs(
  title = "Point-estimates vs. posterior predictive distributions",
  subtitle = "Differences in Kullback-Leibler divergencies from true PMFs")+
  theme_tufte(base_size = 10)+theme(plot.title.position = "plot")+ylab("empirical density")+xlab("difference in klds")
```
\caption{Differences in Kullback-Leibler divergencies from the true distributions, comparing the distributions obtained using point estimates and posterior predictive distributions. Positive values indicate the point-estimate-based PMF was further from the true distribution than the posterior predictive distribution.}
\label{fig:kldsPlots}
\end{figure}


## Conceptual issues




Accuracy considerations aside, we will now engage with the more  conceptual points. 
@Taroni2015Dismissal argue that since first-order probabilities capture your uncertainty about a proposition of interest, second-order probabilities are supposed to capture your uncertainty about how uncertain you are, and that "estimating" your first-order uncertainties is unnecessary.  They think that you can simply figure out your fair odds in a suitable bet on the proposition in question, and the fair odds track your unique, first-order uncertainty without any uncertainty about it. But this point can be questioned. For one thing, the betting interpretation of probability is not uncontroversial.^[See textbooks in formal epistemology [@bradley2015critical; @Titelbaum2020Fundamentals-of].] 
<!---we prefer to respond without assuming it, just working with the assumption that an agent's probabilities or densities are to represent or capture their uncertainty. So is estimation of one's uncertainty pointless?
-->
In addition, even granting the betting interpretation, 
<!---There are more controversial claims that philosophers have argued for than the infallibility of introspection. Those involved with epistemic logic often argue that it is not even the case that if an agent knows (or doesn't know) $p$, they also automatically know that they (don't) know $k$. Analogously, 
--->
it need not be implausible to say that we sometimes are uncertain about what we think the fair bets are or what our first-order uncertainties are.^[On a related score, the introspective axioms in epistemic logic---that is,  if an agent knows (or doesn't know) $p$, they also know that they (don't) know $p$---are by no means uncontroversial. See, for example, Williamson 2000 (chapter 5)'s argument against the KK principle of positive introspection.]
<!---This is one possible path of answering this.--->

Think again about an expert who gathers information about the allelic frequency $f$ of DNA matches in an available database, and starts with a defensible \s{beta} prior with parameters $\alpha, \beta$. Say the expert observes $s$ matches in a database of size $n$. So the population relative frequency the experts is estimating should follow the $\s{beta}(\alpha + s + 1 ,\beta + n - s)$ distribution. <!--- over the possible RMP values.---> So far, nothing controversial happens---the expert is estimating the relevant population frequency. Assuming the conditions are pristine (the expert has no modeling uncertainty, rules out laboratory errors, and so on), the beta distribution can be used to inform the expert's subjective uncertainty. But uncertainty about what? The (estimated) population frequency can serve to attach a probability to the proposition \emph{a match is observed  if another person, unrelated to the suspect, is the source of the trace}. Admittedly, if only this proposition is being considered, it is yet not clear what second-order uncertainties would be uncertainties about. But the expert also considers a continuum of propositions, each of the form \emph{the true population frequency is $\theta$} for each $\theta\in [0,1]$. A density over $\theta$ models the comparative plausibility that the expert assigns to such propositions in light of the evidence. <!--- and normalization allows them to calculate their subjective probabilities for $\theta$ belonging to various sub-intervals of $[0,1]$.---> So if one were worried that there were no propositions that the expert could be "second-order" uncertain about, there actually are plenty. In particular, if $\theta$ is a population frequency, gauging which density captures the extent to which the evidence justifies various estimates of that frequency is the same as gauging the comparative plausibility of the corresponding propositions about the population frequencies.^[Perhaps, this should no longer be called "estimation", but the the connection with estimation is strong enough to justify this terminology. In the end, this is a verbal discussion that we will not get into. <!---, language users' linguistic behavior will decide this for us anyway.--->]

More generally, evidence justifies first-order probability assignments to various degrees. Suppose there is no evidence about the bias of a coin. Then, each first-order uncertainty about it would be equally (un)-justified. (If you like to think in terms of bets, the evidence would give no reason to prefer any particular odds as fair.) If, instead, we know the coin is fair, the evidence clearly selects one preferred value, .5. (Again, if you like the betting metaphor, 1:1 would be the preferred betting odds.) But often the evidence is stronger than the former case and weaker than the latter case. Consider, for example, propositions about population frequencies in light of the results of observations. In such circumstances, the evidence justifies different values of first-order uncertainty to various degrees, and densities simply capture the extent to which different first-order uncertainties are supported by the evidence. <!---So what is it that you are estimating? The ideal evidential support you would get if you had all the evidence. Why is the estimate off? Because your evidence is limited and the best you can do is gauge your uncertainty going where this limited evidence leads you.--->

We conclude this section by examining two additional points raised by  @Taroni2015Dismissal. The first---which we already alluded to earlier---is that first-order probabilities  are not "states of nature" and so cannot be estimated.  <!---We are not sure about "nature" and why the requirement of being part of nature is important for estimation. ---> It is unclear why only states of nature can be estimated. Mathematicians use approximate methods to estimate answers to fairly abstract questions ,not obviously related to "states of nature", whatever these are.  So, estimation should make sense whenever there are some objective answers that we can approximate to a greater or lesser extent.<!---But if there is some objectivity to what the ideal evidence would support, or to the extent to which the actual evidence supports various competing hypotheses, we can be more or less wrong about such things, and so it is not implausible to say that  there is a clear sense in which we can estimate them.--->^[@Taroni2015Dismissal makes the same point for likelihood ratios. They argue that there is no "meaningful state of nature equivalent for the likelihood ratio in its entirety, as it is given by a ratio of two conditional probabilities?" But if it is meaningful to estimate two conditional probabilities (that is, frequencies in the population), or to  compare the relative plausibility of various propositions about them in terms of density, it is equally meaningful to estimate any function of the numbers involved.  Otherwise it would also be meaningless to try to estimate the body mass index (BMI) of an average 21 years old male student in the USA just because BMI is a ratio of other quantities.]

Second, @Taroni2015Dismissal argue that once we allow second-order probability, we run into the threat of infinite regress. But do we? Surely, they would agree that one can be uncertain about a statistical model. But this can be the case even if this model spits out a point estimate rather than a density. If you think the possibility of putting uncertainty on top of propositions about possible values of a first-order parameter leaves us in an  epistemically hopeless situation, you might have hard time explaining why your point estimation is in a better situation. After all, if asking further questions about probabilities up the hierarchy is always justified, we can keep asking about the probability of a point-estimate-spitting model, the probability of that probability, <!---(and the way we have reached it) being adequate---> and so on.  <!---Alternatively, we could concede that while our situation is often epistemically difficult, it is not hopeless. Of course, the mathematical and statistical models we use are exactly those: models, which can be more or less epistemically helpful. And when we decide which models to use, we always face a trade-off between various factors.---> 

Perhaps the problem at issue is just one of complexity. Admittedly, second-order estimation is more complex than relying on point estimates. But we hope to have convinced the reader this complexity is worth the effort. What about more complex models going third-order? If a workable approach can accomplish that---and the additional complexity pays off---we are all for going third-order. <!---In fact, some of the model selection methods can be thought of this way. ---> The fact that more complex models can always be built hardly lead us into a vicious infinite regress. Rather, it is an indication that our models of uncertainty can---in principle---always be improved. 























































#  References {-}